### 2015-02-10 papers/data_paper/datapaper.tex
first

\documentclass[preprint2]{emulateapj}

\usepackage{natbib}
\bibliographystyle{apj}
\usepackage{longtable}
\usepackage[]{graphicx}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{tabularx}
\usepackage{bm}

%% Sometimes a paper's abstract is too long to fit on the
%% title page in preprint2 mode. When that is the case,
%% use the longabstract style option.

%% \documentclass[preprint2,longabstract]{aastex}

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{gsavorgn@astro.swin.edu.au}
\newcommand{\fitfigurewidth}{0.8\textwidth}

\shorttitle{Data paper}
\shortauthors{Savorgnan et al.}

\begin{document}

\title{Data paper}

\author{G. A. D. Savorgnan\altaffilmark{1} and A. W. Graham\altaffilmark{1}}
\affil{Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia.}
\email{gsavorgn@astro.swin.edu.au}

%\and

%\author{A. W. Graham\altaffilmark{1}}
%\affil{Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia.}

%% Notice that each of these authors has alternate affiliations, which
%% are identified by the \altaffilmark after each name.  Specify alternate
%% affiliation information with \altaffiltext, with one command per each
%% affiliation.

%\altaffiltext{1}{Visiting Astronomer, Cerro Tololo Inter-American Observatory.
%CTIO is operated by AURA, Inc.\ under contract to the National Science
%Foundation.}
%\altaffiltext{2}{Society of Fellows, Harvard University.}
%\altaffiltext{3}{present address: Center for Astrophysics,
%    60 Garden Street, Cambridge, MA 02138}
%\altaffiltext{4}{Visiting Programmer, Space Telescope Science Institute}
%\altaffiltext{5}{Patron, Alonso's Bar and Grill}

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.

\begin{abstract}
abstract % 2d worse than 1d say it!!!
\end{abstract}

\keywords{keywords}

\section{Introduction}
\label{sec:int}
Supermassive black holes and host bulges have very different sizes. 
If the event horizon of the Galactic supermassive black hole was as big as a grain of sand of the Sahara desert, 
then the black hole gravitational sphere-of-influence would be as big as the international airport of Cairo, 
and the Galactic bulge would be as big as the Sahara desert itself.
It is thus surprising that the masses of supermassive black holes ($M_{\rm BH}$) 
scale with a number of properties of their host bulges, 
indicating the non-gravitational origin of these correlations 
(e.g.~\citealt{dressler1989,yee1992,kormendyrichstone1995,magorrian1998,ferraresemerritt2000,gebhardt2000,graham2001,
marconihunt2003,haringrix2004,graham2012bent}). {\bf AL: need to talk to you about Laor's work}\\
The tightness (or small scatter) of the observed black hole mass correlations has led to the idea
that black holes and host galaxies have coevolved with some sort of self-regulated growth.
Exploring the evolution of this growth with cosmic time could help identify
the driving mechanisms of the black hole - galaxy coevolution.
Observations at $z=0$ set the local benchmark from which to measure this evolution.
Any complete theory or model describing the coevolution of galaxies and black holes must incorporate 
all of the observed scaling relations, which also have to be consistent with each other.
%Modern semi-analytic models use the observed black hole mass scaling relations 
%to constrain the black hole accretion rate .
Modern hydrodynamical simulations, such as EAGLE \citep{schaye2015}, calibrate the feedback efficiencies 
to the $z=0$ black hole mass - galaxy mass relation.
Scaling relations can also be employed to predict the masses of black holes in other galaxies, where a direct
measure of $M_{\rm BH}$ would be extremely time consuming or simply impossible due to technological limitations.
Many accurate $M_{\rm BH}$ predictions would help derive 
the local black hole mass function (e.g.~\citealt{salucci1999,graham2007smbhmassfunction}) 
and space density (\citealt{grahamdriver2007smbhmassdensity} and references therein; \citealt{comastri2015}). \\
Since the first stellar dynamical detections of black holes were carried out in the '80s 
(see the references in the reviews by \citealt{kormendyrichstone1995} and \citealt{richstone1998}), 
the number of (direct) black hole mass measurements has increased with time 
and it has recently become a statistically meaningful sample -- of the order of 100 objects -- 
with which one can study SMBH demographics. 
It is now generally accepted that supermassive black holes reside at the center of most, if not all, 
massive galaxies, either quiescent or active. \\
Massive, early-type galaxies are often composite systems. 
The knowledge that many elliptical galaxies contain embedded stellar disks can be dated back to 
at least three decades ago 
\citep{capaccioli1987,carter1987,rixwhite1990,bender1990,scorzabender1990,nieto1991,rixwhite1992,scorzabender1995}. 
After examining long-slit spectroscopic observations of morphologically classified elliptical galaxies in the Fornax cluster, 
\cite{donofrio1995} and \cite{graham1998fornax} concluded that only $\sim$$1/3$ of their galaxies do not harbour any disk-like component.
Surveys with integral field spectrographs of early-type galaxies, such as the ATLAS$^{\rm 3D}$ Project \citep{cappellari2011}, 
have further changed our idea of ``elliptical'' galaxies, 
being all but simple and featureless pressure-supported systems.
More than half of all morphologically classified elliptical galaxies in the ATLAS$^{\rm 3D}$ sample are fast rotators \citep{atlas3dIII-MNRAS}.
\cite{krajnovic2013} showed that ``fast rotators'' as a class are simply disks or at least disk-like objects. 
In their magnitude-limited survey, systems 
without any signature of disk-like components (neither in the kinematics nor in the photometry) 
dominate only the most massive end (beyond $10^{11.5}~\rm M_\odot$) of the distribution. 
Given the prevalence of disks, it is important to perform bulge/disk decompositions, 
if one is to properly explore the black hole-bulge connection. \\
Measuring the photometric and structural properties of a galaxy's spheroidal component requires the ability to 
separate it from the rest of the galaxy. 
Galaxy decomposition is a parametric analysis that allows one to fit the surface brightness distribution
of galaxies using a combination of analytic functions (usually one function per
galaxy component, such as bulges, disks, bars, nuclei, etc.).
The one-dimensional (1D) technique involves fitting isophotes to the galaxy image, extracting the (one-dimensional) surface 
brightness radial profile and modeling it with a combination of analytic functions.
With the two-dimensional (2D) technique one fits analytic functions directly to digital images. \\
Over the past eight years, five independent studies 
\citep{grahamdriver2007,sani2011,vika2012,beifiori2012,lasker2014data,lasker2014anal} have attempted galaxy decomposition
in order to derive the bulge parameters and explore their relation with the black hole masses.
Interestingly, the past studies all %investigated the black hole mass scaling relations using 
used almost the same sample of galaxies, 
but they claimed some contradictory conclusions.
One study \citep{grahamdriver2007} obtained a good $M_{\rm BH}-n_{\rm sph}$ correlation
(the bulge S\'ersic index $n_{\rm sph}$ is a measure of the central radial concentration of stars), 
whereas the remaining four did not.
\cite{lasker2014anal} declared that $M_{\rm BH}$ correlates equally well with the total galaxy luminosity and the bulge luminosity, 
as opposed to \cite{beifiori2012}, who claimed that the bulge mass is a better tracer of $M_{\rm BH}$ than the galaxy mass.
The past studies did not converge to the same conclusions 
because their best-fit models for the same galaxy were often 
significantly different and not consistent with each other in terms of fitted components. 
Moreover, none of these studies attempted an individual galaxy-by-galaxy 
comparison of their models with the previous literature. 
We have now made this comparison and performed the optimal decompositions,
using $3.6~\rm \mu m$ \emph{Spitzer} satellite imagery, 
which is an excellent proxy for the stellar mass, superior to the $K$-band (\citealt{sheth2010} and references therein).
We will use these to obtain improved black hole mass scaling relations using the largest to date sample of 72 galaxies.

This paper is structured as follows.
Section \ref{sec:data} presents the data-set used to conduct this study. 
Section \ref{sec:anal} describes how we performed the galaxy decompositions, 
i.e.~how we identified and modeled the sub-components that constitute our galaxies. 
In Section \ref{sec:res} we outline the results from our analysis 
and we discuss the error analysis. 
Section \ref{sec:concl} summarizes our main conclusions. 
The individual galaxy decompositions are made available in the electronic version of this manuscript.

\section{Data}
\label{sec:data}
%\subsection{Initial sample}
Our initial galaxy sample (see Table \ref{tab:sample}) consists of 75 objects for which a dynamical detection of the black hole mass 
has been reported in the literature and for which at least one $3.6~\rm \mu m$ \emph{Spitzer}/IRAC observation was made publicy 
available at the time we began our investigation (2013).
Black hole masses were drawn from the catalog of \citet{grahamscott2013} for 70 galaxies,
from \citet{rusli2013bhmassesDM} for 4 galaxies 
and from \citet{greenhill2003} for 1 galaxy.

%\clearpage
\input{table_sample.tex}
%\noindent
\input{table_observations.tex}

\subsection{\emph{Spitzer}/IRAC observations}
\subsubsection{Data acquisition}
For each of our 75 galaxies, we downloaded from the 
Spitzer Heritage Archive\footnote{\url{http://irsa.ipac.caltech.edu/applications/Spitzer/SHA/}} 
all the available $3.6~\rm \mu m$ IRAC Astronomical Observation Requests (AORs), 
listed in Table \ref{tab:obs}.
Each AOR is an individual Spitzer observation sequence, and
includes a number of data frames (the individual exposures) and the calibration data.
The data frames were selected to be %``enhanced Level 1 products'', or
corrected Basic Calibrated Data (cBCD).
Basic Calibrated Data (BCD) are produced by the IRAC Level 1 pipeline. 
This automatic pipeline takes a single ``raw'' image, removes the scattered light, 
and performs dark subtraction, flatfielding correction and flux calibration (into units of $\rm MJy~sr^{-1}$). 
The final product (the BCD) is a flux-calibrated image which has had all the well-understood instrumental signatures removed.
BCD frames are further processed through an ``artifact correction'' pipeline 
that mitigates the commonly found artifacts of stray light, saturation, ``muxbleed'' and column pulldown\footnote{Stray light 
includes scattered light from stars outside the array location as well as filter ghosts from bright stars. 
Multiplexer bleed, or ``muxbleed'', can be generated by stars, hot pixels, and particle hits.
It looks like a decaying trail of pixels, repeating every fourth column.
Column pulldown is caused by a bright pixel that triggers a bias shift within its respective column, 
creating a lower background value throughout the entire column than in the surrounding columns.}.
After the artifact correction has been applied, the BCD becomes a cBCD.

\subsubsection{Mosaicing}
We performed image mosaicing using the %(command line version of the) 
MOPEX package \citep{makovozmarleau2005mopex}.
Individual cBCD frames with exposure time of $1~\rm sec$ were rejected.
Permanent or semi-permanent bad pixels, contained in a semi-static mask (the ``pmask''), were ignored.
Each AOR is associated to a specific pmask. 
Therefore, when multiple AORs were available for the same galaxy, 
we merged the different pmasks. 
Outlier rejection was performed with the dual outlier and multi-frame techniques.
The pixel size of the mosaic was set to be the same as the input cBCD frames ($1''.22 \times 1''.22$).
For $3.6~\rm \mu m$ observations with this pixel scale, 
the photometric zeropoint magnitude is $m_{\rm zp} = 17.26 \rm~mag$.
The orientation of the mosaic was set to the average rotation angle of the input cBCD frames. 
%{\bf TBD: not n-e oriented or re-sampled because problems with 2d PSF. }
Individual cBCD frames were combined together into a single mosaic with the default linear interpolation algorithm.

\subsubsection{Overlap correction}
Before generating a mosaic, MOPEX can perform background matching among the individual frames
by using the \emph{overlap} module. 
This module calculates and applies an additive correction to the individual frames,
producing a consistent background across the mosaicked image.
According to the \emph{Spitzer Data Cookbook}, the use of the \emph{overlap} module is not particularly recommended for 
$3.6~\rm \mu m$ observations. 
However, after a visual inspection of the mosaics obtained without activating the \emph{overlap} module,
we found that all the mosaics obtained from multiple AORs were affected by patchiness, 
due to bias fluctuations in the CCD array.
For this reason, we re-generated the multiple-AORs mosaics by activating the \emph{overlap} correction,
which successfully removed the ``chessboard'' pattern (see Figure \ref{fig:overlap} for an example).

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.49\columnwidth]{images/m49nooverlap.eps}
\includegraphics[width=0.49\columnwidth]{images/m49overlap.eps}
\caption{An example of the \emph{overlap} correction. 
The image mosaic of the galaxy M49 was obtained by co-adding frames coming from 8 different AORs.
The evident patchiness (left image) was removed (right image) using the \emph{overlap} module.}
\label{fig:overlap}
\end{center}
\end{figure}

\subsubsection{Sigma mosaics}
For each individual cBCD frame, the IRAC Level 1 pipeline calculates the uncertainty associated to each pixel
and produces an uncertainty frame (or sigma frame).
The initial uncertainty is estimated as the Poisson noise in electrons plus the readout noise added in quadrature
($\sigma^2 = \sigma^2_{\rm readoutnoise} + \sigma^2_{\rm Poisson}$).
This initial sigma frame is carried through the pipeline, and additional uncertainties (e.g.~dark current and flat field uncertainties) 
are added in quadrature when appropriate.
When MOPEX generates an image mosaic, it also produces the associated sigma mosaic 
by interpolating the individual uncertainty frames and co-adding them,
following the standard assumption of additive variances.

\subsubsection{Sky subtraction}
Sky subtraction was performed manually on the image mosaics using the tasks {\tt marksky} and {\tt skyfit} 
of the IRAF\footnote{IRAF is the Image Reduction and Analysis Facility, 
distributed by the National Optical Astronomy Observatory, 
which is operated by the Association of Universities for Research in Astronomy (AURA) 
under cooperative agreement with the National Science Foundation.} 
package GALPHOT\footnote{GALPHOT was developed in the IRAF - STSDAS environment mainly by 
W. Freudling, J. Salzer, and M. P. Haynes \citep{haynes1999galphot}.}.
The task {\tt skyfit} also provided an estimate of the sky root mean square ($RMS_{\rm sky}$).
%sky value also checked with Galfit

\subsubsection{Additional aestethic corrections}
The image mosaics of 4 galaxies (NGC 0821, NGC 2974, NGC 4291, NGC 4459) were found to be affected 
by bright, highly saturated stars lying close to the target galaxies.
These stars were modeled and subtracted using the software Galfit \citep{peng2010} 
and the extended IRAC Point Response Function (PRF) image at $3.6~\rm \mu m$.
This helped remove the extended wings and spikes of the saturated stars 
(see Figure \ref{fig:aesth} for an example).

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.49\columnwidth]{images/n4459brightstar.eps}
\includegraphics[width=0.49\columnwidth]{images/n4459nobrightstar.eps}
\caption{An example of aestethic correction.
The images show the mosaic of the galaxy NGC 4459 before (left) and after (right) the removal 
of a bright saturated star.}
\label{fig:aesth}
\end{center}
\end{figure}

\subsubsection{Image masking}
Galactic stars and any other objects different from the target galaxy 
were masked through a two-step procedure.
First, we created an initial mask using the IRAF task 
{\tt objmasks}, %\footnote{{\tt objmasks} is part of the {\tt noao.nproto} package.},
that identifies objects by threshold sigma detection.
Then, we refined each mask by hand, using the software 
SAOImage DS9\footnote{SAOImage DS9 development has been made possible by funding 
from the Chandra X-ray Science Center (CXC) 
and the High Energy Astrophysics Science Archive Center (HEASARC).} 
in conjunction with the IRAF task 
{\tt mskregions}. %\footnote{{\tt mskregions} is part of the {\tt proto} package.}.
We identified and carefully masked not only contaminating sources located in the field of the image mosaic,
but also objects overlapping with the target galaxy, such as foreground stars, globular clusters or red giant stars.

\subsubsection{1D PSF}
The one-dimensional Point Spread Function was charachterized using the IRAF task {\tt imexamine}.
A nonlinear least squares \citep{moffat1969} profile\footnote{The \citep{moffat1969} profile 
has the following form:
\begin{equation}
\mu = \mu_{\rm 0} - 2.5 \log \Biggl[ 1 + \biggl( \frac{R}{\alpha} \biggr)^2 \Biggr]^{-\beta} ,
\end{equation}
where $R$ is the projected galaxy radius, $\mu_{\rm 0}$ is the central surface brightness, and 
$\alpha$ and $\beta$ regulate the width and the shape of the profile.} 
of fixed center and zero background was fit to the 
(background subtracted) pixels of 20 bright stars, belonging to different image mosaics.
%We employed both a Gaussian and a Moffat \citep{moffat1969} profile. 
The best-fit parameters of the 20 stars were then averaged together. 
Doing so, we obtained %$FWHM = 2''.03$ for the Gaussian function,
$(\alpha, \beta) = (2''.38, 4.39)$ %for the Moffat function
(see Appendix \ref{sec:app1} for the analytical expressions of the Moffat function).

\subsubsection{2D PSF}
The IRAC support team provides the users with a two-dimensional instrument Point Response Function (PRF) at $3.6~\rm \mu m$.
However, we found this PRF to be inadequate for the purposes of our analysis.
In fact, the $FWHM$ of the IRAC instrument PRF ($\sim 1''.8$), as measured by the IRAF task {\tt imexamine}, 
is systematically smaller than the average $FWHM$ of ``real'' stars ($\sim 2''.0$). 
Figure \ref{fig:psf} illustrates this issue.
We also tested the IRAC PRF by providing it as the input Point Spread Function (PSF) for Galfit 
and fitting a number of stars in different image mosaics.
A visual inspection of the fit residuals confirmed that the IRAC instrument PRF is narrower than ``real'' point sources.
For this reason, we constructed our 2D PSF according to the following method 
(as directed by C. Peng, private communication).\\
We provided the IRAC instrument PRF as the input PSF for Galfit 
and simultaneously fit 7 bright stars (belonging to different mosaics), 
modeling the stars with Moffat profiles 
and constraining all the profiles to have the same $(\alpha, \beta)$, position angle and axis ratio.
The 2D PSF image was then obtained by taking the best fit Moffat model -- the same best-fit model for all 7 stars, by construction -- 
and convolving it with the IRAC instrument PRF.
The advantage of this method is to obtain a 2D PSF that is wider than the instrument PRF, 
but maintains the asymmetric features of the instrument PRF (e.g.~wings and spikes).
%This procedure was repeated using a Moffat profile in place of the Gaussian profile.
We then tested this 2D PSF 
on a number of stars (these stars were different from the 7 stars employed to build the 2D PSF image)
and verified that it correctly reproduces the shape of ``real'' point sources.

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\columnwidth, trim = 220 30 20 20, clip=True]{images/psf.eps}
%\includegraphics[width=1\columnwidth]{images/star1_cut.eps} \\
%\includegraphics[width=1\columnwidth]{images/PRF_IRACch1_129_129_cut.eps}
\caption{Normalized flux versus radial distance 
from the image centroid of an observed point source (top panel) 
and of the IRAC instrument PRF (bottom panel). 
The normalized flux is given in arbitrary units, 
and the radial distance is in units of pixel size of the IRAC detector (1 pixel = $1''.22$).
The blue solid line shows our 1D Moffat PSF model,
which has been normalized to intersect the data point with the largest flux. 
The $FWHM$ of the IRAC PRF is clearly smaller than that of a ``real'' star.
The inserts display the images of the observed point source and the IRAC PRF.}
\label{fig:psf}
\end{center}
\end{figure}

\subsection{Additional data}
\label{sec:adddata}
\subsubsection{Kinematics}
\label{sec:kinem}
A kinematical classification (slow/fast rotator) is available for 34 of our 75 galaxies 
from the ATLAS$^{\rm 3D}$ survey \citep{atlas3dIII-MNRAS} 
and for 3 additional galaxies from \citet{scott2014}.
This classification (Table \ref{tab:sample}, column 7) 
concerns the kinematic properties of galaxies within the instrument field-of-view, % (typically up to one effective radius),
but does not contain additional information -- crucial for our analysis -- about kinematic substructures, 
such as embedded disks or kinematically decoupled components, 
which can require separate modeling.
For this reason, we also visually inspected the velocity fields of our galaxies, 
when available from the literature.
Velocity maps were taken from the ATLAS$^{\rm 3D}$ survey for 34 galaxies 
(observed with SAURON by \citealt{krajnovic2011}),
from \citet{scott2014} for 2 galaxies 
(observed with WiFeS)
and from the SLUGGS survey for 12 galaxies 
(observed with DEIMOS by \citealt{arnold2014}).
While the field-of-views of SAURON ($33'' \times 41''$) and WiFeS ($25'' \times 38''$) 
reach to about one galaxy effective radius (for our local galaxies),
observations taken with DEIMOS can probe the galaxy kinematics well beyond two effective radii.

\subsubsection{AGNs and nuclear dust}
The X-ray, UV and optical radiation 
emitted by the the accretion disks of Active Galactic Nuclei (AGNs) can stimulate infrared thermal emission 
from circumnuclear dust, if present. 
This means that if a galaxy hosts an optical AGN \emph{and} a certain amount of nuclear dust,
we expected to detect some non-stellar nuclear emission at $3.6~\rm \mu m$.
It is therefore important to identify which of our galaxies have both an optical AGN and circumnuclear dust.
To help with this task, we searched NED\footnote{NED is the NASA/IPAC Extragalactic Database.} 
for the individual galaxies and their associated literature.
Unsurprisingly, dusty AGNs were more frequently found in late-type spiral galaxies.

\subsubsection{S\'ersic/core-S\'ersic classification}
\label{sec:corser}
Core-S\'ersic galaxies \citep{grahamguzman2003,graham2003coresersicmodel,trujillo2004coresersicmodel} are galaxies (or bulges) 
with partially depleted cores, 
i.e.~a central deficit of light relative to the inward extrapolation of their outer S\'ersic light profile. 
S\'ersic galaxies, instead, do not exhibit such central stellar deficits.
Partially depleted cores, as measured from high-resolution observations, have typical sizes of a few tens of parsecs.
The majority are thus unresolved in our image mosaics. 
We masked these unresolved cores by excluding the surface brightness profile 
within 3 PSF's $FWHM$ from the galaxy center.
In case of cores with sizes exceeding the PSF's $FWHM$, we excluded the data points within the size of the core 
plus 3 PSF's $FWHM$.
The S\'ersic/core-S\'ersic classification presented in this work (Table \ref{tab:sample}, column 5)
comes from the compilation of \citet{savorgnangraham2014},
who identified partially depleted cores according to the same criteria used by \citet{grahamscott2013}.
When no high-resolution image analysis was available from the literature, 
they inferred the presence of a partially depleted core based on the stellar velocity dispersion, $\sigma$:
a galaxy is classified as core-S\'ersic if $\sigma > 270\rm~km~s^{-1}$,
or as S\'ersic if $\sigma < 166\rm~km~s^{-1}$.
This resulted in us assigning cores to just 2 galaxies when using this alternative method 
if no high resolution image was available.

\section{Analysis}
\label{sec:anal}
\subsection{Isophotal analysis}
We performed an isophotal analysis of our galaxies using the IRAF task {\tt ellipse} \citep{taskellipse}, 
which fits elliptical isophotes to galaxy images.
The center of the isophotes was held fixed, while the ellipticity ($\epsilon$), the position angle ($P.A.$),  
and the amplitude of the fourth harmonic\footnote{The amplitude of the fourth harmonic deviations from perfect ellipses 
($B4$) parameterizes the diskyness ($B4>0$) or boxyness ($B4<0$) of the isophotes.} ($B4$) were allowed to vary with radius.
The step in semi-major axis length between successive ellipses was set to increase linearly, 
and then geometrically in our second run\footnote{In case of linear steps, 
the semi-major axis length for the next ellipse was calculated by adding $1$ pixel to the current length.
In case of geometric steps, the semi-major axis length for the next ellipse 
was calculated as $1.1$ times the current length.}.
As a result, for each galaxy we produced respectively a ``linearly sampled'' and a ``logarithmically sampled'' 
surface brightness profile along the major-axis. 
Major-axis surface brightness profiles were then projected along the equivalent-axis,
i.e.~the geometric mean of the major ($a$) and minor ($b$) axis ($R_{\rm eq} = \sqrt{ab}$), 
also called the circularized radius.
Isophotes corresponding to an intensity less than $3 \times RMS_{\rm sky}$ were rejected.
Some surface brightness profiles were truncated at our discretion before the $3 \times RMS_{\rm sky}$ limit, 
according to specific technical reasons (e.g.~contamination from light of a neighboring galaxy, 
disturbed morphology in the galaxy outskirts, etc.).
Individual cases are discussed in Section \ref{sec:indgal}.

\subsection{Unsharp masking}
Unsharp masking is an image sharpening technique useful to reveal asymmetric structures in galaxies, 
such as bars or (inclined) embedded disks. 
First, the original galaxy image was smoothed with a Gaussian filter.
Then, the original image was divided by the smoothed one. 
The result of such an operation is the ``unsharp mask''. 
The asymmetric features revealed by this technique have sizes 
comparable to the $FWHM$ of the Gaussian kernel used for the smoothing.
Therefore, for each galaxy, we produced a set of different unsharp masks by varying the $FWHM$ of the filter,
to identify all the asymmetric features that could bias the fitting process and may need modeling.
This information was used in combination with kinematic and AGN information discussed in Section \ref{sec:adddata}.

\subsection{1D fitting routine}
The decomposition of surface brightness profiles was performed with software written by G. Savorgnan.
This software can fit an observed surface brightness profile with any linear combination of 
a set of analytical functions (S\'ersic, exponential, Gaussian, Moffat, Ferrer, symmetric Gaussian ring etc. 
see Appendix \ref{sec:app1} for a description of the analytical form of these profiles).
At each iteration, 
the model is numerically convolved with a Moffat filter, to account for PSF effects, 
and then matched to the data.
The minimization routine is based on the Levenberg-Marquardt algorithm.
During the fit, we deliberately did not make use of any weighting scheme on the data points that constitute the 
surface brightness profile.

\subsection{Smoothing technique}
\label{sec:smooth}
Some nearby galaxies in our sample have very large apparent sizes 
and for them we obtained surface brightness profiles more extended than $8~\rm arcmin$.
This means that their outermost (significant) isophote 
corresponds to a projected galactic radius $R$ 
of more than $240$ times the $FWHM$ of the instrumental PSF.
Such level of spatial resolution is unnecessary for the purposes of our analysis 
and, especially in the case of a clumpy star forming galaxy, 
it results in a ``noisy'' surface brightness profile.
Moreover, in the case of a linearly sampled light profile,
it significantly prolongs the computational time of the fitting routine 
(because, at each iteration, the PSF convolution is performed numerically on a large array).
To overcome this problem, we introduced a method to which we refer as the ``smoothing technique''.
This method was applied to the galaxies M31, M81, NGC 4945 and NGC 5128.
For each of these galaxies, we took the image mosaic and we convolved it with a Gaussian filter 
whose $FWHM$ was larger than the $FWHM$ of the instrumental PSF.
We then ran {\tt ellipse} on the convolved image and extracted 
``linearly sampled'' and ``logarithmically sampled'' surface brightness profiles. 
For the ``linearly sampled'' profiles, we set the radial step between contiguous isophotes to be 
comparable to the $FWHM$ of the smoothing Gaussian filter.
Doing so, we reduced the number of fitted isophotes and we also produced smoother surface brightness profiles.
Before the software fit a smoothed light profile, the model to be fit was convolved twice:
the first time to account for PSF effects, 
and the second time to account for the artificial Gaussian smoothing applied to the image mosaic.

\subsection{Identifying and modeling sub-components}
\label{sec:cpts}
In this Section we give a general overview of the guidelines that we followed to identify and model 
the sub-components that constitute our galaxies. 
However, given the level of accuracy and detail to which each galaxy decomposition has been performed in our analysis,
it is hard to encompass all aspects of this matter in a few paragraphs.
The modeling of each galaxy represented a particular and original problem, 
and we remand the reader to Section \ref{sec:indgal},
where we provide individual descriptions of the galaxies that we analyzed. \\
As stressed in Section \ref{sec:int}, our investigation is primarily focused on the spheroidal components 
of galaxies.
%This is naively identified with the whole galaxy in case of an elliptical galaxy,
%and with the bulge of a lenticular or spiral galaxy.
We assumed \emph{a priori} that each of our galaxies has a unique spheroidal component, 
that we modeled with a S\'ersic profile. \\
Disks were normally fit with the exponential model, 
although in the case of highly inclined or edge-on systems we preferred using $n < 1$ S\'ersic functions. 
\citet{pastrav2013a,pastrav2013b} showed that, due to projection effects, their simulated images of inclined galaxy disks 
are better fit by S\'ersic functions with $n < 1$ than pure exponential profiles.
The inclined embedded disks of some ``elliptical'' galaxies were described with Ferrer functions, rather than $n < 1$ S\'ersic functions. 
This choice is motivated by the fact that a S\'ersic + Ferrer model is less degenerate than a S\'ersic + S\'ersic model, 
since the S\'ersic profile can assume any positive ($n > 1$) or negative ($n < 1$) curvature, 
whereas the Ferrer profile can only have a negative curvature as required for an inclined disk. \\
The presence of large-scale disks, such as those of lenticular and spiral galaxies, 
were known \emph{a priori} from the galaxy morphological classification (as listed on NED), 
although some of them were re-classified by us as having intermediate-scale embedded disks.
These were identified in a number of different ways. 
If highly inclined, they can obviously be spotted from the galaxy image or the unsharp mask.
Local maxima in the ellipticity and fourth harmonic profiles 
can provide footprints of less obvious inner disks.
In particular, the ellipticity profile helps distinguish embedded disks from large-scale ones.
Galaxy disks typically have fixed ellipticity, reflecting their inclination to our line of sight.
On the other hand, spheroids can have their ellipticities varying with radius,
but they are usually rounder than inclined disks, thus their average ellipticities are lower than those of inclined disks.
If the ellipticity profile of a galaxy increases with radius, 
this can be ascribed to an inclined disk that becomes progressively more important over the spheroid,
whereas a radial decrease of ellipticity signifies the opposite case.
Therefore, in a situation where a disk is identified from the galaxy image, 
but its extent (large- or intermediate-scale) is ambiguous,
the shape of the ellipticity profile can be decisive. 
Another way to establish the presence of an embedded disk is to look at the velocity map of a galaxy,
following the approach of \citet{arnold2014}. 
A local angular momentum decrease with increasing radius is indicative of an intermediate-scale disk 
that fades toward larger radii. \\
%As an illustration, we consider the galaxy NGC 3377.
%For this galaxy, \citet{arnold2014} reported a strong decline in local specific angular momentum beyond $\sim 35''$, 
%with a disk-to-bulge transition occurring at about $\sim 50''$.
%They did attempt a three-component decomposition (accounting for a bulge, an embedded disk and a central cusp), 
%but unfortunately they fit the spheroidal component with a limited two-parameter $R^{1/4}$ profile {\bf TBD: cite de vaucouleurs} 
%rather than a three-parameter $R^{1/n}$ S\'ersic function.
%Here we used a surface brightness profile twice as extended as that used by \citet{arnold2014} 
%and we identified a nuclear disk and an intermediate-scale disk, both embedded in a larger spheroidal component (see Figure \ref{fig:n3377ex}).
%The presence of an intermediate-scale disk is evident from the unsharp mask of the image mosaic 
%and also from the ellipticity profile, which increases with increasing radius until $R \sim 45''$ (along the major-axis) 
%and dramatically drops at larger radii.
%The nuclear disk can be spotted from the peak at $R \sim 5''$ in the $B4$ profile.
%In our galaxy model for NGC 3377, 
%the difference between the S\'ersic profile of the bulge and the exponential profile of the intermediate-scale disk is minimum in correspondence 
%of the maximum ellipticity, 
%thus our model correctly reproduces the isophotal properties of the galaxy. 
%The residuals of the fit, especially along the major-axis, display some minor structures which indicate that 
%the bulge S\'ersic index might have been slightly overestimated.
%This could be due to the fact that the embedded disk would be better described by a truncated exponential function.
%Alternatively, a contribution from a stellar halo (not included in our model) at large radii could be the reason of a 
%too large S\'ersic index for the bulge. 
%Our equivalent-axis decomposition results in a bulge component that accounts for 94\% of the total luminosity, 
%an intermediate-scale embedded disk contributing to 5\% 
%and a nuclear disk with 1\% of the total light. \\
%A similar situation occurs for the galaxy M59 (see Figure \ref{fig:m59ex}), 
%for which our equivalent-axis model returns a spheroidal component and an embedded disk 
%contributing to 92\% and 8\% of the total light respectively. \\
%The edge-on embedded disk of the S\'ersic galaxy NGC 1332 (see Figure \ref{fig:n1332ex}) was hard to model with an exponential function, 
%leading us to prefer a low-$n$ S\'ersic profile. 
%Our equivalent-axis decomposition returns a spheroidal component accounting for 95\% of the total light,
%and an embedded disk accounting for the remaining 5\%.
%In passing, we note that the bulge-disk decomposition performed by \citet{rusli2011} on NGC 1332 
%is significantly different from our best-fit model.
%\citet{rusli2011} did not identify an intermediate-scale embedded disk, 
%but instead proposed a model featuring a S\'ersic bulge ($n \sim 2.3$, $R_{\rm e} \sim 8''$), accounting for 43\% of the total light, 
%and a large-scale exponential disk. 
%This result led them to the conclusion that NGC 1332 is a disk dominated lenticular galaxy 
%and is displaced from the $M_{\rm BH} - L_{\rm K,sph}$ (black hole mass vs. $K$-band spheroid luminosity) relation of \citet{marconihunt2003} 
%by an order of magnitude along the $M_{\rm BH}$ direction.
%Assuming instead that, as suggested by our model, 
%the galaxy spheroid is a factor of two more luminous than what claimed by \citet{rusli2011}, 
%NGC 1332 would no longer be an outlier in the $M_{\rm BH} - L_{\rm K,sph}$ diagram (see Figure \ref{fig:n1332pl}), 
%perfectly fitting within the $1\sigma$ scatter of the $M_{\rm BH} - L_{\rm K,sph}$ relation for S\'ersic bulges of \citet{grahamscott2013}. \\
Bars are usually recognizable from galaxy images and unsharp masks, 
although local maxima/minima or abrupt changes in the radial profiles of the isophotal parameters 
can provide additional evidence for less obvious bars. 
As noted, we fit bars with Ferrer functions or low-$n$ S\'ersic profiles ($n \sim 0.2$). \\
The presence of a nuclear component -- either resolved or unresolved -- was generally expected in 
(but not restricted to) galaxies that host an optical AGN and circumnuclear dust.
Nuclear stellar disks and nuclear star clusters fall into the category of nuclear components too,
but their identification can be more subtle than for AGNs.
Nuclear clusters have typical sizes of a few parsecs, 
therefore for the majority of our galaxies they are unresolved in \emph{Spitzer}/IRAC $3.6~\rm \mu m$ observations.
If an identification from high-resolution observations was available from the literature, we relied on that, 
otherwise we concluded that a galaxy was nucleated from an excess of nuclear light in the residuals of the 
fit\footnote{This conclusion was drawn after going through the following steps. 
First we identified all the sub-components of a galaxy 
(assuming that the galaxy was not nucleated), built a model accordingly and fit it to the data.
If the residuals of the fit showed a nuclear light excess, 
we repeated the fit by excluding the data points within the nuclear region.
Only after checking that the outcome of the last fit was consistent with the previous one, 
we inferred the presence of a stellar nuclear component, and accounted for it in the final galaxy model.}.
Unresolved nuclear components were fit with PSF-like profiles,
whereas resolved nuclear components were modeled with (PSF-convolved) narrow Gaussian functions. \\
Rings were identified from galaxy images and unsharp masks, and modeled with symmetric Gaussian ring profiles. \\
As an illustration, we consider the galaxy NGC 2974, 
a spiral galaxy which has been misclassified as an elliptical galaxy in the RC3 catalog \citep{RC3}. 
This galaxy hosts a Seyfert AGN \citep{veroncettyveron2006} and filamentary dust in its center \citep{tran2001}.
NGC 2974 is classified as fast rotator by the ATLAS$^{\rm 3D}$ survey, 
and indeed the velocity map obtained by the SLUGGS survey shows that the galaxy kinematics is rotation-dominated 
well beyond three effective radii ($R>150''$), 
as expected from a large-scale disk. 
%\cite{sani2011} modeled this galaxy with a S\'ersic-bulge + Gaussian-AGN 
%($R_{\rm e} = 27''.2$, $n = 3.0$). 
From an inspection of the unsharp mask, we identified a ring at $R \sim 50''$, 
which might be a residual of two tightly wound spiral arms,
and an elongated bar-like component within $R \lesssim 30''$,
that produces a peak in the ellipticity and position angle profiles at $R \sim 20''$.
Our 1D galaxy decomposition for NGC 2974 (see Figure \ref{fig:n2974ex}) consists of a S\'ersic bulge, an exponential large-scale disk,
a Ferrer bar, a Gaussian nuclear component (AGN) and a Gaussian ring.
Although the ring is extremely faint, it is important to account for it in the galaxy decomposition. 
A model without the ring component results in a ``steeper'' exponential profile for the disk 
(i.e.~the exponential model has smaller scale length and more negative central surface brightness) 
and produces bad residual structures within $R \lesssim 40''$.
Our best-fit model returns a bulge major-axis effective radius $R_{\rm e,sph}^{\rm maj} = 8.3\rm~arcsec$, 
a bulge equivalent-axis S\'ersic index $n_{\rm sph}^{\rm eq} = 1.2$ and a $3.6~\rm \mu m$ bulge apparent magnitude $m_{\rm sph} = 8.65~\rm mag$.
\cite{sani2011} modeled NGC 2974 with a S\'ersic bulge and a Gaussian nuclear component (AGN), 
but did not account for the large-scale disk.
From their best-fit 2D model, they obtained a three times larger bulge major-axis effective radius ($R_{\rm e,sph}^{\rm maj} = 27.2\rm~arcsec$), 
a 2.5 times larger bulge S\'ersic index ($n_{\rm sph} = 3$),
and a significantly brighter $3.6~\rm \mu m$ bulge apparent magnitude ($m_{\rm sph} = 7.28~\rm mag$).

%\begin{figure*}[h]
%\begin{center}
%\includegraphics[width=\fitfigurewidth]{images/n3377_1Dfit.eps}
%\caption{Best-fit model and isophotal parameters for the galaxy NGC 3377. 
%The left panels refer to the major-axis $R_{\rm maj}$, 
%whereas the right panels refer to the equivalent-axis $R_{\rm eq}$.
%The top panels display the galaxy surface brightness radial profiles obtained with a linear sampling. 
%The black points are the observed data.  
%The color lines represent the individual (PSF-convolved) model components:
%red=S\'ersic (bulge), blue=exponential (disk). 
%The best-fit parameters for the S\'ersic profile are included.
%The total (PSF-convolved) model is shown with a black dashed line, 
%but it is hard to distinguish because it almost perfectly matches the data.
%The horizontal grey dashed line indicates the $3~RMS_{\rm sky}$ level.
%$\Delta$ denotes the rms scatter of the fit in units of $\rm mag~arcsec^{-2}$.
%$\Delta \mu$ are the residuals from the fit.
%The six panels below show the ellipticity ($\epsilon$), position angle ($PA$) and $B4$ radial profiles.
%}
%\label{fig:n3377ex}
%\end{center}
%\end{figure*}

\begin{figure*}[h]
\begin{center}
\includegraphics[width=\fitfigurewidth]{images/n2974_1Dfit.eps}
\caption{Best-fit model and isophotal parameters for the galaxy NGC 2974. 
The left panels refer to the major-axis $R_{\rm maj}$, 
whereas the right panels refer to the equivalent-axis $R_{\rm eq}$.
The top panels display the galaxy surface brightness radial profiles obtained with a linear sampling. 
The black points are the observed data.  
The color lines represent the individual (PSF-convolved) model components:
red=S\'ersic (bulge), blue=exponential (disk), green=Gaussian (AGN), cyan=Ferrer (bar), gray=Gaussian ring (ring). 
The best-fit parameters for the S\'ersic profile are included.
The total (PSF-convolved) model is shown with a black dashed line, 
but it is hard to distinguish because it almost perfectly matches the data.
The horizontal grey dashed line indicates the $3~RMS_{\rm sky}$ level.
$\Delta$ denotes the rms scatter of the fit in units of $\rm mag~arcsec^{-2}$.
$\Delta \mu$ are the residuals from the fit.
The six panels below show the ellipticity ($\epsilon$), position angle ($PA$) and fourth harmonic ($B4$) radial profiles.}
\label{fig:n2974ex}
\end{center}
\end{figure*}

%\begin{figure*}[h]
%\begin{center}
%\includegraphics[width=\fitfigurewidth]{images/m59_1Dfit.eps}
%\caption{Best-fit model and isophotal parameters for the galaxy M59. 
%The figure caption is the same as Figure \ref{fig:n3377ex}. 
%Empty data points represent data excluded from the fit.}
%\label{fig:m59ex}
%\end{center}
%\end{figure*}
%
%\begin{figure*}[h]
%\begin{center}
%\includegraphics[width=\fitfigurewidth]{images/n1332_1Dfit.eps}
%\caption{Best-fit model and isophotal parameters for the galaxy NGC 1332.
%The figure caption is the same as Figure \ref{fig:n3377ex}. 
%Empty data points represent data excluded from the fit.}
%\label{fig:n1332ex}
%\end{center}
%\end{figure*}
%
%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=\columnwidth]{images/mbhmagk_n1332.pdf}
%\caption{The location of the S\'ersic galaxy NGC 1332 in the black hole mass versus $K$-band spheroid absolute magnitude diagram,
%according to our galaxy decomposition. 
%The red and blue solid lines show the scaling relations for core-S\'ersic and S\'ersic galaxies, respectively, 
%reported by \citet{grahamscott2013}, 
%while the shaded areas mark the $1\sigma$ errors.}
%\label{fig:n1332pl}
%\end{center}
%\end{figure}

\subsection{2D fits}
\label{sec:2d}
Two-dimensional decompositions were carried out using the software \textsc{Imfit} \citep{imfit}.
For each galaxy, we built a 2D model that was consistent with the corresponding 1D model in terms of number and type of components. 
Because the Ferrer profile is not made available in \textsc{Imfit},
bars were fit with a 2D Gaussian function. \\
The 2D decomposition of NGC 2974 is presented in Figure \ref{fig:n29742d}. 
The galaxy was modeled with a S\'ersic bulge, an exponential disk and a Gaussian bar.
The nuclear component was masked and the ring was not 
modeled\footnote{We built the 2D model first including and then omitting a gaussian ring component, 
but both models converged to the same solution, i.e.~the fit ``ignored'' the presence of the faint ring. 
This did not happen in the 1D decomposition because of the different weighting scheme used by the fitting routines.}. 
Our 2D best-fit model returns a bulge major-axis effective radius $R_{\rm e,sph}^{\rm maj} = 10.5~\rm arcsec$, 
a bulge S\'ersic index $n_{\rm sph} = 1.3$ and a $3.6~\rm \mu m$ bulge apparent magnitude $m_{\rm sph} = 8.39~\rm mag$, 
in fairly good agreement with our 1D decomposition.

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{images/n2974_imgmsk.pdf}
\includegraphics[width=\columnwidth]{images/n2974_model.pdf}
\includegraphics[width=\columnwidth]{images/n2974_resid.pdf}
\caption{The image of the galaxy NGC 2974 with its mask (top panel), the best-fit 2D model (middle panel), 
and the residual image after the subtraction of the 2D model (bottom panel).
The ring has not been modeled, and therefore appears in the residual image.}
\label{fig:n29742d}
\end{center}
\end{figure}

\section{Results}
\label{sec:res}
For each galaxy, after we identified its various components and built a model accordingly, 
we simultaneously performed a set of four 1D fits.
All four fits use a Moffat-convolved model.
Two fits use the major-axis surface brightness profile and the remaining two use the equivalent-axis surface brightness profile. 
For each of these pairs, we use a logarithmically sampled surface brightness profile 
and a linearly sampled surface brightness profile. 
Because our fitting routine does not employ any error weighting scheme on the data points that constitute the surface brightness profile, 
a fit to a logarithmically sampled profile puts more weight on the inner region of the galaxy 
and poorly constrains the outskirts.
On the other hand, a fit to a linearly sampled surface brightness profile equally treats inner and outer regions. 
We found that the fits are, in general, more sensitive to the choice of the initial parameters 
when using logarithmically sampled profiles than linearly sampled profiles. 
In addition, a visual examination of the residuals demonstrated that the goodness of the fit within one galaxy effective radius 
is superior when tighter constraints are put on the galaxy outskirts.
In other words, the better quality of the residuals led us to prefer fits that use linearly sampled surface brightness profiles. \\
Among the initial sample of 75, we did not attempt to model 3 galaxies: M32, NGC 4486A and the Circinus Galaxy. 
The first two... {\bf AL: can you write a short sentence about these 2 galaxies, that have uncertain morphology,
because they have been stripped, blah blah, maybe put one reference? please and thank you.}
The Circinus Galaxy lies at only 4 degrees from the Galactic Plane, 
therefore its image mosaic is contaminated by a large number of foreground stars.
Of the remaining 72 galaxies, 
we obtained satisfactory 1D decompositions for 66, 
whereas the models of 6 galaxies were judged not reliable and were thus excluded. 
We also performed 2D decompositions for 31 galaxies. \\
A galaxy-by-galaxy comparison between our best-fit models and those from the previous literature 
helped identify the optimal decompositions and past problems. 
We compared our best-fit models with those of \cite{grahamdriver2007}, \cite{sani2011}, \cite{beifiori2012}, 
\cite{vika2012}, \cite{lasker2014data}. 
These five studies all investigated black hole mass scaling relations.
We also considered the best-fit models of \cite{laurikainen2010} and \cite{rusli2013} because,   
although they did not specifically deal with black hole - galaxy scaling relations, 
their galaxy samples significantly overlap with ours. \\
%In Section \ref{sec:indgal} we discuss individual cases, explaining the discrepancies between different decompositions. \\
Table \ref{tab:fitres} lists the results from the 1D and 2D fits.

\input{table_fitresults.tex}

\subsection{1D versus 2D decompositions}
%In Figures \ref{fig:re1d2d}, \ref{fig:n1d2d} and \ref{fig:mag1d2d}, we compare the bulge best-fit parameters obtained 
%from the 1D decompositions with those obtained from the 2D decompositions.
How do 1D and 2D decompositions compare with each other? 
Although an exhaustive analysis is beyond the scope of the present work, 
we believe that some of the readers might be interested in our practical knowledge, 
having dealt with 1D and 2D techniques of galaxy modeling at the same time. \\
We summarize our experience in the following points:
\begin{itemize}
\item A visual inspection of galaxy images and their unsharp masks is often not sufficient to accurately identify all galaxy components. 
Embedded disks or weak bars can easily be missed, and (inclined) intermediate-scale disks can be confused with large-scale ones.
In this regards, the 1D isophotal analysis is extremely helpful. 
Local minima/maxima or abrupt changes in the ellipticity, position angle and fourth harmonic profiles contain precious information 
about the galaxy constituents.
\item The ellipticity and position angle of bulges can vary with radius.
The analytic functions used by 2D decomposition codes to fit galaxy components have fixed ellipticity and position angle, 
thus they cannot account for their radial gradients. 
This problem is overcome with 1D decomposition techniques.
\item The interpretation of the residuals is often crucial to identify the optimal decomposition for a galaxy. 
In our experience, we found that the reading of 1D residuals was easier than that of 2D residuals.
\end{itemize}
Although we attempted 2D modeling for all galaxies in our sample, 
nearly half of the 2D decompositions were not successful or did not converge.
For the 31 galaxies that had successful 1D and 2D decompositions, 
we compare their 1D and 2D best-fit parameters in Figures \ref{fig:re1d2d}, \ref{fig:n1d2d} and \ref{fig:mag1d2d}.
%The software \textsc{Imfit} returns major-axis measurements of effective radii, 
%which are plotted in Figure \ref{fig:re1d2d} versus the corresponding 1D major-axis measurements.
%2D measurements of S\'ersic indices can be approximated to 1D equivalent-axis measurements, 
%and are therefore compared to them in Figure \ref{fig:n1d2d}.
%Figure \ref{fig:mag1d2d} shows 2D versus 1D measurements of bulge magnitudes. 
The agreement between 1D and 2D effective radii and magnitudes is remarkable, 
whereas a larger amount of scatter in the S\'ersic indices plot might be caused by the fact that 
2D measurements do not exactly correspond to 1D equivalent-axis measurements. 
No systematic effects are observed in any of these three plots, 
which indicates that 1D and 2D techniques of galaxy modeling -- when performed on the same galaxy -- 
give consistent results. \\
In conclusion, since we found that the best-fit parameters do not depend on the decomposition method (1D or 2D) used, 
but we obtained more successful 1D decompositions than 2D, 
we will base our analysis only on the results from 1D fits.

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{images/comparison_r_e.eps} 
\caption{2D versus 1D major-axis measurements of effective radii. 
The dashed line displays the 1:1 relation.}
\label{fig:re1d2d}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{images/comparison_n.eps} 
\caption{2D measurements of S\'ersic indices can be approximated to 1D equivalent-axis measurements, 
and are therefore compared to them in this plot. The dashed line displays the 1:1 relation.}
\label{fig:n1d2d}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{images/comparison_mag.eps} 
\caption{2D versus 1D measurements of $3.6~\rm \mu m$ bulge magnitudes. The dashed line displays the 1:1 relation.}
\label{fig:mag1d2d}
\end{center}
\end{figure}

\subsection{Parameter uncertainty}
\label{sec:err}
Estimating the uncertainties associated with the best-fit parameters of our 1D galaxy decompositions is not straightforward.
Monte Carlo simulations could be used for this purpose, 
but they would take into account only random errors and not unknown systematic ones.
Systematic errors include incorrect sky subtraction, 
inaccurate masking of contaminating sources, 
imprecise description of the PSF, 
erroneous choice of model components 
(for example, when failing to identify a galaxy sub-component and thus omitting it in the model, 
or when describing a galaxy sub-component with an inadequate function),
the radial extent of the surface brightness profile and its sampling.
%(if no weighting scheme is applied to the data points during the minimization routine, 
%then, for the same galaxy, the fit of a logarithmically sampled light profile will naturally put more weight on the inner part of the profile 
%than in the case of linear sampling).
When performing multi-component decomposition of high signal-to-noise images of nearby -- therefore well resolved -- galaxies, 
errors are dominated by systematics rather than Poisson noise.
For this reason, we decided to estimate the uncertainties of the bulge best-fit parameters 
with a method that took into account systematic errors.

\subsubsection{Goodness of the bulge modeling}
To do this, we first had to evaluate the goodness of the modeling of the bulge component.
For each of our fits, we calculated the associated $RMS$ scatter\footnote{The $RMS$ scatter of a fit was computed as:
\begin{equation}
\Delta = \sqrt{\frac{\sum_{\rm i=0}^{N} (\mu_{\rm i}^{\rm observed} - \mu_{\rm i}^{\rm model})^2}{N_{\rm DOF}} } ,
\end{equation} 
where %$N$ is the number of data points used in the fit, 
$N_{\rm DOF}$ is the number of degrees-of-freedom, 
$\mu^{\rm observed}$ is the observed surface brightness 
and $\mu^{\rm model}$ is the model surface brightness.}.
Although useful to evaluate the overall quality of a galaxy decomposition, 
the $RMS$ scatter alone cannot be used to assess the goodness of the fit for the spheroidal component only, 
unless the galaxy is a pure spheroid and has consequently been modeled with a single S\'ersic profile.
To illustrate this point with an example, 
one can imagine a situation in which a galaxy is thought to be made of a small bulge and a much more extended disk.
This galaxy is decomposed with a S\'ersic + exponential model.
The exponential function provides an excellent description of the light profile of the disk, 
whereas the S\'ersic function does not do the same for the bulge.
The residuals of the fit will then be flat and close to zero at large radii, 
where the emission of the disk dominates over that of the bulge,
while they will display significant structures at small radii, 
in correspondence of the poorly fit spheroidal component.
In case of linear sampling, 
because the part of the surface brightness profile pertaining to the disk contains more data points 
than the part pertaining to the bulge, 
the global $RMS$ scatter will be relatively small, 
but it obviously will not reflect the accuracy of the fit to the spheroidal component only. \\
A simple but powerful way to get a feeling of how precisely the global model describes the spheroidal component 
is to look at the major- and equivalent-axis fits of each galaxy 
and visually inspect the structures of the residuals within $\sim 1 - 2$ bulge effective radii. 
We did this using a grade from 1 to 3, assigned according to the following criteria.
\begin{itemize}
\item [1)] A grade of 1 was given to the best fits, i.e.~fits that do not exhibit any of the problems listed below.
\item [2)] A grade of 2 was used in the following cases: 
the residuals in correspondence of the bulge component are not randomly distributed around zero, 
being symptomatic of a S\'ersic model having a curvature (regulated by the S\'ersic index $n$) 
that does not match the real ``shape'' of the bulge; 
when we identified, but were not able to accurately model a galaxy sub-component;
in the case of significant inconsistencies between the model and the observed galaxy properties 
(e.g.~an embedded disk modeled with an exponential function, 
whose scale length does not match the size of the disk as expected from the ellipticity profile or the velocity map);
when the S\'ersic profile used to describe the bulge component has a size -- as measured by the effective radius -- 
comparable to a few times the $FWHM$ of the PSF.
\item [3)] A grade of 3 was assigned to the poorer and more anomalous fits, 
or those affected by an obvious degeneracy between the bulge S\'ersic profile and the remaining model components 
(e.g.~when the bulge S\'ersic index varies for more than 50\% {\bf TBD: how often what?} among the four different realizations of the fit, 
or when the output of the fit strongly depends on the choice of the initial parameters).
\end{itemize}
As a result, we classified 27 galaxies (38\% of the 72 galaxies for which we attempted a 1D decomposition) with grade 1, 
29 galaxies (40\%) with grade 2 and 10 galaxies (14\%) with grade 3.
%The absolute values of the uncertainties of the S\'ersic parameters will be estimated in a forthcoming paper, 
%according to the grades assigned here to the fits.

\subsubsection{Uncertainties on $n_{\rm sph}$}
%\label{sec:sigman}
In Figure \ref{fig:compn}, for 58 galaxies, 
we compare the measurements of the bulge S\'ersic index obtained by different authors 
with those obtained by us. 
For each galaxy, we computed the average value $\langle \log(n_{\rm sph}) \rangle$ of all the available measurements, 
and we plot it against the scatter of the individual measurements around $\langle \log(n_{\rm sph}) \rangle$.
The measurements are deliberately heterogeneous, 
in the sense that they were obtained from 1D or 2D decompositions of data in different wavelengths, 
and they refer either to the major-axis or to the equivalent-axis.
In Figure \ref{fig:compn}, the black histogram shows the distribution of the scatter around $\langle \log(n_{\rm sph}) \rangle$ 
only for our measurements. 
38\% of this distribution lie within $-\sigma_1^- = -0.08~\rm dex$ and $+\sigma_1^+ = +0.06~\rm dex$, 
78\% lie within $-\sigma_2^- = -0.21~\rm dex$ and $+\sigma_2^+ = +0.17~\rm dex$,
and 92\% lie within $-\sigma_3^- = -0.25~\rm dex$ and $+\sigma_3^+ = +0.25~\rm dex$. 
We elect to use $\pm \sigma_1^\pm$, $\pm \sigma_2^\pm$ and $\pm \sigma_3^\pm$ as $1\sigma$ uncertainties 
for our measurements of $\log(n_{\rm sph})$ obtained from ``grade 1'', ``grade 2'' and ``grade 3'' fits, respectively. \\

\subsubsection{Uncertainties on $R_{\rm e,sph}$}
The uncertainties on the bulge effective radii were computed with the same method used for the S\'ersic indices. 
However, in Figure \ref{fig:compre} we include only major-axis measurements of $R_{\rm e,sph}$. 
The associated $1\sigma$ uncertainties for our measurements of $\log(R_{\rm e,sph})$ are 
$-\sigma_1^- = -0.11~\rm dex$ and $+\sigma_1^+ = +0.07~\rm dex$, 
$-\sigma_2^- = -0.30~\rm dex$ and $+\sigma_2^+ = +0.32~\rm dex$,
and $-\sigma_3^- = -0.39~\rm dex$ and $+\sigma_3^+ = +0.42~\rm dex$.

\subsubsection{Uncertainties on $m_{\rm sph}$}
To estimate the uncertainties on the bulge magnitudes, 
we compared (see Figure \ref{fig:compmag}) only measurements coming from $K$-band or $3.6~\rm \mu m$ observations.
$3.6~\rm \mu m$ magnitudes were converted into $K$-band magnitudes by an additive factor of $+0.27\rm~mag$, 
which was estimated using the stellar population models of \cite{worthey1994},
assuming a 13 Gyr old single-burst stellar population with solar metallicity. 
The associated $1\sigma$ uncertainties for our measurements of $m_{\rm sph}$ are 
$-\sigma_1^- = -0.11~\rm mag$ and $+\sigma_1^+ = +0.18~\rm mag$, 
$-\sigma_2^- = -0.58~\rm mag$ and $+\sigma_2^+ = +0.66~\rm mag$,
and $-\sigma_3^- = -0.66~\rm mag$ and $+\sigma_3^+ = +0.88~\rm mag$. \\
{\bf NOTE: When converting from K-band to 3.6, 
Galactic absorption was not taken into account!!!!! But I have the feeling that it's not significant...}

\subsubsection{Uncertainties on $\mu_{\rm e,sph}$}
As for the bulge magnitudes, we estimated the uncertainties on the bulge effective surface brightnesses $\mu_{\rm e,sph}$ 
by comparing only $K$-band or $3.6~\rm \mu m$ measurements (see Figure \ref{fig:compmue}), 
and accounting for the corrective factor. 
Because not explicitly reported by their authors, literature effective surface brightnesses were calculated by us as:
\begin{multline}
\mu_{\rm e,sph} = m_{\rm sph} + 5\log \Bigl(R_{\rm e,sph}^{\rm maj}\sqrt{(b/a)_{\rm sph}} \Bigr) + \\
+ 2.5\log \Bigl[2\pi n_{\rm sph} e^{b_n} b_n^{-2n_{\rm sph}} \Gamma(2n_{\rm sph})\Bigr], 
\end{multline}
where $b_n$ and $\Gamma(2n_{\rm sph})$ are defined in Section \ref{sec:app1}, 
and $(b/a)_{\rm sph}$ is the bulge axis ratio.
While \cite{laurikainen2010} and \cite{sani2011} reported their estimates of $(b/a)_{\rm sph}$, 
\cite{vika2012} and \cite{lasker2014data} did not. 
For the last two studies, we used the values of $(b/a)_{\rm sph}$ reported by \cite{sani2011}.
The associated $1\sigma$ uncertainties for our measurements of $\mu_{\rm e,sph}$ are 
$-\sigma_1^- = -0.33~\rm mag$ and $+\sigma_1^+ = +0.57~\rm mag$, 
$-\sigma_2^- = -0.84~\rm mag$ and $+\sigma_2^+ = +1.59~\rm mag$,
and $-\sigma_3^- = -1.06~\rm mag$ and $+\sigma_3^+ = +1.74~\rm mag$.

\subsubsection{Uncertainties on $\mu_{\rm 0,sph}$}
From the values of $\mu_{\rm e,sph}$, we derived the central surface brightnesses $\mu_{\rm 0,sph}$ as
\begin{equation}
\mu_{\rm 0,sph} = \mu_{\rm e,sph} - \frac{2.5 b_n}{\ln(10)}.
\end{equation}
The associated $1\sigma$ uncertainties for our measurements of $\mu_{\rm 0,sph}$ are 
$-\sigma_1^- = -0.20~\rm mag$ and $+\sigma_1^+ = +0.64~\rm mag$, 
$-\sigma_2^- = -0.70~\rm mag$ and $+\sigma_2^+ = +1.24~\rm mag$,
and $-\sigma_3^- = -1.05~\rm mag$ and $+\sigma_3^+ = +1.57~\rm mag$.

\begin{figure}%[h]
\begin{center}
\includegraphics[width=1.1\columnwidth]{images/comparison_all_n.eps} 
\caption{The bottom panel shows 58 galaxies for which at least one measurement of the bulge S\'ersic index $n_{\rm sph}$ is available 
from the literature, in addition to that measured by us.
In ordinate, for each galaxy, we plot the average value $\langle \log(n_{\rm sph}) \rangle$ of (the logarithm of) the individual measurements. 
In abscissa we plot the difference between (the logarithm of) the individual measurements of a galaxy 
and the average value for that same galaxy.
Each data point corresponds to an individual measurement, 
which comes from one of the following studies. 
Red points are from \cite{grahamdriver2007}, 
blue points are from \cite{laurikainen2010}, green points are from \cite{sani2011}, 
yellow points are from \cite{vika2012}, gray points are from \cite{beifiori2012},
white points are from \cite{rusli2013}, orange points are from \cite{lasker2014data},
and black points are measurements obtained from the 1D fits presented in this work 
(using linearly sampled surface brightness profiles, along the major-axis).
In the top panel, the white histogram shows the distribution of $\log(n_{\rm sph}) - \langle \log(n_{\rm sph}) \rangle$ 
for all measurements, 
whereas the black histogram refers only to measurements obtained by us.
The (asymmetric) error bars $\sigma_1$, $\sigma_2$ and $\sigma_3$ enclose 38\%, 78\% and 92\% 
of the black histogram, respectively.}
\label{fig:compn}
\end{center}
\end{figure}

\begin{figure}%[h]
\begin{center}
\includegraphics[width=1.1\columnwidth]{images/comparison_all_r_e.eps} 
\caption{The bottom panel shows 52 galaxies for which at least one measurement 
of the major-axis effective radius $R_{\rm e}^{\rm maj}$ is available 
from the literature, in addition to that measured by us.
In ordinate, for each galaxy, we plot the average value $\langle \log(R_{\rm e}^{\rm maj}) \rangle$ 
of (the logarithm of) the individual measurements. 
In abscissa we plot the difference between (the logarithm of) the individual measurements of a galaxy 
and the average value for that same galaxy.
Each data point corresponds to an individual measurement, 
which comes from one of the following studies. 
Red points are from \cite{grahamdriver2007}, 
blue points are from \cite{laurikainen2010}, green points are from \cite{sani2011}, 
yellow points are from \cite{vika2012}, gray points are from \cite{beifiori2012},
orange points are from \cite{lasker2014data},
and black points are measurements obtained from the 1D fits presented in this work 
(using linearly sampled surface brightness profiles, along the major-axis).
In the top panel, the white histogram shows the distribution of $\log(R_{\rm e}^{\rm maj}) - \langle \log(R_{\rm e}^{\rm maj}) \rangle$ 
for all measurements, 
whereas the black histogram refers only to measurements obtained by us.
The (asymmetric) error bars $\sigma_1$, $\sigma_2$ and $\sigma_3$ enclose 38\%, 78\% and 92\% 
of the black histogram, respectively.}
\label{fig:compre}
\end{center}
\end{figure}

\begin{figure}%[h]
\begin{center}
\includegraphics[width=1.1\columnwidth]{images/comparison_all_mag_sph.eps} 
\caption{The bottom panel shows 51 galaxies for which at least one measurement 
of the bulge apparent magnitude $m_{\rm sph}$ -- either in $K$-band or $3.6~\rm \mu m$ -- is available 
from the literature, in addition to that measured by us. 
$3.6~\rm \mu m$ magnitudes were converted to $K$-band magnitudes.
In ordinate, for each galaxy, we plot the average value $\langle m_{\rm sph} \rangle$ 
of the individual measurements. 
In abscissa we plot the difference between the individual measurements of a galaxy 
and the average value for that same galaxy.
Each data point corresponds to an individual measurement, 
which comes from one of the following studies. 
Blue points are from \cite{laurikainen2010}, green points are from \cite{sani2011}, 
yellow points are from \cite{vika2012}, 
orange points are from \cite{lasker2014data},
and black points are measurements obtained from the 1D fits presented in this work 
(using linearly sampled surface brightness profiles, along the equivalent-axis).
In the top panel, the white histogram shows the distribution of $m_{\rm sph} - \langle m_{\rm sph} \rangle$ 
for all measurements, 
whereas the black histogram refers only to measurements obtained by us.
The (asymmetric) error bars $\sigma_1$, $\sigma_2$ and $\sigma_3$ enclose 38\%, 78\% and 92\% 
of the black histogram, respectively.}
\label{fig:compmag}
\end{center}
\end{figure}

\begin{figure}%[h]
\begin{center}
\includegraphics[width=1.1\columnwidth]{images/comparison_all_mu_e.eps} 
\caption{The bottom panel shows 49 galaxies for which at least one measurement 
of the bulge effective surface brightness $\mu_{\rm e,sph}$ -- either in $K$-band or $3.6~\rm \mu m$ -- is available 
from the literature, in addition to that measured by us. 
$K$-band magnitudes were converted to $3.6~\rm \mu m$ magnitudes.
In ordinate, for each galaxy, we plot the average value $\langle \mu_{\rm e,sph} \rangle$ 
of the individual measurements. 
In abscissa we plot the difference between the individual measurements of a galaxy 
and the average value for that same galaxy.
Each data point corresponds to an individual measurement, 
which comes from one of the following studies. 
Blue points are from \cite{laurikainen2010}, green points are from \cite{sani2011}, 
yellow points are from \cite{vika2012}, 
orange points are from \cite{lasker2014data},
and black points are measurements obtained from the 1D fits presented in this work 
(using linearly sampled surface brightness profiles, along the major-axis).
In the top panel, the white histogram shows the distribution of $\mu_{\rm e,sph} - \langle \mu_{\rm e,sph} \rangle$ 
for all measurements, 
whereas the black histogram refers only to measurements obtained by us.
The (asymmetric) error bars $\sigma_1$, $\sigma_2$ and $\sigma_3$ enclose 38\%, 78\% and 92\% 
of the black histogram, respectively.}
\label{fig:compmue}
\end{center}
\end{figure}

\begin{figure}%[h]
\begin{center}
\includegraphics[width=1.1\columnwidth]{images/comparison_all_mu_0.eps} 
\caption{The bottom panel shows 49 galaxies for which at least one measurement 
of the bulge central surface brightness $\mu_{\rm 0,sph}$ -- either in $K$-band or $3.6~\rm \mu m$ -- is available 
from the literature, in addition to that measured by us. 
$K$-band magnitudes were converted to $3.6~\rm \mu m$ magnitudes.
In ordinate, for each galaxy, we plot the average value $\langle \mu_{\rm 0,sph} \rangle$ 
of the individual measurements. 
In abscissa we plot the difference between the individual measurements of a galaxy 
and the average value for that same galaxy.
Each data point corresponds to an individual measurement, 
which comes from one of the following studies. 
Blue points are from \cite{laurikainen2010}, green points are from \cite{sani2011}, 
yellow points are from \cite{vika2012}, 
orange points are from \cite{lasker2014data},
and black points are measurements obtained from the 1D fits presented in this work 
(using linearly sampled surface brightness profiles, along the major-axis).
In the top panel, the white histogram shows the distribution of $\mu_{\rm 0,sph} - \langle \mu_{\rm 0,sph} \rangle$ 
for all measurements, 
whereas the black histogram refers only to measurements obtained by us.
The (asymmetric) error bars $\sigma_1$, $\sigma_2$ and $\sigma_3$ enclose 38\%, 78\% and 92\% 
of the black histogram, respectively.}
\label{fig:compmu0}
\end{center}
\end{figure}

\section{Conclusions}
\label{sec:concl}
The widespread presence of embedded components -- such as intermediate-scale disks -- in massive early-type galaxies 
makes galaxy decomposition an essential tool to properly investigate the scaling relations between black hole masses 
and host bulge properties. \\
The bulge best-fit parameters are model-dependent. 
Past studies often used different models for the same galaxy, and obtained significantly discrepant results,  
which led them to draw contrasting conclusions about the black hole-bulge correlations.
These inconsistencies motivated our effort to secure and refine the measure of bulge properties in a sample of 75 galaxies 
with a dynamical estimate of the black hole mass. 
Using $3.6~\rm \mu m$ \emph{Spitzer} satellite imagery, 
we performed state-of-the-art galaxy decompositions. 
We compared our best-fit models with those from the past literature, 
and explained discrepancies when present. 
Table \ref{tab:lit} summarizes the main characteristics of the five past studies that, since 2007, attempted galaxy decomposition 
to derive black hole-bulge scaling relations, 
and highlights why our endeavor represents a substantial improvement over the past literature.
Our sample of galaxies is the largest ever used. 
The $3.6~\rm \mu m$ band is an excellent tracer for the stellar mass, superior to optical bands or the $K$-band. 
\cite{lasker2014data} and this work identified and modeled more galaxy components than any other study, 
and therefore obtained more accurate decompositions,
but we are the first that did it with the assistance of kinematical information.

\begin{table*}%[ht] 
\begin{center}  				      
\caption{{\bf Summary of previous investigations of black hole mass scaling relations.} 
GD07 = \cite{grahamdriver2007}, 
S+11 = \cite{sani2011}, V+12 \cite{vika2012}, B+12 = \cite{beifiori2012}, L+14 = \cite{lasker2014data}.}
\begin{tabular}{lllllll}
\hline
              & {\bf GD07}  & {\bf S+11}   & {\bf V+12} & {\bf B+12}   & {\bf L+14} & {\bf This work}	      \\ %%[0.5ex] 
\hline
{\bf Galaxies with successful fit} & 27		     & 57	       & 25      & 19          & 35      & 66               \\
{\bf Wavelength}        	   & $R$-band	     & $3.6~\rm \mu m$ & $K$-band& $i$-band    & $K$-band& $3.6~\rm \mu m$  \\
{\bf Decomposition}     	   & 1D 	     & 2D	       & 2D	 & 2D	       & 2D	 & 1D \& 2D	    \\
{\bf Nuclear components}       	   & masked	     & modeled         & modeled & not treated & modeled & modeled/masked   \\
{\bf Partially depleted cores} 	   & masked	     & masked	       & masked  & not treated & masked  & masked	    \\ 
{\bf Bars}	        	   & excluded	     & modeled         & modeled & excluded    & modeled & modeled	    \\ 
{\bf Other components}  	   & no 	  & no  	       & no	 & no	       & yes	 & yes  	    \\ 
{\bf Kinematics}        	   & no 	     & no	       & no	 & no	       & no	 & yes  	    \\ %%[0.5ex] 
\hline 
\end{tabular}
\label{tab:lit} 
\end{center}	
\end{table*}

Our principal conclusions are summarized in the following points.
\begin{itemize}
\item One-dimensional and two-dimensional techniques of galaxy decomposition return the same results when applied to the same galaxy.
However, in our practical experience, two-dimensional decompositions {\bf HOW DO YOU WRITE THIS IN ENGLISH?? fail twice as much as one-dimensional ones}, 
either because the fit does not converge or because the result is unphysical. 
A strong limitation of two-dimensional codes is their inability to accommodate the radial gradients of ellipticity and position angle 
often observed in galaxy bulges.
\item We found the interpretation of one-dimensional residuals easier than that of two-dimensional residual images. 
A correct interpretation of the residuals is fundamental to determine the optimal model for a galaxy.
\item A one-dimensional isophotal analysis was extremely helpful and sometimes even necessary to accurately identify galaxy components.
\item Given the level of detail to which each galaxy decomposition was performed, 
we believe that our analysis cannot be reproduced by any current automatic routine. 
\item The uncertainties associated with the bulge best-fit parameters are dominated by systematic errors (e.g.~incorrect sky subtraction, 
inaccurate masking of contaminating sources, erroneous choice of model components), and only marginally affected by random errors. 
For this reason, we developed a method to estimate the uncertainties on the best-fit parameters that takes into account systematic errors.
\end{itemize}

We will use the results from our one-dimensional galaxy decompositions to obtain improved black hole mass scaling relations. 
We will also derive and explore some unpublished correlations. 
These results will be presented in a series of forthcoming papers.

%\clearpage

% \onecolumngrid
% \subsection{individual galaxies}
% \label{sec:indgal}
%
% individual galaxies \\
%
% \clearpage\newpage\noindent
% {\bf IC 1459 \\}
%
% IC 1459 is an elliptical galaxy with a fast counterrotating stellar component \citep{franxillingworth1988ic1459,cappellari2002ic1459},
% nuclear dust and an indication of a central stellar disk \citep{forbes1994ic1459}.
% The kinematically decoupled component cannot be identified as a separate structure in photometric observations.
% This galaxy has also a partially depleted core ($R_{\rm break} = 0''.69$, \citealt{rusli2013}).
%
% %\cite{sani2011} fit this galaxy with a S\'ersic-bulge + Gaussian-AGN model ($R_{\rm e} = 61''.1$, $n=6$).
% %\cite{lasker2014data} used a S\'ersic-bulge + mask-core model ($R_{\rm e} = 62''.4$, $n=8.25$).
% %\cite{rusli2013} adopted a core-S\'ersic model ($R_{\rm e} = 45''.43$, $n=7.6$).
%
% \emph{1D fit -- }
% Our isophotal analysis confirms a simple morphology for IC 1459, with no evident embedded components.
% We masked the innermost $6''.1$ affected by the partially depleted core 
% and we fit this galaxy with a S\'ersic profile. 
%
% \emph{Other models -- } 
%
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/ic1459_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for IC 1459.}
% \end{center}
% \end{figure}
%
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for IC 1459.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c & $63.1$  &  $6.6$ \\
% T.W. 1D eq.  & S-bul + m-c & $57.3$  &  $7.0$ \\
% T.W. 2D      & S-bul + m-c & $87.5$  &  $8.3$ \\
% \hline
% S+11         & S-bul + G-n   & $61.1$  &  $6.0$ \\
% R+13         & core-S\'ersic & $45.4$  &  $7.6$ \\
% L+14         & S-bul + m-c   & $62.4$  &  $8.3$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:ic1459}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf IC 2560 \\}
%
% IC 2560 is a barred spiral galaxy with a Seyfert AGN \citep{veroncettyveron2006}
% and dust within the central $2''.5$ \citep{martini2003}.
%
% %\cite{sani2011} used a S\'ersic-bulge + exponential-disk + Gaussian-AGN model ($R_{\rm e} = 27''.5$, $n=2$).
%
% \emph{1D fit -- }
% A visual inspection of the image of IC 2560 reveals a boxy bulge and
% a large scale bar that extends out to $R \lesssim 43''$.
% The disk appears to be slightly lopsided along the direction of the bar,
% due to two non-symmetric ansae, but it becomes symmetric beyond $R \gtrsim 60''$. 
% This is why the surface brightness profile deviates from a perfect exponential
% in the radial range $42'' \lesssim R \lesssim 60''$, which has been excluded from the fit.
% Motivated by the presence of a strong optical AGN and dust in the nucleus, 
% we accounted for an excess of non-stellar light by adding a central Gaussian component to the model.
% %Our final 1D model consists of a S\'ersic-bulge + exponential-disk + Ferrer-bar + Gaussian-AGN.
% %Similarly, our 2D model consists of a S\'ersic-bulge + exponential-disk + Gaussian-bar + Gaussian-AGN.\\
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/ic2560_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for IC 2560.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for IC 2560.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + G-n & $6.1$  &  $0.8$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + G-n & $4.5$  &  $0.6$ \\
% %T.W. 2D	&  &  & \\
% \hline
% S+11         & S-bul + e-d + G-n & $27.5$  &  $2.0$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:ic2560}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf IC 4296 \\}
%
% IC 4296 is an elliptical galaxy. % with a radio jet \citep{liuxie1992}.
% Due to its high stellar velocity dispersion, this galaxy is expected to host a partially depleted core. 
%
% %\cite{sani2011} fit this galaxy with a S\'ersic-bulge + Gaussian-AGN model ($R_{\rm e} = 33''.6$, $n = 4$).
% %\cite{lasker2014data} employed a S\'ersic-bulge + mask-core model ($R_{\rm e} = 97''.8$, $n=8.2$).
%
% \emph{1D fit -- }
% After masking the innermost $6''.1$, % affected by the (unresolved) partially depleted core,
% we found that a single S\'ersic profile provides a fairly good description of this galaxy.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/ic4296_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for IC 4296.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for IC 4296.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c & $65.9$  &  $5.8$ \\
% T.W. 1D eq.  & S-bul + m-c & $68.1$  &  $6.2$ \\
% T.W. 2D      & S-bul + m-c & $82.3$  &  $6.6$ \\
% \hline
% S+11         & S-bul + G-n & $33.6$  &  $4.0$ \\
% L+14         & S-bul + m-c & $97.8$  &  $8.2$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:ic4296}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf M31 -- NGC 0224 \\}
%
% M31, the Andromeda galaxy, is a spiral galaxy. 
% Although for decades this galaxy had been classified as an unbarred spiral,
% recent works have unraveled the presence of a bar \citep{athanassoulabeaton2006m31,beaton2007m31,morrison2011m31}.
% M31 also features a pseudo-ring-like structure at $R \sim 50'$ \citep{athanassoulabeaton2006m31}. 
%
%
% \emph{1D fit -- }
% We applied the \emph{smoothing} technique described in Section \ref{sec:smooth} to the analysis of M31.
% The region $2000'' \lesssim R \lesssim 3400''$, where the pseudo-ring is observed, was excluded from the fit.
% A S\'ersic + exponential fit is clearly not adequate to describe the light profile of M31.
% The residuals of such fit display a structure in correspondence of the bar ($R \lesssim 1500''$).
% The bar was thus fit with a Ferrer function.
%
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m31_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M31.}
% \end{center}
% \end{figure}
%
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M31.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar & $418.6$  &  $2.2$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar & $173.6$  &  $1.3$ \\
% %T.W. 2D	& $$  & $$  &  $$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m31}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf M49 -- NGC 4472 \\}
%
% M49, the brightest member of the Virgo cluster, is a giant elliptical galaxy 
% with a partially depleted core ($R_{\rm break} = 1''.8$, \citealt{rusli2013}).
% %M49 is a slow-rotator \citep{atlas3dIII-MNRAS}.
%
% %\cite{ferrarese2006acsvcs} modeled this galaxy with a core-S\'ersic profile ($R_{\rm e} = 208''$, $n = 5.5$).
% %Also \cite{rusli2013} used a core-S\'ersic model ($R_{\rm e} = 199''$, $n = 5.6$).
%
% \emph{1D fit -- }
% We excluded from the fit the innermost $12''$.
% We found no evidence for embedded components and
% we fit M49 with a single S\'ersic profile.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m49_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M49.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M49.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c & $190.2$  &  $6.6$ \\
% T.W. 1D eq.  & S-bul + m-c & $135.3$  &  $5.4$ \\
% T.W. 2D      & S-bul + m-c & $151.9$  &  $5.5$ \\
% \hline
% R+13         & core-S\'ersic & $199.0$  &  $5.6$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m49}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf M59 -- NGC 4621 \\}
%
% M59 is an elliptical galaxy with an edge-on, intermediate-scale embedded disk \citep{scorzabender1995}
% and a thin, faint nuclear stellar disk \citep{ferrarese2006acsvcs,ledo2010}.
% %M59 is a fast-rotator \citep{atlas3dIII-MNRAS}.
%
% %\cite{sani2011} modeled this galaxy with a single S\'ersic-bulge model ($R_{\rm e} = 70''.1$, $n = 5$).
% %\cite{vika2012} used a S\'ersic-bulge + mask-core model ($R_{\rm e} = 54''.7$, $n = 5.7$). 
%
% \emph{1D fit -- }
% The intermediate-scale embedded disk is clearly visible in our unsharp mask image, and 
% the velocity field confirms the presence of a rapidly rotating component. 
% Although faint, it is important to account for the presence of this disk.
% A single S\'ersic fit produces bad residuals.
% The fit notably improves if an embedded exponential disk function 
% is added to the final model.
% We chose not to account for the nuclear stellar disk by excluding the innermost $6''.1$ from the fit.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m59_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M59.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M59.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c + e-id & $48.0$  &  $5.5$ \\
% T.W. 1D eq.  & S-bul + m-c + e-id & $90.9$  &  $8.8$ \\
% %T.W. 2D	&    & $$  &  $$ \\
% \hline 
% S+11         & S-bul       & $70.1$  &  $5.0$ \\
% V+12         & S-bul + m-c & $54.7$  &  $5.7$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m59}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
%
% {\bf M60 -- NGC 4649 \\}
%
% exclude it, doesn't make sense
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf M64 -- NGC 4826 \\}
%
% M64 is a dusty spiral galaxy with a Seyfert AGN \citep{veroncettyveron2006}. 
%
% %B12 model the galaxy image with a 2D S\'ersic function + exponential disk ($n = 1.49$). \\
% \emph{1D fit -- }
% A large-scale bar ($R \lesssim 50''$) can be identified in the unsharp mask 
% and produces corresponding peaks in the ellipticity, PA and B4 profiles.
% We modeled it with a Ferrer function.
% Additional peaks in the ellipticity, PA and B4 profiles at $R \sim 15''$ 
% signal the presence of an embedded disky component, that we described with a low-$n$
% S\'ersic function.
% Although the galaxy is dusty and hosts an AGN, we do not observe any nuclear excess of light 
% in the fit residuals and therefore we consider unnecessary to add a nuclear component to our model.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m64_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M64.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M64.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + S-id  & $3.8$  &  $0.8$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + S-id  & $4.3$  &  $1.4$ \\
% %T.W. 2D	&   & $$  & $$  &  $$ \\
% \hline
% B+12      & S-bul + e-d  & $5.0$  &  $1.5$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m64}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf M81 -- NGC 3031 \\}
%
% M81 is an early-type spiral galaxy with a LINER AGN.
% This galaxy features two bars \citep{elmegreen1995m81,gutierrez2011,erwindebattista2013}.
%
% \emph{1D fit -- }
% We applied the \emph{smoothing} technique described in Section \ref{sec:smooth} to the analysis of M81.
% The bars are not particularly evident in the galaxy image or in the unsharp mask.
% Any attempt to account for the bars in the fit was unsuccessful. 
% For this reason, we used a simple S\'ersic + exponential model to describe M81.
%
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m81_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M81.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M81.}
% \begin{center}
% \begin{tabular}{llccc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d  & $31.0$  &  $1.7$ \\
% T.W. 1D eq.  & S-bul + e-d  & $33.2$  &  $2.1$ \\
% %T.W. 2D	&   & $$  & $$  &  $$ \\
% \hline
% GD07      & S-bul + e-d	 & $68.1$   &  $3.3$ \\
% S+11      & S-bul + e-d + G-n  & $127.3$  &  $3.0$ \\
% B+12      & S-bul + e-d	 & $50.0$   &  $2.6$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m81}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf M84 -- NGC 4374 \\}
%
% M84 is an elliptical galaxy with a partially depleted core ($R_{\rm break} = 1''.5$, \citealt{rusli2013}). 
%
% %GD07 model the galaxy light profile with a S\'ersic function ($n = 5.6$).
% %V12 model the galaxy image with a 2D S\'ersic function + core-mask ($n = 3.5$).
% %S11 model the galaxy image with a 2D S\'ersic function + Gaussian ($n = 7.0$).
% %B12 model the galaxy image with a 2D S\'ersic function ($n = 4.10$). \\
%
% \emph{1D fit -- }
% The unsharp mask reveals a faint inner component ($R \lesssim 12''$). 
% Here we excluded the data within $R < 13''.8$ and modeled the galaxy with a S\'ersic profile.
%
% %After masking the innermost $13''.8$, we modeled the galaxy with a S\'ersic profile. 
% %However, we notice that 
% %the 1D and 2D fits are not much affected by the presence of the embedded component.
% %In fact, reducing the size of the central mask to uncover the inner component 
% %(but still masking the region affected by the partially depleted core) does not
% %alter significantly the bulge parameters. 
% %The structures observed in the residuals are caused by the negative ellipticity gradient 
% %(from $\epsilon \sim 0.17$ to $\epsilon \sim 0.03$, over a radial range $10'' < R < 130''$),
% %which cannot be reproduced by the 2D model because of its fixed ellipticity ($\epsilon = 0.10$). 
%   
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m84_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M84.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M84.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c   & $101.6$  &  $7.8$ \\
% T.W. 1D eq.  & S-bul + m-c   & $129.8$  &  $7.9$ \\
% T.W. 2D      & S-bul + m-c   & $181.8$  &  $8.4$ \\
% \hline
% GD07         & S-bul         & $75.1$   &  $5.6$ \\
% S+11         & S-bul + G-n   & $105.9$  &  $7.0$ \\
% V+12         & S-bul + m-c   & $28.7$   &  $3.5$ \\
% B+12         & S-bul         & $63.6$   &  $4.1$ \\
% R+13         & core-S\'ersic & $126.2$  &  $7.1$ \\
% L+14         & S-bul + m-c   & $139.0$  &  $8.3$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m84}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf M87 -- NGC 4486 \\}
%
% M87 is a giant elliptical galaxy belonging to the Virgo cluster.
% This galaxy has a large partially depleted core ($R_{\rm break} = 7''.2$, \citealt{ferrarese2006acsvcs}).
%
% \emph{1D fit -- }
% The morphology of M87 is similar to that of M49.
% The innermost $18''.5$ were excluded from the fit.
% We used a single S\'ersic model to characterize M87.
%
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m87_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M87.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M87.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c  & $203.9$  &  $10.0$ \\
% T.W. 1D eq.  & S-bul + m-c  & $87.1$   &  $5.9$ \\
% T.W. 2D      & S-bul + m-c  & $88.3$   &  $4.3$ \\
% \hline
% GD07      & S-bul	    & $-$     &  $6.9$ \\
% S+11      & S-bul + G-n   & $99.5$  &  $4.0$ \\
% V+12      & S-bul + m-c   & $34.6$  &  $2.4$ \\
% R+13      & core-S\'ersic & $180.9$ &  $8.9$ \\
% L+14      & S-bul + m-c   & $122.0$ &  $5.6$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m87}
% \end{table}
%
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
%
% {\bf M89 -- NGC 4552 \\}
%
% m87 has ellipticity still increasing, the halo is not prevailing
% m89 instead has a plateau in the ell prof, bcz the halo is predominant
% we need to fit the halo
%
% BTW, if we wanna test the two phase formation scenario, we should care more about the halos...
% at which effective radius does the halo begin?
% what is the ratio of the mass in the galaxy or in the halo?
%
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m89_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M89.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M89.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-halo  & $29.0$  &  $4.6$ \\
% T.W. 1D eq.  & S-bul + e-halo  & $28.2$  &  $5.1$ \\
% %T.W. 2D	&   & $$  & $$  &  $$ \\
% \hline
% &   & $$  &  $$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m89}
% \end{table}
%
%
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf M94 -- NGC 4736 \\}
%
% M94 is an almost face-on spiral galaxy, with a Seyfert AGN \citep{veroncettyveron2006} and 
% circumnuclear dust \citep{elmegreen2002m94,peeplesmartini2006m94}.
% This galaxy has a ring-like structure at $R \sim 50''$ \citep{munoztunon1989m94}.
%
% %\cite{fisherdrory2010} fit M94 with a S\'ersic-bulge + exponential-disk model ($R_{\rm e} = 7''.7$, $n = 1.3$). 
%
% \emph{1D fit -- }
% The ellipticity and B4 profiles display a local maximum at $R \sim 10''$, 
% indicating the presence of a disky component embedded in the bulge,
% as already noted by \cite{fisherdrory2010}.
% We modeled this component with a Ferrer function.
% We excluded the data within the innermost $6''.7$, due to the AGN contribution.
% Any attempt to model the ring was unsuccessful, resulting in a degenerate model.
% Therefore we preferred to exclude also the data in the range $20'' \lesssim R \lesssim 80''$, 
% where the ring light dominates over the rest of the galaxy.
% %Our final model for M94 consists of a S\'ersic-bulge + exponential-disk + Ferrer-bar.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m94_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M94.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M94.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar & $11.4$  &  $0.9$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar & $8.4$   &  $1.1$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m94}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf M96 -- NGC 3368 \\}
%
% M96 is a spiral galaxy with a large scale bar and no AGN activity \citep{martini2003,nowak2010n3368n3489}. 
%
% %\cite{sani2011} fit this galaxy with a S\'ersic-bulge + exponential-disk + Gaussian-bar model 
% %($R_{\rm e} = 45''.6$, $n = 1.0$). 
% \cite{nowak2010n3368n3489} argued for the presence of a large pseudo-bulge and a tiny classical bulge 
% ($R_{\rm e} = 1''.6$, $n = 2.3$). 
%
% \emph{1D fit -- }
% M96 has a complex morphology. 
% The large-scale disk does not exhibit a perfect exponential profile, 
% mainly because of wound spiral arms resembling a ring ($R \sim 185''$),
% which however are too faint and irregular to be described with a Gaussian ring model. 
% The large scale bar extends out to $R \lesssim 80''$ and has been modeled with a Ferrer function.
% A peak at $R \sim 25''$ in the ellipticity profile unravels an inner component embedded in the bulge.
% The isophotal twist observed within $R \lesssim 25''$ suggests that the inner component can be identified 
% with a second bar, as already noted by \cite{erwin2004}.
% We accounted for this embedded component with a low-$n$ S\'ersic profile.
% We built our final model with a S\'ersic-bulge + exponential-disk + Ferrer-bar + S\'ersic-bar.
% The residuals do not display any significant light excess in the nuclear region, 
% in agreement with the lack of AGN activity.
%
% \cite{nowak2010n3368n3489} came to their conclusion by modeling only the innermost $8''.5$ of the 
% light profile of M96 with a S\'ersic-classical-bulge and an exponential-pseudo-bulge, 
% but without subtracting the contribution of the large-scale disk and the large-scale bar.
% In addition, they acknowledged the presence of a secondary inner bar, but did not include it in their fit.
%
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m96_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M96.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M96.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + S-bar  & $7.5$  &  $1.5$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + S-bar  & $5.3$  &  $1.3$ \\
% T.W. 2D      & S-bul + e-d + 2 G-bars       & $8.3$  &  $2.0$ \\
% \hline
% S+11         & S-bul + e-d + G-bar  & $45.6$  &  $1.0$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m96}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
%
% {\bf M104 -- NGC 4594 \\} 
%
% OK
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m104_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M104.}
% \end{center}
% \end{figure}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf M105 -- NGC 3379 \\}
%
% M105 is an elliptical galaxy with a partially depleted core ($R_{\rm break} = 1''.1$, \citealt{rusli2013}).
%
% \emph{1D fit -- }
% M105 exhibits a mild isophotal twist and an abrupt change of ellipticity at $R \sim 30''$.
% The velocity map shows rotation at least within $R \lesssim 30''$, 
% but no embedded disk can be recognized in the unsharp mask 
% and the ellipticity profile is quite unusual for a spheroidal system with an inner disk.
% After masking the innermost $6''.1$, we noticed that a single S\'ersic model 
% does not provide a good description of the galaxy light profile.
% However, the addition of a second function (of any analytical form) to the model did not improve the fit.
% We thus employed a single S\'ersic profile for M105.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m105_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M105.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M105.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c & $57.2$  &  $5.2$ \\
% T.W. 1D eq.  & S-bul + m-c & $50.9$  &  $5.3$ \\
% T.W. 2D      & S-bul + m-c & $73.6$  &  $7.0$ \\
% \hline
% GD07         & S-bul & $58.3$  &  $4.3$ \\
% S+11         & S-bul & $46.0$  &  $5.0$ \\
% R+13         & core-S\'ersic & $55.1$  &  $5.8$ \\
% L+14         & S-bul + m-c & $96.3$  &  $9.3$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m105}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf M106 -- NGC 4258 \\} 
%
% M106 is a barred spiral galaxy, harbouring a Seyfert AGN \citep{veroncettyveron2006} 
% with %a radio jet \citep{liuzhang2002} and 
% circumnuclear dust \citep{martini2003}. 
%
% %\cite{grahamdriver2007} modeled the galaxy light profile with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 14''.9$, $n = 2.0$). 
% %\cite{sani2011} used a S\'ersic-bulge + exponential-disk + Gaussian-AGN model 
% %($R_{\rm e} = 111''.7$, $n = 2.0$). 
% %\cite{vika2012} proposed two different fits: the first with a S\'ersic-bulge + exponential-disk + PSF-nucleus
% %($R_{\rm e} = 17''.0$, $n = 3.5$), 
% %and the second with a S\'ersic-bulge + exponential-disk + PSF-nucleus + S\'ersic-bar
% %($R_{\rm e} = 6''.3$, $n = 2.2$).
%
% \emph{1D fit -- }
% We excluded from the fit the innermost $6''.1$ because affected by the AGN emission.
% The boxy bar, responsible for the minimum in the B4 profile at $R \sim 65''$, 
% is faint compared to the large-scale disk and in the light profile is only detectable as a slight swelling
% within $65'' \lesssim R \lesssim 110''$.
% This region has been excluded from the fit.
% A peak in the ellipticity profile at $R \sim 10''$ suggests that a disk-like component might be 
% embedded in the bulge.
% Since the galaxy is barred, this component could be identified with a peanut bulge. 
% However, we found that adding an extra component to the model introduces degeneracies, 
% thus we preferred not to do so.
% While the fit along the major-axis is satisfactory, 
% the equivalent-axis residuals present a sneaky structure within $R \lesssim 20''$,
% due to the unsubtracted embedded component.
% %Our final model consists of a S\'ersic-bulge + exponential-disk.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/m106_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for M106.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for M106.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d & $15.3$  &  $2.0$ \\
% T.W. 1D eq.  & S-bul + e-d & $8.3$   &  $1.2$ \\
% %T.W. 2D	& & $$  & $$  &  $$ \\
% \hline
% GD07     & S-bul + e-d		 & $14.9$   &  $2.0$ \\
% S+11     & S-bul + e-d + G-n  	 & $111.7$  &  $2.0$ \\
% V+12 (1) & S-bul + e-d + PSF-n	 & $17.0$   &  $3.5$ \\
% V+12 (2) & S-bul + e-d + PSF-n + S-bar & $6.3$    &  $2.2$ \\
% L+14     & S-bul + e-d + PSF-n + e-id + S-bar + spiral arms & $6.3$	 &  $3.3$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:m106}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 0524 \\} 
%
% NGC 0524 is an unbarred face-on lenticular galaxy. 
%
% %\cite{laurikainen2010} modeled this galaxy with a S\'ersic-bulge + exponential-disk + 2 Ferrer-lenses 
% %($R_{\rm e} = 8''.9$, $n = 2.7$).
% %\cite{sani2011} opted for a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 26''.8$, $n = 3.0$).
% \emph{1D fit -- }
% The unsharp mask of NGC 0524 unravels a faint multi-ring structure in the galaxy disk,
% with a major ring peaking at $R \sim 20''$.
% We accounted only for this brightest ring using a Gaussian ring profile.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n0524_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 0524.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 0524.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + G-r & $6.0$  &  $1.1$ \\
% T.W. 1D eq.  & S-bul + e-d + G-r & $5.8$  &  $1.1$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% L+10      & S-bul + e-d + 2 F-l & $8.9$  &  $2.7$ \\
% S+11      & S-bul + e-d	  & $26.8$  &  $3.0$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n0524}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
%
% {\bf NGC 0821 \\}
%
% ok
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n0821_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 0821.}
% \end{center}
% \end{figure}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 1023 \\}
%
% NGC 1023 is a barred lenticular galaxy. % with no partially depleted core \cite{faber1997}. \\
%
% %\cite{grahamdriver2007} modeled the galaxy light profile with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 17''.7$, $n = 2.0$). 
% %\cite{sani2011} used a S\'ersic-bulge + exponential-disk + Gaussian-bar model 
% %($R_{\rm e} = 24''.0$, $n = 3.0$).
% %\cite{lasker2014data} employed a S\'ersic-bulge + exponential-disk + S\'ersic-bar 
% %($R_{\rm e} = 9''.6$, $n = 3.1$).
% \emph{1D fit -- }
% The bar extends out to $R \lesssim 40''$.
% %The light profile of NGC 1023 is well described by a S\'ersic-bulge + exponential-disk + Ferrer-bar model.
% Small deviations of the disk from a perfect exponential profile, 
% within the region $60'' \lesssim R \lesssim 125''$ (that has been excluded from the fit), 
% can be ascribed to some faint residual spiral arms, also noticeable in the 2D residual image. 
% The peak at $R \sim 10''$ in the B4 profile signals the presence of an embedded (and faint) disk.
% The 1D residuals show a structure within $R \lesssim 10''$ caused by this unsubtracted 
% component, also noted by \cite{lasker2014data}.
% Attempts to account for the inner disk by adding a fourth function to the model 
% did not significantly modify the best-fit bulge parameters,
% thus we elected not to model the embedded disk.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n1023_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 1023.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 1023.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar & $9.2$  &  $2.1$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar & $7.4$  &  $2.0$ \\
% T.W. 2D      & S-bul + e-d + G-bar & $6.6$  &  $2.3$ \\
% \hline
% GD07    & S-bul + e-d 	& $17.7$  &  $2.0$ \\
% S+11    & S-bul + e-d + G-bar & $24.0$  &  $3.0$ \\
% L+14    & S-bul + e-d + S-bar & $9.6$   &  $3.1$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n1023}
% \end{table}
%
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 1300 \\}
%
% NGC 1300 is a face-on, barred spiral galaxy. % with no partially depleted core \citep{graham2012MNRAS}. 
%
% %\cite{sani2011} fit NGC 1300 with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 85''.4$, $n = 3.0$).
% %\cite{lasker2014data} built their model with a S\'ersic-bulge + exponential-disk + PSF-nucleus + exponential-inner-disk + 
% %S\'ersic-bar + spiral arms
% %($R_{\rm e} = 10''.4$, $n = 4.3$).
%
% \emph{1D fit -- }
% The morphology of NGC 1300 is quite complex. 
% In addition to a large scale disk and a bulge, the galaxy is composed of 
% a large scale bar that extends up to $R \lesssim 90''$, two prominent spiral arms 
% and an inner disk-like component ($R \lesssim 5''$), disclosed by the peaks in the ellipticity and PA profiles.
% We truncate the light profile at $R \sim 50''$ and we fit the inwards combination of the large-scale bar and the disk 
% as a single component, using an exponential function.
% The embedded component is described with a low-$n$ S\'ersic profile.
% %Our final model consists of a S\'ersic-bulge + exponential-(disk+bar) + S\'ersic-inner disk.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n1300_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 1300.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 1300.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-(d+bar) + S-id & $9.9$  &  $3.8$ \\
% T.W. 1D eq.  & S-bul + e-(d+bar) + S-id & $8.1$  &  $3.6$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% S+11    & S-bul + e-d & $85.4$  &  $3.0$ \\
% L+14    & S-bul + e-d + PSF-n + e-id + S-bar + spiral arms & $10.4$  &  $4.3$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n1300}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 1316 \\}
%
% ok
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n1316_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 1316.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 1316.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + 2 G-r + G-n & $21.5$  &  $2.0$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + 2 G-r + G-n & $15.9$  &  $1.8$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% S+11      & S-bul + e-d + G-n & $93.0$  &  $5.0$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n1316}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 1332 \\}
%
% ok
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n1332_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 1332.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 1332.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + S-id + m-c & $34.7$  &  $5.1$ \\
% T.W. 1D eq.  & S-bul + S-id + m-c & $18.0$  &  $3.7$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n1332}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 1374 \\}
%
% ok
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 1399 \\}
%
% NGC 1399, the central galaxy of the Fornax cluster, is an elliptical galaxy with a partially depleted core 
% ($R_{\rm break} \sim 2''.4$, \citealt{rusli2013,dullograham2014cores}).
% The nuclear activity of NGC 1399 is classified as Seyfert \citep{veroncettyveron2006},
% but the galaxy lacks of dust \citep{tran2001}.
%
% %\cite{grahamdriver2007} fit the galaxy light profile with a single S\'ersic-bulge profile 
% %($n = 16.8$).  
% %\cite{rusli2013} modeled NGC 1399 with a core-S\'ersic profile and the sum of a S\'ersic and an exponential
% %function accounting for the galaxy halo 
% %($R_{\rm e} = 36''.2$, $n = 7.38$). 
% %\cite{lasker2014data} used a S\'ersic-bulge + mask-core model 
% %($R_{\rm e} = 154''.0$, $n = 11.1$). 
%
% \emph{1D fit -- }
% The ellipticity and PA profiles display a steep decline within $R \lesssim 60''$,
% suggesting the presence of an embedded disk.
% This inner component is also visible, although very faint, in the unsharp mask.
% TBD!! velocity map, the disk is face on
% We note that, after excluding the innermost $6''.1$ affected by the partially depleted core,
% a single S\'ersic profile is not sufficient to describe the whole galaxy light profile.
% The addition of an inner exponential function to the model 
% notably improves the fit.
% %We build our final model with a S\'ersic-bulge + mask-core + exponential-inner disk.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n1399_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 1399.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 1399.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c + e-id & $405.2$  &  $10.0$ \\
% T.W. 1D eq.  & S-bul + m-c + e-id & $338.1$  &  $10.0$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% GD07         & S-bul  		    & $-$      &  $16.8$ \\
% R+13         & core-S\'ersic + (S+e)-halo & $36.2$   &  $7.4$ \\
% L+14         & S-bul + m-c		    & $154.0$  &  $11.1$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n1399}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 2273 \\}
%
% NGC 2273 is a barred spiral galaxy with a Seyfert AGN \citep{contini1998} 
% and circumnuclear dust \citep{simoeslopes2007}. 
% Its bar is surrounded by two tightly wound star-forming spiral arms that resemble a ring
% \citep{comeron2010}. 
%
% \cite{laurikainen2005} pointed out the presence of an inner disk embedded in the bulge of NGC 2273,
% but they deliberately chose not to model it due to the degeneracy between the functions describing the bulge
% and the inner disk. 
% %Their final model consists of a S\'ersic-bulge + exponential-disk + Ferrer-bar 
% %($R_{\rm e} = 2''.6$, $n = 1.8$). 
%
% \emph{1D fit -- }
% The bar of NGC 2273 extends out to $R \lesssim 25''$. 
% The pseudo-ring does not produce any evident swelling in the light profile, 
% therefore we do not account for it in the galaxy model.
% The isophotal parameters confirm the presence of a nuclear disk-like component within $R \lesssim 5''$.
% However, as noted by \cite{laurikainen2005}, any attempt to account for the embedded disk resulted in a 
% degenerate model. 
% This is not surprising if one considers the poor spatial resolution of the galaxy image, 
% the effective radius of the bulge being comparable to the FWHM of the PSF.
% Although NGC 2273 hosts an optical AGN and nuclear dust, no central excess of light is observed in the 1D residuals.
% The addition of a nuclear component to the model does not improve the fit and does not significantly affect
% the bulge parameters.
% %Our final model consists of a S\'ersic-bulge + exponential-disk + Ferrer-bar.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n2273_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 2273.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 2273.}
% \begin{center}
% \begin{tabular}{llccc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar & $1.6$  &  $2.1$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar & $1.9$  &  $2.7$ \\
% %T.W. 2D	&  & $$  &  $$ \\
% \hline
% L+05         & S-bul + e-d + F-bar & $2.6$  &  $1.8$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n2273}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 2549 \\}
%
% NGC 2549 is an edge-on barred lenticular galaxy. 
%
% %\cite{sani2011} modeled this galaxy with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 11''.6$, $n = 7.0$). 
%
% \emph{1D fit -- }
% Although the edge-on inclination of NGC 2549 complicates the identification of embedded components, 
% a large scale bar ($R \lesssim 45''$) can be recognized in the galaxy image itself and -- more easily -- in the light profile.
% We describe the large scale bar with a Ferrer profile.
% A peak in the ellipticity profile discloses the presence of a disky component
% embedded in the bulge ($R \lesssim 10''$).
% This inner component can be spotted also by looking at the velocity map.
% We fit it with a Ferrer profile.
% %We build our final model with a S\'ersic-bulge + exponential-disk + 2 Ferrer-bars.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n2549_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 2549.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 2549.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + F-id & $6.1$  &  $2.3$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + F-id & $3.1$  &  $1.5$ \\
% T.W. 2D      & S-bul + e-d + G-bar + G-id & $5.6$  &  $2.1$ \\
% \hline
% S+11         & S-bul + e-d & $11.6$  &  $7.0$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n2549}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 2778 \\}
%
% NGC 2778 is an almost face-on lenticular galaxy. 
%
% %\cite{grahamdriver2007} modeled NGC 2778 with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 2''.3$, $n = 1.6$). 
% %\cite{sani2011} used a S\'ersic-bulge + exponential-disk model
% %($R_{\rm e} = 2''.5$, $n = 2.5$). 
% %\cite{vika2012} employed a S\'ersic-bulge + exponential-disk model 
% %($R_{\rm e} = 1''.5$, $n = 2.7$). 
% %\cite{lasker2014data} fit the galaxy with a S\'ersic-bulge + exponential-disk + S\'ersic-bar
% %($R_{\rm e} = 2''.8$, $n = 4.0$). 
%
% \emph{1D fit -- }
% The peak in the ellipticity profile at $R \sim 5''$ unravels the existence of a nuclear component 
% embedded in the galaxy bulge.
% After an inspection of the unsharp mask, we identified this component with a small bar,
% that we modeled with a Ferrer function.
% We also accounted for some nuclear light excess by adding a PSF component to the model.
% %We then fit NGC 2778 with a S\'ersic-bulge + exponential-disk + Ferrer-bar + PSF-nucleus.
% We note that excluding the PSF component from our galaxy model does not significantly affect the 
% bulge effective radius, but it increases the S\'ersic index to $n \sim 2$.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n2778_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 2778.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 2778.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + PSF-n & $2.3$  &  $1.3$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + PSF-n & $2.2$  &  $1.2$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% GD07         & S-bul + e-d	     & $2.3$  &  $1.6$ \\
% S+11         & S-bul + e-d	     & $2.5$  &  $2.5$ \\
% V+12         & S-bul + e-d	     & $1.5$  &  $2.7$ \\
% L+14         & S-bul + e-d + S-bar & $2.8$  &  $4.0$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n2778}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 2787 \\}
%
% NGC 2787 is a barred lenticular galaxy with LINER nuclear activity \citep{veroncettyveron2006}.
% An inner disk is embedded in the bulge of this galaxy \citep{erwin2003n2787}.
% NGC 2787 features a spectacular structure of dust rings at its center \citep{erwinsparke2003}
% and a nuclear stellar disk ($R \lesssim 1''.8$, \citealt{ledo2010}).
%
% %\cite{grahamdriver2007} modeled the galaxy light profile with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 4''.6$, $n = 2.0$). 
% %\cite{fisherdrory2008} decomposed NGC 2787 in a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 9''.1$, $n = 1.2$).
% %\cite{laurikainen2010} employed a S\'ersic-bulge + exponential-disk + Ferrer-bar + Ferrer-lens model
% %($R_{\rm e} \sim 4''$, $n = 1.3$).
% %\cite{sani2011} used a S\'ersic-bulge + exponential-disk + Gaussian-bar + Gaussian-AGN 
% %($R_{\rm e} = 15''.7$, $n = 3.0$). 
% %\cite{lasker2014data} described NGC 2787 with a S\'ersic-bulge + truncated exponential-disk + 
% %truncated S\'ersic-bar + S\'ersic-inner disk + PSF-nucleus
% %($R_{\rm e} = 14''.3$, $n = 2.8$). 
%
% \emph{1D fit -- }
% NGC 2787 has an undoubtedly complex morphology. 
% This galaxy is made of a bulge, a large scale disk, a bar (see the peaks in the ellipticity, PA and B4 profiles
% at $R \sim 30''$) with two ansae, an inner disk (or bar?) embedded in the bulge (see the peak in the ellipticity profile
% at $R \sim 17''$), and a bright nucleus.
%
% Beyond $R \gtrsim 50''$, where the galaxy large scale disk extends without being affected by the bar ansae,
% the surface brightness profile is not a perfect exponential, being better described by a S\'ersic function
% with $n < 1$.
% Curiously, at optical wavelengths the disk has an exponential decline with no truncation or anti-truncation 
% (see \citealt{erwin2003n2787} their Figure 14, and \citealt{erwin2008}).
% Here we propose two different models for NGC 2787.
% In both cases we used a S\'ersic profile for the bulge, a Ferrer function for the bar, a second Ferrer function for the inner disk (or bar) 
% and a PSF for the nucleus. 
% We then described the large scale disk using a S\'ersic function with $n < 1$ 
% or an exponential function. 
% In the last case we needed to exclude the data within $R \sim 30'' - 50''$ where the ansae produce a hollow.
% The two different decompositions result in the same bulge best-fit parameters ($R_{\rm e} \sim 5''$, $n \sim 1$)
% along the major-axis, whereas the outcome is significantly different along the equivalent-axis 
% ($n = 1.3$ when using the exponential-disk and $n = 2.2$ when using the S\'ersic-disk).
%
% might just pick one TBD!!
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n2787_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 2787.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 2787.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% \multicolumn{4}{c}{\emph{exponential-disk}} \\
% %\hline
% T.W. 1D maj. & S-bul + e-d + F-bar + F-id + PSF-nucleus & $4.8$  &  $1.1$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + F-id + PSF-nucleus & $3.3$  &  $1.3$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% \multicolumn{4}{c}{\emph{S\'ersic-disk}}\\
% %\hline
% T.W. 1D maj. & S-bul + S-d + F-bar + F-id + PSF-nucleus & $4.6$  &  $1.1$ \\
% T.W. 1D eq.  & S-bul + S-d + F-bar + F-id + PSF-nucleus & $4.4$  &  $2.2$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% GD07         & S-bul + e-d					  & $4.6$   &  $2.0$ \\
% L+10         & S-bul + e-d + F-bar + F-l			  & $4.0$   &  $1.3$ \\
% S+11         & S-bul + e-d + G-bar + G-n			  & $15.7$  &  $3.0$ \\
% L+14         & S-bul + trunc. e-d + trunc. S-bar + S-id + PSF-n & $14.3$  &  $2.8$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n2787}
% \end{table}
%
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 2974 \\}
%
% NGC 2974 is a spiral galaxy which has been misclassified as an elliptical galaxy in the RC3 catalog.
% The galaxy hosts a Seyfert AGN \citep{veroncettyveron2006} and filamentary dust in its center \citep{tran2001}.
%  
% %\cite{sani2011} modeled this galaxy with a S\'ersic-bulge + Gaussian-AGN 
% %($R_{\rm e} = 27''.2$, $n = 3.0$). 
%
% \emph{1D fit -- }
% From an inspection of the unsharp mask, we identified a ring structure at $R \sim 50''$, 
% probably a residual of two tightly wound spiral arms,
% and an elongated bar-like component within $R \lesssim 30''$, that we fit with a Ferrer function.
% The last one produces a peak in the ellipticity profile at $R \sim 20''$.
% Although the ring is extremely faint, it is important to account for it in the galaxy decomposition. 
% A model without the ring component results in a ``steeper'' exponential-disk (the disk has a smaller scale length 
% and a lower central surface brightness) and produces bad residuals within $R \lesssim 40''$.
% The nuclear component is fit with a Gaussian profile.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n2974_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 2974.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 2974.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + G-n + G-r & $8.3$   &  $1.4$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + G-n + G-r & $6.9$   &  $1.2$ \\
% T.W. 2D      & S-bul + e-d + G-bar + m-n	 & $10.6$  &  $1.3$ \\
% \hline
% S+11         & S-bul + G-n & $27.2$  &  $3.0$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n2974}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
%   
% {\bf NGC 3079 \\}
%
% inner component?? ok fuck it
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3079_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3079.}
% \end{center}
% \end{figure}
%
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 3091 \\}
%
% NGC 3091 is an elliptical galaxy with a partially depleted core ($R_{\rm break} = 0''.62$, \citealt{rusli2013}).
%
% %\cite{rusli2013} modeled NGC 3091 with a single core-S\'ersic profile 
% %($R_{\rm e} = 91''.0$, $n = 9.3$). 
%
% \emph{1D fit -- }
% The data within the innermost $6''.1$ are excluded from the fit.
% A faint embedded component ($R \lesssim 10''$) can be recognized in the unsharp mask and has a corresponding peak in the
% ellipticity profile.
% This extra component is clearly visible in the residual image obtained by subtracting a 2D S\'ersic model from the  
% galaxy image.
% However, any attempt to account for the embedded component resulted in an unsatisfactory fit and did not 
% significantly affect the bulge parameters.
% We thus elected not to model the inner component.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3091_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3091.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 3091.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c & $100.5$  &  $7.6$ \\
% T.W. 1D eq.  & S-bul + m-c & $51.2$	&  $6.6$ \\
% T.W. 2D      & S-bul + m-c & $67.1$	&  $6.7$ \\
% \hline
% R+13         & core-S\'ersic & $91.0$  &  $9.3$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n3091}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 3115 \\}
%
% NGC 3115 is an edge-on elliptical/lenticular galaxy. 
% Its nuclear activity is classified as passive in the optical wavelengths \citep{gavazzi2013nuclspec}.
% This galaxy features a nuclear stellar disk ($R \sim 1''$, \citealt{scorzabender1995,ledo2010}).
%
% \cite{grahamdriver2007} TBD!!! ask Alister what did they do, not clear from appendix
% ($n = 13.0$). 
% \cite{fisherdrory2008} decomposed the galaxy in a S\'ersic-bulge + exponential disk 
% ($R_{\rm e} = 31''.0$, $n = 3.9$). 
% %\cite{sani2011} modeled NGC 3115 with a S\'ersic-bulge + exponential disk 
% %($R_{\rm e} = 27''.3$, $n = 3.0$). 
% %\cite{lasker2014data} used a S\'ersic-bulge + exponential disk + S\'ersic-envelope 
% %($R_{\rm e} = 3''.9$, $n = 3.0$). 
%
% \emph{1D fit -- }
% NGC 3115 is an example of a galaxy with an intermediate scale disk.
% The surface brightness profile of NGC 3115 does not have an exponential decline at large radii 
% because the disk component does not dominate the outer regions of the galaxy.
% This is immediately evident by looking at the ellipticity and B4 profiles.
% Going from the galaxy center to the outskirts, the ellipticity increases until it reaches its maximum 
% ($\epsilon \sim 0.65$) at $R \sim 90''$. 
% After this point, it starts declining, being as low as $\epsilon \sim 0.55$ at $R \sim 240''$. 
% The galaxy isophotes display a positive diskyness within $R \sim 90''$, but become perfect ellipses 
% ($B4 \sim 0$) beyond that point.
% Now, because galaxy disks typically have fixed ellipticity, reflecting their inclination to our line of sight,
% the previous observations tell us that, going from the galaxy center to the outer regions,
% the disk of NGC 3115 becomes increasingly important relative to the spheroid light, reaching its maximum at $R \sim 90''$.
% Beyond $R \gtrsim 90''$, the contribution from the disk light starts declining more rapidly than the spheroid light.
% This interpretation is supported by the results of \cite{arnold2011n3115},
% who found that within $R \sim 110''$ the bulge of NGC 3115 is flattened and rotates rapidly ($v/\sigma \gtrsim 1.5$),
% whereas at larger radii the rotation declines dramatically (to $v/\sigma \sim 0.7$), 
% but remains well aligned with the inner regions. \\
% From an inspection of the unsharp mask, we identify a faint nuclear edge-on ring ($R \sim 15''$), 
% which produces a corresponding peak in the ellipticity profile. \\
% %We build our model with a S\'ersic-bulge + exponential disk + Gaussian-ring. \\
% We do not observe any nuclear excess of light in the 1D residuals, 
% contrary to what we expected given the presence of a nuclear stellar disk. 
% The addition of a PSF component to the final model does not significantly change the outcome of the fit. 
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3115_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3115.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 3115.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + G-r & $43.6$  &  $4.4$ \\
% T.W. 1D eq.  & S-bul + e-d + G-r & $34.4$  &  $5.1$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% GD07         & ??? TBD	       & $-$	&  $13.0$ \\
% S+11         & S-bul + e-d	       & $27.1$ &  $3.0$ \\
% L+14         & S-bul + e-d + S-halo  & $3.9$  &  $3.0$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n3115}
% \end{table}
%   
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 3227 \\}
%
% NGC 3227 is a spiral galaxy with a large-scale bar. 
% This galaxy hosts a Seyfert AGN \citep{khachikian1974} and circumnuclear dust \citep{martini2003}. 
%
% %\cite{laurikainen2004} decomposed NGC 3227 into a S\'ersic-bulge + exponential-disk + Ferrer-bar 
% %($R_{\rm e} = 1''.8$, $n = 2.2$). 
% %\cite{sani2011} described this galaxy with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 82''.9$, $n = 4.0$). 
% %\cite{lasker2014data} used a S\'ersic-bulge + exponential disk + S\'ersic-bar 
% %($R_{\rm e} = 0''.7$, $n = 4.1$). 
%
% \emph{1D fit -- }
% The large-scale bar produces an evident bump in the major-axis surface brightness profile,
% which is curiously absent in the equivalent-axis profile. \\
% We modeled the bar with a Ferrer function and the nucleus with a Gaussian function. 
% As expected, the Ferrer component is rejected by the equivalent-axis fit.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3227_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3227.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 3227.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + G-n   & $8.1$  &  $1.7$ \\
% T.W. 1D eq.  & S-bul + e-d + [F-bar] + G-n & $4.6$  &  $1.1$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% L+04         & S-bul + e-d + F-bar & $1.8$  &  $2.2$ \\
% S+11         & S-bul + e-d	     & $82.9$ &  $4.0$ \\
% L+14         & S-bul + e-d + S-bar & $0.7$  &  $4.1$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n3227}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 3245 \\}
%
% NGC 3245 is a lenticular galaxy.
%
% %\cite{grahamdriver2007} modeled this galaxy with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 11''.3$, $n = 4.3$). 
% %\cite{laurikainen2010} described the galaxy with a S\'ersic-bulge + exponential-(disk+lens) + Ferrer-inner disk 
% %($R_{\rm e} \sim 4''$, $n = 2.4$).  
% %\cite{sani2011} used a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 4''.6$, $n = 2.5$). 
% %\cite{vika2012} employed a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 3''.5$, $n = 2.6$). 
% %\cite{beifiori2012} decomposed the galaxy into a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 4''.00$, $n = 1.6$). 
% %\cite{lasker2014data} built their model with a S\'ersic-bulge + exponential-disk + S\'ersic-bar 
% %($R_{\rm e} = 1''.95$, $n = 1.6$). 
%
% \emph{1D fit -- }
% The unsharp mask reveals the presence of an embedded disk ($R \lesssim 15''$), 
% confirmed by the corresponding peaks in the ellipticity and B4 profiles.
% We model this inner component with a Ferrer function.
% %We model the galaxy with a S\'ersic-bulge + exponential-disk + Ferrer-inner disk.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3245_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3245.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 3245.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-id & $4.4$  &  $2.9$ \\
% T.W. 1D eq.  & S-bul + e-d + F-id & $2.4$  &  $1.7$ \\
% T.W. 2D      & S-bul + e-d + G-id & $1.9$  &  $1.8$ \\
% \hline
% GD07         & S-bul + e-d		& $11.3$ &  $4.3$ \\
% L+10         & S-bul + e-(d+l) + F-id & $4.0$  &  $2.4$ \\
% S+11         & S-bul + e-d		& $4.6$  &  $2.5$ \\
% V+12         & S-bul + e-d		& $3.5$  &  $2.6$ \\
% B+12         & S-bul + e-d		& $4.0$  &  $1.6$ \\
% L+14         & S-bul + e-d + S-bar	& $2.0$  &  $1.6$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n3245}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
%
% {\bf NGC 3377 \\}
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3377_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3377.}
% \end{center}
% \end{figure}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 3384 \\}
%
% NGC 3384 is a barred lenticular galaxy that hosts 
% an embedded disk ($R \lesssim 11''$, \citealt{erwin2004}) and 
% a nuclear stellar disk ($R \lesssim 0''.8$, \citealt{ledo2010}).
%
% %\cite{grahamdriver2007} fit this galaxy with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 2''.5$, $n = 1.7$). 
% %\cite{laurikainen2010} described the galaxy with a S\'ersic-bulge + exponential-disk + 2 Ferrer-bars
% %($R_{\rm e} \sim 4''$, $n = 1.5$).  
% %\cite{sani2011} employed a S\'ersic-bulge + exponential-disk + Gaussian-bar 
% %($R_{\rm e} = 4''.4$, $n = 2.5$). 
% %\cite{beifiori2012} modeled the galaxy with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 8''.3$, $n = 2.3$). 
% %\cite{lasker2014data} decomposed the galaxy into a S\'ersic-bulge + exponential-disk + S\'ersic-inner disk + 
% %S\'ersic-nuclear disk + S\'ersic-bar
% %($R_{\rm e} = 5''.9$, $n = 2.5$). 
%
% \emph{1D fit -- }
% The isophotal parameters clearly show the presence of a boxy bar that extends out to $R \lesssim 18''$.
% We model the bar with a Ferrer function.
% An embedded disk can be seen in the unsharp mask and produces the peak in the ellipticity profile at $R \sim 4''$.
% modeling the embedded disk turned out to be a complicated task, resulting in a degenerate model.
% We elected not to fit the embedded disk (nor the nuclear disk) and we therefore excluded from the fit the innermost $3''.7$,
% where there two components are expected to dominate.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3384_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3384.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 3384.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + m-n & $5.5$  &  $1.6$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + m-n & $5.6$  &  $1.8$ \\
% %T.W. 2D	&  & $$  &  $$ \\
% \hline
% GD07         & S-bul + e-d		      & $2.5$  &  $1.7$ \\
% L+10         & S-bul + e-d + 2 F-bar        & $4.0$  &  $1.5$ \\
% S+11         & S-bul + e-d + G-bar	      & $4.4$  &  $2.5$ \\
% B+12         & S-bul + e-d		      & $8.3$  &  $2.3$ \\
% L+14         & S-bul + e-d + 2 S-id + S-bar & $5.9$  &  $2.5$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n3384}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 3393 \\}
%
% NGC 3393 is an edge-on spiral galaxy hosting a Seyfert AGN \citep{diaz1988n3393} and 
% circumnuclear dust \citep{martini2003}.
% The galaxy has a large-scale bar and a nuclear bar ($R \lesssim 3''$, \citealt{erwin2004}).  
%
% \emph{1D fit -- }
% The large-scale bar ($R \lesssim 20''$) is modeled with a Ferrer function.
% The poor spatial resolution of the data does not allow to fit the nuclear bar and/or the AGN component 
% without incurring degeneracies.
% We thus elected not to fit the nuclear components.
% We note that the addition to the model of a central component would lower the bulge S\'ersic index,
% whereas the the addition of a nuclear bar component would increase the bulge S\'ersic index.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3393_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3393.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 3393.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar & $1.4$  &  $3.4$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar & $1.4$  &  $2.6$ \\
% T.W. 2D      & S-bul + e-d + G-bar & $1.2$  &  $1.9$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n3393}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 3414 \\}
%
% NGC 3414 is a peculiar lenticular/elliptical galaxy. 
% Different interpretations have been proposed about the morphology of this galaxy.
% \cite{whitmore1990} suggested that NGC 3414 is an edge-on galaxy 
% with a large scale polar ring, rather than a face-on barred lenticular.
% \cite{laurikainen2010} decomposed NGC 3414 as a barred lenticular galaxy, but 
% they cautioned against the uncertainty of their solution due to the possible misinterpretation of the 
% galaxy morphology.
%
% \emph{1D fit -- }
% The unsharp mask clearly shows that the edge on component does not extend all the way through the 
% center of the galaxy, but instead it is truncated at $R \sim 20''$.
% However, an elongated structure in the galaxy center resembles a nuclear edge-on disk ($R \lesssim 5''$).
% The velocity map shows rotation within $R \lesssim 5''$ and no rotation within $5'' \lesssim R \lesssim 15''$.
% Beyond $R \gtrsim 15''$ one can recognize a sign of rotation in the opposite direction of the nuclear region.
% The velocity pattern is consistent with the galaxy being a spheroidal system containing a polar ring and a nuclear disk
% both seen edge-on.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3414_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3414.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 3414.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-id + G-r & $28.0$  &  $4.8$ \\
% T.W. 1D eq.  & S-bul + e-id + G-r & $25.5$  &  $4.5$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% L+10         & S-bul + e-d + F-bar	  & $5.0$  &  $2.6$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n3414}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 3489 \\}
%
% NGC 3489 is a barred lenticular galaxy. 
% Its disk has a smooth antitruncation at $R \sim 85''$ \citep{erwin2008}.
%
% \cite{nowak2010n3368n3489} argued for the presence of both a pseudo-bulge and a classical bulge. 
% %\cite{laurikainen2010} fit this galaxy with a S\'ersic-bulge + exponential-disk + Ferrer-bar
% %($R_{\rm e} \sim 2''$, $n = 2.1$). 
% %\cite{sani2011} used a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 4''.6$, $n = 1.5$). 
%
% \emph{1D fit -- }
% We excluded the data beyond $R \gtrsim 38''$ to avoid fitting the antitruncation region.
% The bar is fit with a Ferrer function.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3489_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3489.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 3489.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar & $2.2$  &  $1.5$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar & $1.7$  &  $1.3$ \\
% T.W. 2D      & S-bul + e-d + G-bar & $1.7$  &  $2.1$ \\
% \hline
% L+10         & S-bul + e-d + F-bar & $2.0$  &  $2.1$ \\
% S+11         & S-bul + e-d & $4.6$  &  $1.5$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n3489}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
%
% {\bf NGC 3585 \\}
%
% ok
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3585_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3585.}
% \end{center}
% \end{figure}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
%
% {\bf NGC 3607 \\}
%
% ok
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3607_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3607.}
% \end{center}
% \end{figure}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
%
% {\bf NGC 3608 \\}
%
% ok
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3608_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3608.}
% \end{center}
% \end{figure}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 3842 \\}
%
% NGC 3842 is the brightest member of the Leo Cluster (Abell 1367).
% It is an elliptical galaxy with a partially depleted core ($R_{\rm break} = 0''.7$, 
% \citealt{rusli2013,dullograham2014cores}).
%
% %\cite{rusli2013} modeled this galaxy with a core-S\'ersic model 
% %($R_{\rm e} = 58''.8$, $n = 6.3$). 
%
% \emph{1D fit -- }
% The light profile presents a swelling at $R \sim 20''$ and the ellipticity profile has a peak in the same position.
% One can also glimpse a faint elongated oval in the unsharp mask.
% The innermost $6''.1$ are excluded from the fit due to the presence of a partially depleted core.
% As expected, a single S\'ersic model is not sufficient to provide a good description of the galaxy light profile 
% and we need to add a second component (Ferrer function) to account for the embedded lens.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3842exp_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3842.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 3842.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + F-l + m-c  & $100.7$  &  $8.1$ \\
% T.W. 1D eq.  & S-bul + F-l + m-c  & $73.6$   &  $8.2$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% R+13         & core-S\'ersic & $58.8$  &  $6.3$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n3842}
% \end{table}  
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 3998 \\}
%
% NGC 3998 is a barred lenticular galaxy with a Seyfert AGN and dust in its nucleus \citep{knapp1996n3998}.
% Despite its large stellar velocity dispersion, this galaxy does not have a partially depleted core TBD!! ref??. 
% \cite{gutierrez2011} identified a bar at $R \lesssim 8''$, a ring between $30'' \lesssim R \lesssim 50''$, and  
% an antitruncation in the light profile of the disk at $R \sim 122''$.
% \cite{laurikainen2010} found that NGC 3998 features a weak bar at $R < 8''$,
% a bright lens at $R < 15''$ and a weak bump in the surface brightness profile at $R \sim 40''$ 
% %($R_{\rm e} \sim 5''$, $n = 2.0$).
% %\cite{sani2011} decomposed this galaxy into a S\'ersic-bulge + exponential-disk + Gaussian-bar + Gaussian-AGN 
% %($R_{\rm e} = 4''.7$, $n = 1.5$).
% %\cite{beifiori2012} modeled this galaxy with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 5''.7$, $n = 2.3$).
% %\cite{lasker2014data} used a S\'ersic-bulge + exponential-disk (with inner truncation) + PSF-nucleus + 
% %S\'ersic-inner disk + S\'ersic-bar 
% %($R_{\rm e} = 2''.0$, $n = 1.1$).
%
% \emph{1D fit -- }
% We see three distinct peaks in the ellipticity profile.
% The first two peaks occur at $R \sim 7''$ and $R \sim 15''$, and they correspond to a weak bar and to a faint 
% oval component respectively.
% These components are both fit with a Ferrer function.
% The third peak at $R \sim 35''$ coincides with a bump in the surface brightness profile and is produced by a ring.
% The AGN component is modeled with a Gaussian profile.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n3998_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 3998.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 3998.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + F-l + G-n + G-r & $5.8$  &  $1.2$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + F-l + G-n + G-r & $4.8$  &  $1.3$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% L+10         &  & $5.0$  &  $2.0$ \\
% S+11         & S-bul + e-d + G-bar + G-n & $4.7$  &  $1.5$ \\
% B+12         & S-bul + e-d & $5.7$  &  $2.3$ \\
% L+14         & S-bul + trunc. e-d + PSF-n + S-bar + S-id & $2.0$  &  $1.1$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n3998}
% \end{table} 
%
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4026 \\}
%
% NGC 4026 is an edge-on lenticular galaxy with a nuclear stellar disk ($R \lesssim 0''.5$, \citealt{ledo2010}).
%
% %\cite{sani2011} modeled this galaxy with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 11''.4$, $n = 3.5$).
%
% \emph{1D fit -- }
% The unsharp mask reveals the presence of a bar ($R \lesssim 30''$) and of a disky component 
% embedded in the bulge, that is responsible for the peak at $R \sim 5''$ in the B4 profile.
% The last can be also recognized in the velocity map.
% The bar is fit with a Ferrer function and the inner disk with a low-$n$ S\'ersic profile.
% Due to the poor spatial resolution of our data, we were not able to model the nuclear stellar disk.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4026_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4026.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 4026.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + S-id & $3.4$  &  $2.4$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + S-id & $6.3$  &  $2.1$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% S+11         & S-bul + e-d & $11.4$  &  $3.5$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n4026}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4151 \\}
%
% NGC 4151 is a face-on barred spiral galaxy that hosts a Seyfert AGN \citep{veroncettyveron2006} 
% and circumnuclear dust \citep{pott2010}.
%
% \cite{sani2011} fit this galaxy with a S\'ersic-bulge + exponential-disk + Gaussian-AGN model
% ($R_{\rm e} = 5''.4$, $n = 3.5$).
% \cite{gadotti2008} decomposed it into a S\'ersic-bulge + exponential-disk + S\'ersic-bar + PSF-AGN 
% ($R_{\rm e} = 4''.7$, $n = 3.0$).
%
% \emph{1D fit -- }
% embedded disk in the bulge,
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4151_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4151.}
% \end{center}
% \end{figure}
%
%
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4261 \\}
%
% NGC 4261 is an elliptical galaxy with %a radio jet \citep{liuxie1992}, 
% a LINER nucleus \citep{veroncettyveron2006}
% and a dusty nuclear disk \citep{tran2001}.
% The galaxy features a partially depleted core ($R_{\rm break} = 1''.2$, \citealt{rusli2013}).
%
% %\cite{grahamdriver2007} fit NGC 4261 with a single S\'ersic-bulge 
% %($R_{\rm e} = 88''.6$, $n = 7.3$).
% %\cite{sani2011} used a S\'ersic-bulge + exponential-disk + Gaussian-AGN model 
% %($R_{\rm e} = 22''.6$, $n = 4.0$).
% %\cite{vika2012} employed a S\'ersic-bulge + mask-core 
% %($R_{\rm e} = 24''.2$, $n = 3.5$).
% %\cite{beifiori2012} adopted a single S\'ersic-bulge model 
% %($R_{\rm e} = 48''.8$, $n = 4.3$).
% %\cite{rusli2013} modeled this galaxy with a core-S\'ersic profile 
% %($R_{\rm e} = 77'.1$, $n = 6.3$).
% %\cite{lasker2014data} employed a S\'ersic-bulge + mask-core model 
% %($R_{\rm e} = 68'.4$, $n = 6.5$).
%
% \emph{1D fit -- }
% The ellipticity profile has a peak at $R \sim 10''$, unraveling the presence of an embedded component.
% We identify such component with an inner disk, possibly seen face-on, and we model it with a Ferrer function.
% The innermost $6''.1$ are excluded from the fit due to the presence of a partially depleted core.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4261_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4261.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 4261.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c + F-id & $52.6$  &  $4.7$ \\
% T.W. 1D eq.  & S-bul + m-c + F-id & $47.3$  &  $4.3$ \\
% T.W. 2D      & S-bul + m-c + e-id & $50.4$  &  $4.4$ \\
% \hline
% GD07      & S-bul		& $88.6$  &  $7.3$ \\
% S+11      & S-bul + e-d + G-n & $22.6$  &  $4.0$ \\
% V+12      & S-bul + m-c	& $24.2$  &  $3.5$ \\
% B+12      & S-bul		& $48.8$  &  $4.3$ \\
% R+13      & core-S\'ersic	& $77.1$  &  $6.3$ \\
% L+14      & S-bul + m-c	& $68.4$  &  $6.5$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n4261}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4291 \\}
%
% NGC 4291 is an elliptical galaxy with a partially depleted core ($R_{\rm break} = 0''.3$, \citealt{rusli2013}).
%
% %\cite{grahamdriver2007} fit this galaxy with a S\'ersic-bulge 
% %($R_{\rm e} = 14'.8$, $n = 4.0$).
% %\cite{rusli2013} modeled this galaxy with a core-S\'ersic profile 
% %($R_{\rm e} = 15''.3$, $n = 5.6$).
% %\cite{lasker2014data} adopted a S\'ersic-bulge + mask-core model
% %($R_{\rm e} = 21''.3$, $n = 8.6$).
%
% \emph{1D fit -- }
% After excluding from the fit the innermost $6''.1$, because affected by the partially depleted core,
% we found that NGC 4291 can be modeled with a single S\'ersic profile.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4291_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4291.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 4291.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c & $15.0$  &  $4.2$ \\
% T.W. 1D eq.  & S-bul + m-c & $15.4$  &  $5.9$ \\
% T.W. 2D      & S-bul + m-c & $20.8$  &  $7.7$ \\
% \hline
% GD07         & S-bul         & $14.8$  &  $4.0$ \\
% R+13         & core-S\'ersic & $15.3$  &  $5.6$ \\
% L+14         & S-bul + m-c   & $21.3$  &  $8.6$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n4291}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4388 \\}
%
% NGC 4388 is an edge-on spiral galaxy with a Seyfert AGN \citep{veroncettyveron2006} 
% and copious nuclear dust \citep{martini2003}.
%
% \emph{1D fit -- }
% The peaks at $R \sim 30''$ in the ellipticity and B4 profiles signal the presence of a bar.
% The data between $35'' \lesssim R \lesssim 65''$ have been excluded because cannot be accommodated by the Ferrer function 
% used to describe the bar. 
% The AGN component is fit with a Gaussian profile.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4388_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4388.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 4388.}
% \begin{center}
% \begin{tabular}{llccc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + G-n & $4.6$  &  $0.6$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + G-n & $4.2$  &  $1.3$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n4388}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4459 \\}
%
% NGC 4459 is a lenticular galaxy.
% Its disk profile is reported to have an antitruncation at $R \sim 119''$ \citep{gutierrez2011}.
%
% %\cite{laurikainen2010} decomposed this galaxy into a S\'ersic-bulge + exponential-disk + Ferrer-lens 
% %($R_{\rm e} \sim 7'' $, $n = 3.0$).
% %\cite{sani2011} used a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 10''.3$, $n = 2.5$).
% %\cite{vika2012} modeled this galaxy with a S\'ersic-bulge + Moffat-nucleus 
% %($R_{\rm e} = 25''.0$, $n = 3.9$).
% %\cite{beifiori2012} adopted a single S\'ersic-bulge model 
% %($R_{\rm e} = 155''.2$, $n = 7.4$).
%
% \emph{1D fit -- }
% The ellipticity constantly increases across all the observed radial range ($R \lesssim 100''$).
% This is an indication that, going from the galaxy center to the outskirts,
% the disk component becomes increasingly more important than the spheroidal component.
% However, the lack of a plateau in the ellipticity profile at large radii implies that at $R \sim 100''$ 
% the contribution from the bulge is still significant compared to that of the disk.
% In fact, the surface brightness profile does not display an exponential decline at large radii.
% We note that the antitruncation reported by \cite{gutierrez2011} could be an artificial feature 
% produced by the transition between the radial range where the disk becomes increasingly more important 
% over the bulge and the truly exponential decline of the disk beyond $R \gtrsim 119''$.
% We do not find evidence for any embedded components in our data.
% A nuclear light excess is observed in the 1D residuals from a S\'ersic + exponential fit 
% and is consequently modeled with a Gaussian profile.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4459_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4459.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 4459.}
% \begin{center}
% \begin{tabular}{llccc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + G-n & $18.4$  &  $3.1$ \\
% T.W. 1D eq.  & S-bul + e-d + G-n & $13.0$  &  $2.6$ \\
% %T.W. 2D	&  & $$  &  $$ \\
% \hline
% L+10      & S-bul + e-d + F-l & $7.0$   &  $3.0$ \\
% S+11      & S-bul + e-d	& $10.3$  &  $2.5$ \\
% V+12      & S-bul + M-n	& $25.0$  &  $3.9$ \\
% B+12      & S-bul		& $155.2$ &  $7.4$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n4459}
% \end{table}
%
%
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4473 \\}
%
% NGC 4473 is an elliptical galaxy with an embedded disk \citep{scorzabender1998,ledo2010}.
%
% \cite{grahamdriver2007} modeled this galaxy with a S\'ersic-bulge 
% ($R_{\rm e} = 39''.6$, $n = 2.7$). 
% \cite{sani2011} used a S\'ersic-bulge model
% ($R_{\rm e} = 49''.3$, $n = 7.0$). 
% \cite{vika2012} fit this galaxy with a S\'ersic-bulge + mask-core 
% ($R_{\rm e} = 21''.3$, $n = 4.3$). 
% \cite{beifiori2012} adopted a S\'ersic-bulge + exponential-disk model 
% ($R_{\rm e} = 10''.6$, $n = 2.2$).  
% \cite{lasker2014data} described this galaxy with a single S\'ersic-bulge 
% ($R_{\rm e} = 27''.9$, $n = 5.1$).  
%
%
% \emph{1D fit -- }
% TBD!! explain ellipticity profile...
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4473_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4473.}
% \end{center}
% \end{figure}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4564 \\}
%
% NGC 4564 is an edge-on lenticular galaxy.
%
% %\cite{grahamdriver2007} decomposed this galaxy into a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 4''.3$, $n = 3.2$). 
% \cite{fisherdrory2008} fit this galaxy with a S\'ersic-bulge + exponential-disk 
% ($R_{\rm e} = 8''.7$, $n = 3.7$). 
% %\cite{sani2011} modeled this galaxy with a S\'ersic-bulge + exponential-disk 
% %($R_{\rm e} = 25''.0$, $n = 7.0$).
% %\cite{vika2012} used a S\'ersic-bulge + exponential-disk model 
% %($R_{\rm e} = 3''.0$, $n = 3.7$). 
%
% \emph{1D fit -- }
% The isophotal parameters do not show any evidence for embedded components.
% However, fitting NGC 4564 with a S\'ersic + exponential model results in bad residuals.
% In the unsharp mask, one can hardly glimpse an oval structure extending out to $R \lesssim 15''$.
% The addition of a Ferrer function to the model dramatically improves the fit and smoothes the residuals.
% Although the lack of strong evidence for the presence of a (faint) third component, 
% we prefer to model the lens with a Ferrer function 
% because of the notable improvement in the goodness of the global fit.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4564_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4564.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 4564.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-l & $5.0$  &  $2.6$ \\
% T.W. 1D eq.  & S-bul + e-d + F-l & $6.0$  &  $3.0$ \\
% %T.W. 2D	&  & $$  &  $$ \\
% \hline
% GD07      & S-bul + e-d & $4.3$  &  $3.2$ \\
% S+11      & S-bul + e-d & $25.0$ &  $7.0$ \\
% V+12      & S-bul + e-d & $3.0$  &  $3.7$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n4564}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4596 \\}
%
% NGC 4596 is an edge-on barred lenticular galaxy.
%
% %\cite{laurikainen2005} decomposed this galaxy into a S\'ersic-bulge + exponential-disk + Ferrer-bar + Ferrer-lens 
% %($R_{\rm e} = 2''.8$, $n = 1.4$). 
% %\cite{sani2011} modeled this galaxy with a S\'ersic-bulge + exponential-disk + Gaussian-bar 
% %($R_{\rm e} = 28''.0$, $n = 3.0$). 
% %\cite{vika2012} decomposed it into a S\'ersic-bulge + exponential-disk + S\'ersic-bar 
% %($R_{\rm e} = 13''.2$, $n = 3.6$). 
% %\cite{beifiori2012} adopted a S\'ersic-bulge + exponential-disk model
% %($R_{\rm e} = 44''.9$, $n = 4.4$).
%
% \emph{1D fit -- }
% The morphology of this galaxy is quite complex. 
% The bar extends out to $R \lesssim 50''$ and concludes in two evident ansae.
% The large-scale disk features a wide ring that is responsible for the curvature in the light profile observed within 
% $60'' \lesssim R \lesssim 130''$ (see also \citealt{comeron2014}).
% An additional embedded disk-like component ($R \lesssim 15''$) can be recognized in the B4 profile and in the velocity map.
% The bump in the light profile within $60'' \lesssim R \lesssim 70''$ corresponds to the ansae of the bar.
% This data range has therefore been excluded from the fit.
% The large scale bar is fit with a Ferrer function and the inner disk with a S\'ersic profile.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4596_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4596.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 4596.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + G-r + S-id & $6.6$  &  $2.7$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + G-r + S-id & $9.0$  &  $3.0$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% L+05      & S-bul + e-d + F-bar + F-l & $2.8$  &  $1.4$ \\
% S+11      & S-bul + e-d + G-bar & $28.0$  &  $3.0$ \\
% V+12      & S-bul + e-d + S-bar & $13.2$  &  $3.6$ \\
% B+12      & S-bul + e-d & $44.9$  &  $4.4$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n4596}
% \end{table}
%
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4697 \\}
%
% NGC 4697 is an elliptical galaxy with an embedded disk \cite{scorzabender1995}.
%
% %\cite{grahamdriver2007} modeled this galaxy with a S\'ersic-bulge model
% %($n = 4.0$).
% %\cite{sani2011} described it with a S\'ersic-bulge 
% %($R_{\rm e} = 100'.5$, $n = 5.0$). 
% %\cite{vika2012} proposed a single S\'ersic-bulge model 
% %($R_{\rm e} = 39''.1$, $n = 3.8$) 
% %or a S\'ersic-bulge + exponential-disk model 
% %($R_{\rm e} = 10''.0$, $n = 2.9$).
% %\cite{beifiori2012} chose a single S\'ersic-bulge 
% %($R_{\rm e} = 128''.7$, $n = 5.0$).
% %\cite{lasker2014data} decomposed NGC 4697 into a S\'ersic-bulge + exponential-disk + PSF-nucleus + S\'ersic-envelope 
% %($R_{\rm e} = 6''.3$, $n = 2.1$).
%
% \emph{1D fit -- }
% The velocity map and the unsharp mask of NGC 4697 clearly show 
% the presence of an intermediate-size disk embedded in the main spheroid.
% However, the ellipticity profile presents two peaks. 
% The peak at $R \sim 40''$ corresponds to the embedded disk already mentioned,
% while the peak at $R \sim 10''$ unveils a second inner disk-like component.
% After testing different decomposition models, 
% in which we fit the embedded disks with a set of different functions (exponential, S\'ersic, Ferrer)
% but we always described the main spheroidal component with a S\'ersic profile,
% we noticed that the bulge parameters did not vary significantly among the various decompositions.
% Our preferred model for NGC 4697 consists of a S\'ersic-bulge + 2 S\'ersic-inner disks.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4697_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4697.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 4697.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + 2 S-id & $239.3$  &  $7.2$ \\
% T.W. 1D eq.  & S-bul + 2 S-id & $226.4$  &  $6.7$ \\
% T.W. 2D      & S-bul + e-id	& $121.4$  &  $5.0$ \\
% \hline
% GD07  & S-bul & $-$  &  $4.0$ \\
% S+11  & S-bul & $100.5$  &  $5.0$ \\
% V+12  & S-bul & $39.1$  &  $3.8$ \\
% V+12  & S-bul + e-d & $10.0$  &  $2.9$ \\
% L+14  & S-bul + e-d + PSF-n + S-halo & $6.3$  &  $2.1$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n4697}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4889 \\}
%
% NGC 4889 is the brightest member of the Coma cluster. 
% It is an elliptical galaxy with a partially depleted core ($R_{\rm break} = 1''.6$, \citealt{rusli2013}).
%
% %\cite{rusli2013} modeled NGC 4889 with a core-S\'ersic model 
% %($R_{\rm e} = 169''.2$, $n = 9.8$). 
%
% \emph{1D fit -- }
% We excluded from the fit the innermost $6''.1$ because affected by the partially depleted core and
% we modeled the galaxy with a single S\'ersic profile.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4889_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4889.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 4889.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c & $119.7$  &  $8.1$ \\
% T.W. 1D eq.  & S-bul + m-c & $60.8$	&  $6.8$ \\
% T.W. 2D      & S-bul + m-c & $104.3$  &  $7.8$ \\
% \hline
% R+13         & core-S\'ersic & $169.2$  &  $9.8$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n4889}
% \end{table}
%   
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 4945 \\}
%
% TBD!!! \\
% NGC 4945 is an edge-on, dusty spiral galaxy that hosts a Seyfert AGN \citep{lin2011}.
%
% \cite{kormendy2010} found that the bulge of NGC 4945 is best modeled with a low-$n$ S\'ersic function
% ($n = 1.3$).
%
% \emph{1D fit -- }
% Given the edge-on inclination of this object, the identification of a possible large-scale bar is complicated.
% The light profile has an obvious bump that could be interpreted in two different ways.
% The bump could be produced by a large-scale bar; 
% in this case, either the disk must have a truncation happening at a radius smaller than the length of the bar,
% or its light profile cannot be exponential (this condition can be caused by the high inclination of the galaxy).
% Alternatively, the bump could be due by the galaxy spiral arms, the disk has an exponential profile with no truncation
% and the galaxy has no bar.
% After checking that both scenarios result in similar best-fit parameters for the bulge, 
% we choose the second and simpler hypothesis.
% We exclude form the fit the data between $R \sim 100''-400''$, where the spiral arm would contribute to the galaxy light,
% and we also exclude the innermost $6''.4$ due to the presence of the AGN.
% Our final model for NGC 4945 consists of a S\'ersic-bulge + exponential-disk.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4945_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 4945.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 4945.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d & $13.9$  &  $1.4$ \\
% T.W. 1D eq.  & S-bul + e-d & $9.5$   &  $1.7$ \\
% T.W. 2D      & S-bul + e-d & $16.2$  &  $0.8$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n4945}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 5077 \\}
%
% ok
%
% \emph{1D fit -- }
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n5077_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 5077.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 5077.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c & $23.5$  &  $4.2$ \\
% T.W. 1D eq.  & S-bul + m-c & $23.0$  &  $5.7$ \\
% T.W. 2D      & S-bul + m-c & $30.5$  &  $6.8$ \\
% \hline
% S+11         & S-bul + G-n & $29.2$  &  $6.0$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n5077}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
%
% {\bf NGC 5128 - Centaurus A \\}
%
% ok
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n5128_1Dfit.eps}
% \caption{}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 5128.}
% \begin{center}
% \begin{tabular}{llccc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & TBD!! & $61.3$  &  $1.2$ \\
% T.W. 1D eq.  &  & $60.8$  &  $2.2$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% S+11         & S-bul + e-d + G-n & $103.6$  &  $3.5$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n5128}
% \end{table}
%
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
%
% {\bf NGC 5576 \\}
%
% ok
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n4889_1Dfit.eps}
% \caption{}
% \end{center}
% \end{figure}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 5845 \\}
%
% NGC 5845 is an elliptical/lenticular galaxy. 
%
% %\cite{grahamdriver2007} modeled this galaxy with a S\'ersic-bulge 
% %($R_{\rm e} = 4''.1$,$n = 3.2$).
% %\cite{sani2011} fit it with a S\'ersic-bulge 
% %($R_{\rm e} = 3''.7$, $n = 3.0$). 
% %\cite{vika2012} used a S\'ersic-bulge model 
% %($R_{\rm e} = 3''.5$, $n = 2.6$).
% %\cite{beifiori2012} chose a single S\'ersic-bulge 
% %($R_{\rm e} = 4''.1$, $n = 3.5$).
% %\cite{lasker2014data} described this galaxy with a S\'ersic-bulge 
% %($R_{\rm e} = 3''.5$, $n = 2.8$).
%
% \emph{1D fit -- }
% The small apparent size of NGC 5845 (compared to the PSF FWHM) implies low resolution for the data 
% and complications for the modeling of this galaxy.
% Here, looking at the unsharp mask and the velocity field, we identify 
% a disk component, either large-scale or embedded.
% To rule out one of the two cases, a good test would be to check whether
% the light profile of NGC 5845 presents an exponential decline in its outer part.
% Unfortunately, the IRAC image is not deep enough and beyond $R \gtrsim 23''$ the galaxy light 
% is fainter than $3 RMS_{\rm SKY}$.
% %Since the too small radial range prevents us from disentangling between an embedded and a large-scale disk,
% %we choose the most statistically significant solution (lowest $\chi^2$), 
% %but we caution that the best fit parameters might be biased by a non adequate choice
% %of the model components. 
% %However, our results are in agreement with those from \cite{jia2012}, 
% %who study the structure and kinematics of NGC5845 with HST/WFPC2 data 
% %and previously published spatially resolved kinematics, and unambiguously identify a 
% %large scale disk in both the photometry and the kinematics. \\
% The ellipticity profile has a peak at $R \sim 13''$ and therefore suggests that the disk is indeed embedded.
% However, when fitting NGC 5845 with a S\'ersic + exponential model, 
% the lowest-$\chi^2$ solution does not reproduce an inner disk, but rather a large-scale one.
% This is because the S\'ersic and the exponential function are degenerate in this decomposition, 
% due to the small number of data points sampling the galaxy light profile.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n5845_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 5845.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 5845.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d & $3.6$  &  $2.5$ \\
% T.W. 1D eq.  & S-bul + e-d & $3.1$  &  $2.3$ \\
% T.W. 2D      & S-bul + e-d & $2.8$  &  $2.4$ \\
% \hline
% GD07      & S-bul & $4.1$  &  $3.2$ \\
% S+11      & S-bul & $3.7$  &  $3.0$ \\
% V+12      & S-bul & $3.5$  &  $2.6$ \\
% B+12      & S-bul & $4.1$  &  $3.5$ \\
% L+14      & S-bul & $3.5$  &  $2.8$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n5845}
% \end{table}
%
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 5846 \\}
%
% NGC 5846 is an elliptical galaxy with a partially depleted core ($R_{\rm break} = 1''.2$, \citealt{rusli2013}).
% This galaxy has a LINER nucleus \citep{carrillo1999} and filamentary nuclear dust \citep{tran2001}.
%
% %\cite{sani2011} modeled this galaxy with a S\'ersic-bulge
% %($R_{\rm e} = 36''.4$, $n = 3.0$). 
% %\cite{vika2012} used a S\'ersic-bulge + mask-core model 
% %($R_{\rm e} = 46''.3$, $n = 3.7$). 
% %\cite{rusli2013} fit this galaxy with a core-S\'ersic profile 
% %($R_{\rm e} = 113''.2$, $n = 5.3$). 
%
% \emph{1D fit -- }
% This galaxy has a strong isophotal twist between its center and $R \sim 70''$, and
% its light profile presents a slight bump at $R \sim 20''$.
% However, the isophotal parameters, the unsharp mask and the velocity map lack of clear evidence for an embedded component.
% We thus model NGC 5846 with a single S\'ersic profile, after masking the innermost $6''.1$.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n5846_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 5846.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 5846.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c & $105.1$  &  $6.4$ \\
% T.W. 1D eq.  & S-bul + m-c & $83.4$	&  $5.7$ \\
% T.W. 2D      & S-bul + m-c & $85.1$	&  $5.2$ \\
% \hline
% S+11      & S-bul	    & $36.4$   &  $3.0$ \\
% V+12      & S-bul + m-c   & $46.3$   &  $3.7$ \\
% R+13      & core-S\'ersic & $113.2$  &  $5.3$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n5846}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 6251 \\}
%
% NGC 6251 is an elliptical galaxy with a large stellar velocity dispersion ($\sigma \sim 310~\rm km~s^{-1}$),
% which is a clue for the presence of a partially depleted core.
% The galaxy features a nuclear disk of dust \citep{ferrareseford1999n6251} and a Seyfert AGN \citep{panessabassani2002}.
%
% %\cite{grahamdriver2007} fit this galaxy with a S\'ersic-bulge 
% %($R_{\rm e} = 173''.92$, $n = 11.8$). 
% %\cite{sani2011} adopted a S\'ersic-bulge + Gaussian-AGN model 
% %($R_{\rm e} = 42''.4$, $n = 7.0$). 
% %\cite{lasker2014data} modeled this galaxy with a S\'ersic-bulge 
% %($R_{\rm e} = 20''.6$, $n = 5.0$). 
%
% \emph{1D fit -- }
% We masked the innermost $6''.1$ because possibly affected by the combined contribution of the AGN 
% and the partially depleted core. 
% A single S\'ersic profile provides a good description of NGC 6251.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n6251_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 6251.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 6251.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c & $41.7$  &  $6.8$ \\
% T.W. 1D eq.  & S-bul + m-c & $30.1$  &  $5.6$ \\
% T.W. 2D      & S-bul + m-c & $39.3$  &  $7.1$ \\
% \hline
% GD07      & S-bul	     & $173.9$  &  $11.8$ \\
% S+11      & S-bul + G-n    & $42.4$	&  $7.0$ \\
% L+14      & S-bul	     & $20.6$	&  $5.0$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n6251}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 7052 \\}
%
% NGC 7052 is an elliptical galaxy reported to have a partially depleted core 
% ($R_{\rm break} = 0''.8$, \citealt{quillen2000}). 
%
% %\cite{grahamdriver2007} modeled this galaxy with a S\'ersic-bulge 
% %($R_{\rm e} = 70''.4$, $n = 4.6$). 
% %\cite{sani2011} used a S\'ersic-bulge model 
% %($R_{\rm e} = 39''.3$, $n = 5.0$). 
% %\cite{vika2012} fit this galaxy with a S\'ersic-bulge + exponential-disk
% %($R_{\rm e} = 4''.3$, $n = 1.8$).
% %\cite{lasker2014data} adopted a S\'ersic-bulge model 
% %($R_{\rm e} = 26''.6$, $n = 4.2$). 
%
% \emph{1D fit -- }
% A single S\'ersic function provides a fairly good description of the surface brightness profile of NGC 7052.
% However, the residuals of such fit display a curious behaviour, having a nuclear light excess along the major-axis 
% and a nuclear light deficit along the equivalent-axis. 
% This pattern is still present when the innermost $6''.1$ are excluded and the fit is repeated with a S\'ersic profile.
% The reasons for this are unclear. 
% The best-fit parameters obtained from a single S\'ersic fit, without excluding the innermost data points,
% are not significantly different from those obtained when excluding the core region (see Table \ref{tab:n7052}).
% Given the uncertainty connected to the nuclear region of this galaxy,
% we preferred the model that incorporates the exclusion of the innermost $6''.1$, 
% with the caveat about the ambiguous presence of a partially depleted core.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n7052_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 7052.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 7052.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% \multicolumn{4}{c}{\emph{excluding innermost $6''.1$}} \\
% T.W. 1D maj. & S-bul + m-c & $59.4$  &  $4.2$ \\
% T.W. 1D eq.  & S-bul + m-c & $37.0$  &  $5.6$ \\
% T.W. 2D      & S-bul + m-c & $36.2$  &  $4.0$ \\
% \hline
% \multicolumn{4}{c}{\emph{including innermost $6''.1$}} \\
% T.W. 1D maj. & S-bul & $66.1$  &  $4.8$ \\
% T.W. 1D eq.  & S-bul & $35.0$  &  $5.0$ \\
% %T.W. 2D	&  & $$  & $$  &  $$ \\
% \hline
% GD07      & S-bul & $70.4$  &  $4.6$ \\
% S+11      & S-bul & $39.3$  &  $5.0$ \\
% V+12      & S-bul + e-d & $4.3$  &  $1.8$ \\
% L+14      & S-bul & $26.6$  &  $4.2$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n7052}
% \end{table}
%
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 7619 \\}
%
% NGC 7619 is an elliptical galaxy reported to have a partially depleted core ($R_{\rm break} = 0''.5$, \citealt{rusli2013}).
%
% %\cite{rusli2013} fit this galaxy with a core-S\'ersic profile 
% %($R_{\rm e} = 100''.1$, $n = 9.3$). 
%
% \emph{1D fit -- }
% We identified an embedded disk signaled by the peak at $R \sim 10''$ in the ellipticity profile.
% The velocity map of this galaxy confirms the presence of a fast rotating component (Falcon-Barroso private comm.).
% We modeled NGC 7619 with a S\'ersic + inner-exponential model.
% We note that the residuals obtained from our best-fit model do not suggest the presence of a partially depleted core,
% whereas they display a nuclear lack of light when the galaxy is modeled with a single S\'ersic-bulge. 
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n7619_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 7619.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 7619.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-id & $63.2$  &  $5.3$ \\
% T.W. 1D eq.  & S-bul + e-id & $58.0$  &  $5.2$ \\
% \hline
% R+13      & core-S\'ersic & $100.1$  &  $9.3$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n7619}
% \end{table}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf NGC 7768 \\}
%
% NGC 7768 is an elliptical galaxy with a partially depleted core ($R_{\rm break} = 0''.3$, \citealt{rusli2013}).
%
% %\cite{rusli2013} modeled this galaxy with a core-S\'ersic profile 
% %($R_{\rm e} = 46''.1$, $n = 6.2$). 
%
% \emph{1D fit -- }
% The image of NGC 7768 is corrupted by a saturated star, which lies close to the galaxy.
% The sky background is not constant across the image.
% To be safe, we fit only the innermost $R \lesssim 30''$, where the contribution from the background is negligible.
% The data within $R < 6''.1$ are excluded from the fit due to the presence of a partially depleted core.
% Our best-fit model consists of a single S\'ersic profile.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/n7768_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for NGC 7768.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for NGC 7768.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + m-c & $92.9$  &  $8.4$ \\
% T.W. 1D eq.  & S-bul + m-c & $42.1$  &  $6.7$ \\
% \hline
% R+13      & core-S\'ersic & $46.1$  &  $6.2$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:n7768}
% \end{table}
%  
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage\newpage\noindent
% {\bf UGC 03789 \\}
%
% UGC 03789 is a face-on spiral galaxy.
%
% \emph{1D fit -- }
% This galaxy features a ring ($R \sim 20''$) and a nuclear bar ($R \lesssim 4''$),
% which can be seen in the unsharp mask and produces corresponding peaks in the ellipticity and PA profiles.
% The bar is fit with a Ferrer function.
%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\fitfigurewidth]{images/ugc03789_1Dfit.eps}
% \caption{Best-fit model and isophotal parameters for UGC 03789.}
% \end{center}
% \end{figure}
%
% \begin{table}[h]
% \small
% \caption{Best-fit parameters for UGC 03789.}
% \begin{center}
% \begin{tabular}{llcc}
% \hline
% {\bf Work} & {\bf Model}   & $\bm R_{\rm e}$    & $\bm n$ \\
%   &  &  $[\rm arcsec]$ & \\
% \hline
% T.W. 1D maj. & S-bul + e-d + F-bar + G-r & $1.8$  &  $1.9$ \\
% T.W. 1D eq.  & S-bul + e-d + F-bar + G-r & $2.4$  &  $1.4$ \\
% \hline
% \end{tabular}
% \end{center}
% \label{tab:ugc3789}
% \end{table}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgments
GS warmly thanks Chieng Peng, Peter Erwin, Luca Cortese, Elisabete Lima Da Cunha and Gonzalo Diaz 
for useful discussion. \\
This research was supported by Australian Research Council funding through grants
DP110103509 and FT110100263.
This work is based on observations made with the IRAC instrument \citep{fazio2004IRAC} 
on-board the Spitzer Space Telescope, 
which is operated by the Jet Propulsion Laboratory, 
California Institute of Technology under a contract with NASA.
This research has made use of the GOLDMine database \citep{goldmine} and the NASA/IPAC Extragalactic Database (NED) 
which is operated by the Jet Propulsion Laboratory, California Institute of Technology, 
under contract with the National Aeronautics and Space Administration.

\appendix

\section{1D analytical functions}
\label{sec:app1}
Below we give the mathematical expressions of all the analytical functions used in this work to model 
the observed surface brightness profiles ($\mu$) of galaxies.
We indicate with $R$ the projected galactic radius, 
corresponding to the distance of the isophotes from the galaxy center (along the major- or equivalent-axis). \\

The \emph{S\'ersic profile} \citep{sersic1963,sersic1968} is a three-parameter function with the following form:
\begin{equation}
\mu_{\rm S\acute{e}rsic}(\mu_{\rm e},R_{\rm e},n;R) = \mu_{\rm e} + \frac{2.5~b_{\rm n}}{\ln(10)} 
\Biggl[\biggl(\frac{R}{R_{\rm e}}\biggr)^{1/n}-1\Biggr] ,
\end{equation}
\citep{caon1993,andredakis1995,grahamdriver2005}
where $\mu_{\rm e}$ is the surface brightness at the effective radius $R_{\rm e}$ that encloses 
half of the total light from the model, 
the S\'ersic index $n$ is the parameter that measures the curvature of the radial light profile,
and $b_{\rm n}$ is a scalar value defined in terms of the S\'ersic index $n$.
The value of $b_{\rm n}$ is found by solving numerically the following expression:
\begin{equation}
\Gamma(2n) = 2\gamma(2n,b_{\rm n}),
\end{equation}
where $\Gamma$ is the complete gamma function \citep{ciotti1991} 
and $\gamma$ is the incomplete gamma function defined by 
\begin{equation}
\gamma(2n,x) = \int_{0}^{x} {\rm e}^{-t} t^{2n-1} {\rm d}t .
\end{equation} \\

The \emph{exponential profile} is a special case ($n=1$) of the S\'ersic profile.
It is therefore a two-parameter function, with the following form:
\begin{equation}
\mu_{\rm exponential}(\mu_{\rm 0},h;R) = \mu_{\rm 0} + \frac{2.5}{\ln(10)} \biggl(\frac{R}{h} \biggr) ,
\end{equation}
where $\mu_{\rm 0}$ is the central surface brightness and $h$ is the scale length. \\

The \emph{Gaussian profile} is a two-parameter function with the following form:
\begin{equation}
\mu_{\rm Gaussian}(\mu_{\rm 0},FWHM;R) = \mu_{\rm 0} + \frac{2.5}{\ln(10)} 
\biggl[\frac{R^2}{2(FWHM/2.355)^2} \biggr] ,
\end{equation}
where $\mu_{\rm 0}$ is the central surface brightness 
and $FWHM$ is the full width at half maximum of the Gaussian profile. \\

The \emph{Moffat profile} \citep{moffat1969} is a three-parameter function with the following form:
\begin{equation}
\mu_{\rm Moffat}(\mu_{\rm 0},\alpha,\beta;R) = \mu_{\rm 0} - 2.5 
\log \Biggl[ 1 + \biggl( \frac{R}{\alpha} \biggr)^2 \Biggr]^{-\beta} ,
\end{equation}
where $\mu_{\rm 0}$ is the central surface brightness,
$\alpha$ is related to the $FWHM$ through 
\begin{equation}
FWHM = 2\alpha \sqrt{2^{1/\beta}-1}
\end{equation}
and $\beta$ regulates the shape of the profile at large radii. \\

The \emph{Ferrer profile} is a four-parameter function with the following form:
\begin{equation}
\mu_{\rm Ferrer}(\mu_{\rm 0},R_{\rm out},\alpha,\beta;R) = \left\{
  \begin{array}{l l}
    \mu_{\rm 0} - 2.5 \log \Biggl[ 1 - \biggl(\frac{R}{R_{\rm out}} \biggr)^{2-\beta} \Biggr]^{\alpha} 
    & \quad \text{for $R<R_{\rm out}$}\\
    +\infty & \quad \text{for $R\geq R_{\rm out}$}
  \end{array} \right. ,
\end{equation}
where $\mu_{\rm 0}$ is the central surface brightness,
$\alpha$ controls the sharpness of the truncation,
$\beta$ is related to the central slope,
and $R_{\rm out}$ is the limit within which the function is defined. \\

The \emph{symmetric Gaussian ring} is a three-parameter function with the following form:
\begin{equation}
\mu_{\rm Gaussian}(\mu_{\rm 0},R_{\rm 0},FWHM;R) = \mu_{\rm 0} + \frac{2.5}{\ln(10)} 
\biggl[\frac{(R-R_{\rm 0})^2}{2(FWHM/2.355)^2} \biggr] ,
\end{equation}
where $\mu_{\rm 0}$ is the central surface brightness,
$R_{\rm 0}$ is the radius at which the Gaussian profile is centered, 
and $FWHM$ is the full width at half maximum of the Gaussian profile. \\

\bibliography{/Users/gsavorgnan/Dropbox/giulia_e_basta/Literature/SMBHbibliography}

\clearpage

\end{document}