### 2015-02-10 papers/data_paper/table_sample.tex

\begin{table*}                                        
\begin{center}                                        
\caption{{\bf Galaxy sample.}                        
\emph{Column (1):} Galaxy name.                       
\emph{Column (2):} Distance.                                   
\emph{Column (3):} Black hole mass.                                   
\emph{Column (4):} Reference of the black hole mass reported here (G+03 = \citealt{greenhill2003}, GS13 = \citealt{grahamscott2013}; R+13b = \citealt{rusli2013bhmassesDM}).                                   
\emph{Column (5):} Presence of a partially depleted core. 
The question mark is used when the classification has come from the velocity dispersion criteria mentioned in Section \ref{sec:corser}. 
The value of the core break radius is reported in parenthesis when available.  
\emph{Column (6):} Reference of the identification of a partially depleted core (G+94 = \citealt{grillmair1994}; F+97 = \citealt{forbes1997}; Q+00 = \citealt{quillen2000}, 
T+04 = \citealt{trujillo2004coresersicmodel}; F+06 = \citealt{ferrarese2006acsvcs}; J+11 = \citealt{jardel2011}; R+11 = \citealt{richings2011}; 
DG13 = \citealt{dullograham2013cores}; R+13a = \citealt{rusli2013}).  
\emph{Column (7):} Kinematical classification (fast/slow rotator).
\emph{Column (8):} Availability of velocity map (A = ATLAS$^{\rm 3D}$, S = SLUGGS). 
\emph{Column (9):} Completion of 1D fit. 
\emph{Column (10):} Completion of 2D fit. }                                 
\begin{tabular}{llllllllll}                           
\hline                                                
\multicolumn{1}{l}{{\bf Galaxy}} &                   
\multicolumn{1}{l}{{\bf Distance}} &                 
\multicolumn{1}{l}{{\bf $\bm{M_{\rm BH}}$}} &  
\multicolumn{1}{l}{{\bf Ref.}} &                     
\multicolumn{1}{l}{{\bf Core}} &                     
\multicolumn{1}{l}{{\bf Ref.}} &                     
\multicolumn{1}{l}{{\bf Rot.}} &                     
\multicolumn{1}{l}{{\bf Vel. map}} &                 
\multicolumn{1}{l}{{\bf 1D fit}} &                   
\multicolumn{1}{l}{{\bf 2D fit}} \\                
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{[Mpc]} &                           
\multicolumn{1}{l}{$[10^8~\rm M_{\odot}]$} &         
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{$([\rm arcsec])$} &                                
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{} \\                             
\multicolumn{1}{l}{(1)} &                             
\multicolumn{1}{l}{(2)} &                             
\multicolumn{1}{l}{(3)} &                             
\multicolumn{1}{l}{(4)} &                             
\multicolumn{1}{l}{(5)} &                             
\multicolumn{1}{l}{(6)} &                             
\multicolumn{1}{l}{(7)} &                             
\multicolumn{1}{l}{(8)} &                             
\multicolumn{1}{l}{(9)} &                             
\multicolumn{1}{l}{(10)} \\                         
\hline                                                
Circinus   &  $4.0$  &  $0.017_{-0.003}^{+0.004}$   &  G+03  &  no?  &     &      &     &  no  &  no  \\ 
IC 1459  &  $28.4$  &  $24_{-10}^{+10}$   &  GS13  &  yes  $(0.7)$  &  R+13a  &      &     &  yes  &  yes  \\ 
IC 2560  &  $40.7$  &  $0.044_{-0.022}^{+0.044}$   &  GS13  &  no?  &     &      &     &  yes  &  no  \\ 
IC 4296  &  $40.7$  &  $11_{-2}^{+2}$   &  GS13  &  yes?  &     &      &     &  yes  &  yes  \\ 
M104  &  $9.5$  &  $6.4_{-0.4}^{+0.4}$   &  GS13  &  yes   &  J+11  &      &     &  yes  &  no  \\ 
M105  &  $10.3$  &  $4_{-1}^{+1}$   &  GS13  &  yes  $(1.1)$  &  DG13, R+13a  &  FAST   &  A  &  yes  &  yes  \\ 
M106  &  $7.2$  &  $0.39_{-0.01}^{+0.01}$   &  GS13  &  no   &     &      &     &  yes  &  no  \\ 
M31  &  $0.7$  &  $1.4_{-0.3}^{+0.9}$   &  GS13  &  no   &     &      &     &  yes  &  no  \\ 
M32  &  $0.8$  &  $0.024_{-0.005}^{+0.005}$   &  GS13  &  no   &     &      &     &  no  &  no  \\ 
M49  &  $17.1$  &  $25_{-1}^{+3}$   &  R+13b  &  yes  $(1.5)$  &  DG13, R+13a  &   SLOW  &  A  &  yes  &  yes  \\ 
M59  &  $17.8$  &  $3.9_{-0.4}^{+0.4}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  no  \\ 
M60  &  $16.4$  &  $47_{-10}^{+10}$   &  GS13  &  yes  $(2.7)$  &  DG13, R+13a  &  FAST   &  A, S  &  no  &  no  \\ 
M64  &  $7.3$  &  $0.016_{-0.004}^{+0.004}$   &  GS13  &  no?  &     &      &     &  yes  &  no  \\ 
M77  &  $15.2$  &  $0.084_{-0.003}^{+0.003}$   &  GS13  &  no   &     &      &     &  no  &  no  \\ 
M81  &  $3.8$  &  $0.74_{-0.11}^{+0.21}$   &  GS13  &  no   &     &      &     &  yes  &  no  \\ 
M84  &  $17.9$  &  $9.0_{-0.8}^{+0.9}$   &  GS13  &  yes  $(1.9)$  &  F+06  &   SLOW  &  A, S  &  yes  &  yes  \\ 
M87  &  $15.6$  &  $58.0_{-3.5}^{+3.5}$   &  GS13  &  yes  $(7.2)$  &  F+06  &   SLOW  &  A, S  &  yes  &  yes  \\ 
M89  &  $14.9$  &  $4.7_{-0.5}^{+0.5}$   &  GS13  &  yes  $(0.4)$  &  DG13, R+13a  &   SLOW  &  A  &  yes  &  no  \\ 
M94  &  $4.4$  &  $0.060_{-0.014}^{+0.014}$   &  GS13  &  no?  &     &      &     &  yes  &  no  \\ 
M96  &  $10.1$  &  $0.073_{-0.015}^{+0.015}$   &  GS13  &  no   &     &      &     &  yes  &  yes  \\ 
NGC 0253  &  $3.5$  &  $0.10_{-0.05}^{+0.10}$   &  GS13  &  no   &     &      &     &  no  &  no  \\ 
NGC 0524  &  $23.3$  &  $8.3_{-1.3}^{+2.7}$   &  GS13  &  yes  $(0.2)$  &  R+11  &  FAST   &  A  &  yes  &  no  \\ 
NGC 0821  &  $23.4$  &  $0.39_{-0.09}^{+0.26}$   &  GS13  &  no   &     &  FAST   &  A, S  &  yes  &  yes  \\ 
NGC 1023  &  $11.1$  &  $0.42_{-0.04}^{+0.04}$   &  GS13  &  no   &     &  FAST   &  A, S  &  yes  &  yes  \\ 
NGC 1300  &  $20.7$  &  $0.73_{-0.35}^{+0.69}$   &  GS13  &  no   &     &      &     &  yes  &  no  \\ 
NGC 1316  &  $18.6$  &  $1.50_{-0.80}^{+0.75}$   &  GS13  &  no   &     &  FAST   &     &  yes  &  no  \\ 
NGC 1332  &  $22.3$  &  $14_{-2}^{+2}$   &  GS13  &  no   &     &      &     &  yes  &  no  \\ 
NGC 1374  &  $19.2$  &  $5.8_{-0.5}^{+0.5}$   &  R+13b  &  no?  &     &  FAST   &  A  &  yes  &  yes  \\ 
NGC 1399  &  $19.4$  &  $4.7_{-0.6}^{+0.6}$   &  GS13  &  yes  $(2.4)$  &  DG13, R+13a  &   SLOW  &  A  &  yes  &  no  \\ 
NGC 2273  &  $28.5$  &  $0.083_{-0.004}^{+0.004}$   &  GS13  &  no   &     &      &     &  yes  &  no  \\ 
NGC 2549  &  $12.3$  &  $0.14_{-0.13}^{+0.02}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  yes  \\ 
NGC 2778  &  $22.3$  &  $0.15_{-0.10}^{+0.09}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  no  \\ 
NGC 2787  &  $7.3$  &  $0.40_{-0.05}^{+0.04}$   &  GS13  &  no   &     &      &     &  yes  &  no  \\ 
NGC 2974  &  $20.9$  &  $1.7_{-0.2}^{+0.2}$   &  GS13  &  no   &     &  FAST   &  A, S  &  yes  &  yes  \\ 
NGC 3079  &  $20.7$  &  $0.024_{-0.012}^{+0.024}$   &  GS13  &  no?  &     &      &     &  yes  &  no  \\ 
NGC 3091  &  $51.2$  &  $36_{-2}^{+1}$   &  R+13b  &  yes  $(0.6)$  &  R+13a  &      &     &  yes  &  yes  \\ 
NGC 3115  &  $9.4$  &  $8.8_{-2.7}^{+10.0}$   &  GS13  &  no   &     &      &     &  yes  &  no  \\ 
NGC 3227  &  $20.3$  &  $0.14_{-0.06}^{+0.10}$   &  GS13  &  no   &     &      &     &  yes  &  no  \\ 
NGC 3245  &  $20.3$  &  $2.0_{-0.5}^{+0.5}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  yes  \\ 
NGC 3377  &  $10.9$  &  $0.77_{-0.06}^{+0.04}$   &  GS13  &  no   &     &  FAST   &  A, S  &  yes  &  yes  \\ 
NGC 3384  &  $11.3$  &  $0.17_{-0.02}^{+0.01}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  no  \\ 
NGC 3393  &  $55.2$  &  $0.34_{-0.02}^{+0.02}$   &  GS13  &  no   &     &      &     &  yes  &  yes  \\ 
NGC 3414  &  $24.5$  &  $2.4_{-0.3}^{+0.3}$   &  GS13  &  no   &     &   SLOW  &  A  &  yes  &  no  \\ 
NGC 3489  &  $11.7$  &  $0.058_{-0.008}^{+0.008}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  yes  \\ 
NGC 3585  &  $19.5$  &  $3.1_{-0.6}^{+1.4}$   &  GS13  &  no   &     &      &     &  yes  &  no  \\ 
NGC 3607  &  $22.2$  &  $1.3_{-0.5}^{+0.5}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  yes  \\ 
NGC 3608  &  $22.3$  &  $2.0_{-0.6}^{+1.1}$   &  GS13  &  yes  $(0.2)$  &  DG13, R+13a  &   SLOW  &  A, S  &  yes  &  yes  \\ 
NGC 3842  &  $98.4$  &  $97_{-26}^{+30}$   &  GS13  &  yes  $(0.7)$  &  DG13, R+13a  &      &     &  yes  &  no  \\ 
NGC 3998  &  $13.7$  &  $8.1_{-1.9}^{+2.0}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  no  \\ 
NGC 4026  &  $13.2$  &  $1.8_{-0.3}^{+0.6}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  no  \\ 
NGC 4151  &  $20.0$  &  $0.65_{-0.07}^{+0.07}$   &  GS13  &  no   &     &      &     &  yes  &  no  \\ 
\hline         
\end{tabular}   
\label{tab:sample} 
\end{center}    
\end{table*}

\begin{table*}                                        
\begin{center}                                        
\begin{tabular}{llllllllll}                           
\hline                                                
\multicolumn{1}{l}{{\bf Galaxy}} &                   
\multicolumn{1}{l}{{\bf Distance}} &                 
\multicolumn{1}{l}{{\bf $\mathbf{M_{\rm BH}}$}} &  
\multicolumn{1}{l}{{\bf Ref.}} &                     
\multicolumn{1}{l}{{\bf Core}} &                     
\multicolumn{1}{l}{{\bf Ref.}} &                     
\multicolumn{1}{l}{{\bf Rot.}} &                     
\multicolumn{1}{l}{{\bf Vel. map}} &                 
\multicolumn{1}{l}{{\bf 1D fit}} &                   
\multicolumn{1}{l}{{\bf 2D fit}} \\                
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{[Mpc]} &                           
\multicolumn{1}{l}{$[10^8~\rm M_{\odot}]$} &         
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{$([\rm arcsec])$} &                                
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{} &                                
\multicolumn{1}{l}{} \\                             
\multicolumn{1}{l}{(1)} &                             
\multicolumn{1}{l}{(2)} &                             
\multicolumn{1}{l}{(3)} &                             
\multicolumn{1}{l}{(4)} &                             
\multicolumn{1}{l}{(5)} &                             
\multicolumn{1}{l}{(6)} &                             
\multicolumn{1}{l}{(7)} &                             
\multicolumn{1}{l}{(8)} &                             
\multicolumn{1}{l}{(9)} &                             
\multicolumn{1}{l}{(10)} \\                         
\hline                                                
NGC 4261  &  $30.8$  &  $5_{-1}^{+1}$   &  GS13  &  yes  $(1.6)$  &  R+11  &   SLOW  &  A  &  yes  &  yes  \\ 
NGC 4291  &  $25.5$  &  $3.3_{-2.5}^{+0.9}$   &  GS13  &  yes  $(0.3)$  &  DG13, R+13a  &      &     &  yes  &  yes  \\ 
NGC 4342  &  $23.0$  &  $4.5_{-1.5}^{+2.3}$   &  GS13  &  no   &     &  FAST   &  A  &  no  &  no  \\ 
NGC 4388  &  $17.0$  &  $0.075_{-0.002}^{+0.002}$   &  GS13  &  no?  &     &      &     &  yes  &  no  \\ 
NGC 4459  &  $15.7$  &  $0.68_{-0.13}^{+0.13}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  no  \\ 
NGC 4473  &  $15.3$  &  $1.2_{-0.9}^{+0.4}$   &  GS13  &  no   &     &  FAST   &  A, S  &  yes  &  yes  \\ 
NGC 4486A  &  $17.0$  &  $0.13_{-0.08}^{+0.08}$   &  GS13  &  no   &     &  FAST   &  A  &  no  &  no  \\ 
NGC 4564  &  $14.6$  &  $0.60_{-0.09}^{+0.03}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  no  \\ 
NGC 4596  &  $17.0$  &  $0.79_{-0.33}^{+0.38}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  no  \\ 
NGC 4697  &  $11.4$  &  $1.8_{-0.1}^{+0.2}$   &  GS13  &  no   &     &  FAST   &  A, S  &  yes  &  yes  \\ 
NGC 4889  &  $103.2$  &  $210_{-160}^{+160}$   &  GS13  &  yes  $(1.7)$  &  F+97  &      &     &  yes  &  yes  \\ 
NGC 4945  &  $3.8$  &  $0.014_{-0.007}^{+0.014}$   &  GS13  &  no?  &     &      &     &  yes  &  yes  \\ 
NGC 5077  &  $41.2$  &  $7.4_{-3.0}^{+4.7}$   &  GS13  &  yes  $(0.3)$  &  T+04  &      &     &  yes  &  yes  \\ 
NGC 5128  &  $3.8$  &  $0.45_{-0.10}^{+0.17}$   &  GS13  &  no?  &     &      &     &  yes  &  no  \\ 
NGC 5576  &  $24.8$  &  $1.6_{-0.4}^{+0.3}$   &  GS13  &  no   &     &   SLOW  &  A  &  yes  &  yes  \\ 
NGC 5813  &  $31.3$  &  $6.8_{-0.7}^{+0.7}$   &  GS13  &  yes  $(0.4)$  &  DG13, R+13a  &   SLOW  &  A  &  no  &  no  \\ 
NGC 5845  &  $25.2$  &  $2.6_{-1.5}^{+0.4}$   &  GS13  &  no   &     &  FAST   &  A  &  yes  &  yes  \\ 
NGC 5846  &  $24.2$  &  $11_{-1}^{+1}$   &  GS13  &  yes   &  F+97  &   SLOW  &  A, S  &  yes  &  yes  \\ 
NGC 6251  &  $104.6$  &  $5_{-2}^{+2}$   &  GS13  &  yes?  &     &      &     &  yes  &  yes  \\ 
NGC 7052  &  $66.4$  &  $3.7_{-1.5}^{+2.6}$   &  GS13  &  yes  $(0.8)$  &  Q+00  &      &     &  yes  &  yes  \\ 
NGC 7582  &  $22.0$  &  $0.55_{-0.19}^{+0.26}$   &  GS13  &  no   &     &      &     &  no  &  no  \\ 
NGC 7619  &  $51.5$  &  $25_{-3}^{+8}$   &  R+13b  &  yes  $(0.5)$  &  DG13, R+13a  &      &     &  yes  &  no  \\ 
NGC 7768  &  $112.8$  &  $13_{-4}^{+5}$   &  GS13  &  yes   &  G+94  &      &     &  yes  &  no  \\ 
UGC 03789  &  $48.4$  &  $0.108_{-0.005}^{+0.005}$   &  GS13  &  no?  &     &      &     &  yes  &  no  \\ 
\hline         
\end{tabular}   
\end{center}    
\end{table*}