### 2015-10-06 papers/MbhMsph/replytoreferee.txt
kry

-

- The referee is correct in saying that the last point in our Conclusion Section is not a new result and,  
  indeed, we didn't mean to deliver this result as a novelty. The Introductions of Papers I and II present 
  some points about black hole mass scaling relations that are currently under debate. One of these points 
  is the level of intrinsic scatter in the black hole mass - spheroid/galaxy luminosity correlation. Some 
  studies (e.g. Erwin & Gadotti 2012, Beifiori et al. 2012, Kormandy & Ho 2013) claimed that the spheroid 
  luminosity is a better tracer of the black hole mass, but the most recent study by Lasker et al. reported 
  that the spheroid luminosity and the galaxy luminosity are equally good proxies of the black hole mass. 
  Our work solves this controversy by showing that the spheroid luminosity and the galaxy luminosity correlate 
  equally well with the black hole mass when only early-type galaxies are considered (Lasker's sample has only 
  4 spiral galaxies). However, when the galaxy sample contains a reasonable number of spiral galaxies, the 
  correlation with the spheroid luminosity has less intrinsic scatter than that with the galaxy luminosity. 
  To our best knowledge, this has never been shown before in the literature. 
  Additional references have been added to the last point in our Conclusion Section.

- The referee is correct, and we are aware of this. Unlike many papers, we believe that we have not under-
  estimated the uncertainties associated to the bulge best-fit parameters. It is possible that some of the 
  results taken from the literature are "grossly wrong", however the exclusion of extreme outliers would 
  affect only the sigma_3 error bars, i.e. the uncertainties associated to the "grade 3" fits, which constitute 
  only 14% of all fits. The exclusion of extreme outliers would have a very small impact on the sigma_1 and 
  sigma_2 errorbars (we learned this from Savorgnan et al. 2013, where we applied a very similar method). 
  In conclusion, the exclusion of extreme outliers would decrease the uncertainties associated to 14% of 
  all galaxies, with little or no effects on the measurement of the intrinsic scatter.

- We prefer not to reference Burstein et al. because they falsely identified many discs due to poor R^(1/4) 
  decompositions.

- A footnote has been added in Section 4.3 to remind the reader why a slope~1 is consistent with a dry-merging 
  formation scenario, although it simply repeats what already stated in the Introduction.

- A (B-V) galaxy color vs. log(galaxy stellar mass) plot is attached. The early-type galaxies are seen to occupy 
  the region of the red sequence, and the spiral galaxies are located in the region of the blue cloud. 
  The correspondence between galaxy morphological type (early vs. late) and position in the color-mass diagram 
  (red sequence vs. blue cloud) is a very well estabilished result and we don't feel the need of showing this 
  (poor) diagram in the paper.

- We fully agree with the referee on this point, as extensively discussed in Section 4.2.1. We don't think that 
  classical and pseudo-bulges can be distinguished based on their Sersic index. However, this has been common 
  practice in the recent literature, and here we are showing that n<2 bulges are not offset from the correlation
  traced by n>2 bulges. The referee might appreciate that we never refereed to our n<2 bulges merely as pseudo-bulges, 
  but we always specified "claimed by some to be pseudo-bulges". In addition, in the text we have questioned the 
  n<2 criterion used by some to identify pseudo-bulges.

- We never claimed that the break of the black hole - bulge mass correlation is due to the bulge mass. 
  In the Introduction, we anticipated that, for consistency with other known scaling relations (Lsph-sigma and 
  Mbh-sigma), a bent Mbh-Lsph correlation is expected, with the bend occurring around Mbh~10^8. 
  In the Mbh-Lsph and Mbh-Msph diagrams, we identified two different sequences (early vs. late) precisely according 
  to the morphology of galaxies. Indeed, the break is caused by the galaxy morphology. 
  We did not identify pseudo-bulges with n<2 bulges. 
  In Figure 5, the red dashed line marks the correlation for ellipticals+lenticulars (as stated in the caption). 
  The correlation for ellipticals only is not shown in this Figure. The spiral galaxies can be seen not following 
  this correlation. 
  Concerns about the lack of a correlation for spiral galaxies are not correct, and have already been addressed in 
  the last paragraph of Section 4.2. The correlation for spiral galaxies has been measured and quantified with 3 
  different linear regression routines. If there was no correlation for spiral galaxies, the uncertainties associated 
  to the slope and intercept of the linear regression would be extremely large.