### 2015-10-16 papers/data_paper/ApJ_submission_2/replytoreferee_I.txt
grampians

Dear Lee,

The referee's comments and suggestions have been very useful and helped improve the quality of the manuscript. 
We have addressed all the points listed in the reviewer's report. Below are our responses to the report. 
The title of the paper has been modified, as suggested by the referee, and all changes to the text have been printed 
in bold face.

Kind regards,

Giulia Savorgnan

RESPONSES TO THE REVIEWER'S REPORT:

- The tone of the title was not meant to be presumptuous, but we now understand that it is prone to misinterpretation, 
  therefore we have modified it. The new title reads: "Galaxy vivisection".

- We agree that some disk galaxies, like NGC 300, can be bulge-less. However, all galaxies in our sample are either 
  ellipticals/lenticulars or early-type spirals (Sa-Sc), with the only exception of NGC 4945 (Scd). 
  By definition, these galaxies have a spheroid/bulge component, which is what we modelled with a Sersic profile, 
  without attempting to distinguish between classical and pseudo-bulges. The text in Section 3.5 was actually misleading 
  and has been modified.
  Throughout the text, we used the terms "spheroid" and "bulge" as synonyms, although we recognize that some bulges are 
  "disk-like" rather than "spheroidal-like". This point has now been clarified in Footnote 1.

- We did not attempt to fit down- or up-bendings of large-scale disks. We scrutinized the available literature for all 
  individual galaxies, and we paid particular attention to the papers that recorded disk breaks. 
  The surface brightness profiles of galaxies that present this feature were truncated well before the occurrence of the 
  bending. This was not explicit in the text, and we have now added a sentence at the end of Section 3.1.

- The 1D equivalent axis fit does account for the varying ellipticity because the equivalent axis is calculated as 
  Req = Rmaj * (1-ellipticity)^0.5. 
  M87 has a strong ellipticity gradient. The 2D model cannot account for this gradient, and this can be seen in the 2D 
  residual image which looks like a "butterfly pattern". 
  Ideally, galaxies should be fit in 3D, rather than 2D or 1D. However, because the 1D fit tracks the changes in the 
  observed P.A. and ellipticity, the geometric mean axis (equivalent to a circularized axis) enables one to accurately 
  recover the total and component magnitudes in a way that the 2D model cannot.

- The text in Section 3.5 was not accurate, and we have now clarified this point. All bars have been fit with a Ferrer function, 
  except for NGC 4151 where we preferred to use a Sersic model. Disk-like components embedded in the bulges/spheroids have 
  been fit with a low-n Sersic profile and, in a few cases (when the Sersic model was degenerate with the other model-
  components), with a Ferrer function. 
  Gadotti & De Souza (2006) report that flat bars are typical of earlier type spirals, whereas exponential bars are common 
  in later type spirals. This is in agreement with our findings: our galaxy sample lacks of later-type spirals and we did 
  not detect any bar with an exponential profile.

- Sheth et al. (2010, the paper that presents the S4G collaboration) refers to Jun & Im (2008) to point out that the 3.6 um
  band is indeed a better tracer of the stellar mass than the K-band, which still suffers from dust absorption (Driver et al.
  2008, 678, L101).

- In Figure 7 we compared the 2D Sersic indices with the 1D equivalent-axis Sersic indices. These two quantities are not 
  expected to be exactly the same, but they should be approximatively similar. There are four most obvious outliers in this 
  plot: NGC 821, NGC 3377, NGC 4945, and NGC 5576. Their individual cases are discussed in the Appendix. 
  For M87, the 2D Sersic index (4.3) and the 1D equivalent-axis Sersic index (5.9) are in fairly good agreement. 
  The large discrepancy between the M87's 1D major-axis Sersic index and the other two measurements is due to the strong ellipticity
  gradient that this galaxy has. In fact, the 2D model cannot account for such gradient (whereas the 1D equivalent-axis fit can) 
  and the 2D residuals show a significant ``butterfly'' feature caused by this well-understood problem.

- The aim of this work is to investigate scaling relations between the black hole mass, Mbh, and several properties, X, of the host 
  spheroid/bulge. These correlations are commonly expressed as power laws: Mbh \propto X^a, i.e. log(Mbh) = a*log(X) + b. 
  Therefore we are interested in the logarithmic values of the Sersic indices and effective radii, along with their associated 
  uncertainties. 
  In addition, the Sersic index is the "exponent" in the Sersic equation, and that one needs to deal with the log of this quantity.  
  A fixed change in n is very different depending on the value of n, for example 2+/-1 results in a much greater change to the profile 
  than 10+/-1.

- The point that Footnote 2 was trying to make is the following. Sani et al., Vika et al. and Beifiori et al. did not find a 
  correlation between black hole mass and Sersic index. However, after averaging their Sersic indices together, a strong correlation
  was recovered. This is telling us that the measurements obtained by the individual authors were "noisy", i.e. non accurate. 
  Averaging the Sersic indices together reduces the noise and shows that the black hole mass is indeed correlated to the Sersic 
  index. However, we removed the second sentence of Footnote 2; we will come back to this point in a future paper.

- The claim about the anti-truncation of NGC 4459 has been clarified in the Appendix.

- Thanks for pointing out the missing reference to Erwin & Gadotti (2012), which has now been added.

- The Spitzer cookbook does not explain why the overlap correction is not recommended for 3.6 um images, it just states so. 
  It might be that the cookbook does not consider mosaicking images belonging to different "Astronomical Observation Requests", 
  for which we found that the overlap correction was needed to remove the chess-board effect.

- We did check that the PSF does not significantly vary across the various mosaics. In each mosaic, we selected up to 10 bright 
  non-saturated stars and we characterized them with a Moffat function (using IRAF imexamine). For each mosaic, we averaged 
  together the best-fit parameters of the stars and obtained an average FWHM and beta (the PSF parameters specific of each mosaic). 
  The variation of the Moffat FWHM across all mosaics is +-0''.1; the variation of the Moffat beta is +-2.0.
  The use of a universal PSF is justified for the following reasons: 
  i) the PSF-convolution is way more sensitive to the value of the FWHM than beta, therefore having a 50% variation in beta is 
  not an issue as long as the variation in FWHM is small;
  ii) the use of a non-signal-to-noise-weighted fitting scheme minimizes biases from a non-accurate PSF description;
  iii) not all mosaics have enough stars suitable for PSF characterization; rather than having an individual PSF for each mosaic 
  (which would have been based on only 1-2 stars for some mosaics), we preferred using a universal PSF;
  iv) the IRAC instrument handbook guarantees a pretty stable PSF across time.

- A reference to Gadotti (2008) has been added in Section 3.5.

- Regarding the comment on weighting schemes in Section 3.3, we attach two simple, but hopefully illustrative figures. 
  In both figures, the left hand panel shows a perfect R^(1/4) profile.
  In one figure no weighting scheme is used, while in the other a signal-to-noise-weighting scheme is used. 
  The results of the fit are the same. 
  Now in the right hand-panel, exactly the same partially depleted core has been introduced in both figures.  
  When the R^(1/4) model is fit with the different weighting schemes, the signal-to-noise-weighted scheme can be seen to perform worse.  
  In fact, any central deviation from the assumed model (in this case an R^(1/4) model, but we could have used a Sersic model) 
  results in the signal-to-noise-weighted fitting scheme performing worse.  
  Such deviations arise from: additional star clusters, dust, depleted cores, AGN, non accurate PSF, etc.  
  Unless one correctly accounts for all of these things (which realistically doesn't happen), then the signal-to-noise-weighted scheme 
  performs worse than a scheme with no weighting when it comes to matching the global light profile.

- References to Sheth et al. and Kim et al. have been added.

- The contrast in Figure 5 was indeed not appropriate and has been adjusted.

- Section 4.2.1, item 2: the typo has been corrected.

- Section 4.2.2: a sentence has been added to explain the different sigma symbols.

- The edge-on disc of NGC 1332 is clearly identified in the galaxy image. The ellipticity profile of the galaxy isophotes indicates 
  that the disc is intermediate-scale. We modified the caption of Figure 35 to better illustrate this point.

- The fifth paragraph of Section 3.5 explains how we decided to include a nuclear component in our models.

- We tried to model the bar of NGC 4945 with a Ferrer function or a Sersic model, but we did not obtain a satisfactory result. 
  We now realize that NGC 4945 is the only later type (Scd) spiral galaxy in our sample, and its bar might not be well described 
  by a Ferrer function. However, a Sersic model did not provide a good fit of the bar either. It is likely that the edge-on 
  inclination of the galaxy complicated the extraction of the surface brightness profile with ELLIPSE. 
  Our best solution is still to exclude the data in the range where the bar dominates the light and model only the bulge and the 
  large-scale disk. We note that we classified the fit of this galaxy as "grade 2".

- Kormendy & Ho (2013) did not perform bulge/disk decompositions for their galaxies, but they collected results from various sources 
  in the literature. We wanted to compare the results from our decompositions with the results from other systematic studies that 
  performed galaxy modelling in a consistent way.

And finally, I'd like to say thanks for the kind words at the beginning of the report. 
It is actually "refreshing" for myself to see that our effort is acknowledged by someone that fully understands the problematic 
involved in galaxy decomposition, so thanks a lot for the encouragement!