### 2016-02-24 papers/mbhn/replytoreferee.txt
referee report mbhn

I thank the referee for a prompt and insightful review that helped improve the quality of the manuscript. 
I have modified the manuscript as suggested by the referee, and highlighted the changes in boldface. 
Below is my response to the referee report.

Figures 1 and 2 have been modified to show the insets with error bars on the same scale as the main plots. 
I wish to maintain the fancy symbols for continuity with Paper II and because they allow to identify 
the two lenticular galaxies excluded from the linear regression analysis (orange crosses) and
the four galaxies classified as E/S0 or S0/Sp. 
If I was to show only the four-colour error bars, these six data points could not be easily located. 
I tried to combine fancy symbols and error bars in a single panel, but the result was a messy and 
too "crowded" plot. 
Hopefully a two-panel figure - with separate symbols and error bars - is a good compromise.

The linear regression for the "all" sample is now shown with a black dotted line in both Figures 1 and 2. 
Indeed, the Figures are now more informative. 
Looking at the bottom panel of Figure 1, one can notice that the blue and orange error bars 
span basically the same range in spheroid magnitude and Sérsic index, but they are offset from each other. 
In effect, the L-n linear regression for all galaxies seems to almost exactly define a separation line 
in between lenticular and spiral galaxies. 
Going beyond a "by eye" assessment, the linear regression analysis, which takes into account the error bars, 
confirms that the slopes of the correlations for early- and late-type galaxies are not consistent with each other
within their 1 sigma uncertainties. 
This is already a quantitative indication that early- and late-type galaxies define two separate L-n sequences, 
which justifies the use of a 4-parameter model, rather than a 2-parameter one, to describe the L-n diagram. 
I've also explored an independent method, drawn from Information Theory, to measure the relative quality of 
a 2-parameter and a 4-parameter model for the given L-n dataset. 
The Akaike Information Criterion rewards the goodness of a fit and penalises the number of parameter used 
in the model, therefore it provides a means for model selection. 
The Akaike Information Criterion test confirms that a 4-parameter model is preferred over a 2-parameter model 
to describe the L-n diagram, whereas a 2-parameter model is preferred over a 4-parameter model to describe 
the Mbh-n dataset. 
Details about the Akaike Information Criterion are given in the manuscript (Section 3.1).

Indeed, I agree with the referee's concern about the uncertain and sometimes misleading definition of "bulge" 
adopted in the literature, which led us to prefer the use of a more general term such as "spheroid" as opposed 
to disk, although we recognise that "spheroid" might not be fully appropriate to describe disk-like pseudobulges. 
In Papers I, II and III, as long as in Savorgnan & Graham (2016, MNRAS, 457, 320), we adopted and tried to be 
consistent with a point of view similar to that of the ATLAS3D team: local galaxies display a continuous distribution 
of spheroid-to-total flux ratios (e.g. Figure 2 of Cappellari et al. 2011), going from disc-less, "pure" 
elliptical galaxies (slow rotators) to disc-dominated lenticular and spiral galaxies (fast rotators). 
In Paper I, we consistently modelled each galaxy's spheroid with a Sérsic profile. Our identification of the spheroidal 
component can be classified as a "photometric" identification: we associated the spheroidal component with the central 
excess of light after the subtraction of discs, bars, nuclei, rings and disc-like components.

Since it was first presented in 2007 by Graham & Driver, the Mbh-n correlation did not receive enough attention 
over the following years. Vika et al. and Beifiori et al. failed to recover the correlation, whereas Sani et al. 
and Lasker et al. didn't even explore the Mbh-n diagram (and we checked that no strong Mbh-n correlation 
could be obtained using their datasets). 
In addition, none of these works explored the L-n diagram, whose more recent analysis was done by us (to our best knowledge) 
in Savorgnan et al. (2013) using the Graham & Guzman (2003) and Graham & Worley (2008) (outdated) datasets. 
For this reason, we felt that joining the analysis of the Mbh-n correlation with that of the more popular Mbh-L 
correlation would not have given enough credit to the former. 
At present, no simulations of black hole-galaxy coevolution take into account the observed Mbh-n correlation. 
A separate analysis and a dedicated manuscript would therefore give more visibility to this relation.
A separate analysis is also motivated by the following fact. The three linear regression routines 
(BCES, mFITEXY and linmix_err) all gave consistent results for the Mbh-L correlation. 
However, as explained in Section 3 of the manuscript, the modified FITEXY routine performs better than the BCES and 
linmix_err codes when it comes to the analysis of our L-n and Mbh-n datasets. 
To avoid confusion, we preferred to keep the analyses of the Mbh-L and Mbh-n relations separate.