### 2016-02-26 thesis/mn.tex

\chapter{$M_{\rm BH} - n_{\rm sph}$}
\label{ch:mn}

chapter x was dedicated to recovery of mn correlation with literature data

now we can do better with high quality data set

consistency with other correlations

good uncertainties -> robust estimate of intr scatter -> comparison with m-l

The remainder of this chapter comprises the published version of the paper 
``Supermassive Black Holes and Their Host Spheroids. 
III. The $M_{\rm BH} - n_{\rm sph}$ correlation'' 
by G.~A.~D.~Savorgnan,  
as it appears in Volume xxx of the \emph{The Astrophysical Journal}.

\includepdf[pages={1-10}]{mbhn.pdf}