### 2016-03-10 thesis/acknowledgement.tex
all chapters but concl

\chapter*{Acknowledgements}

alister, duncan, c blake, greg

karl,

gonzalo, rob, piero, glenn, anna, angela,

elisa, shaun, silvia, james

gym friends, rock climbing friends