### 2016-03-11 papers/MbhMsph/erratum/mm.tex
erratum

\documentclass{aastex6}

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

\begin{document}

\title{Erratum: ``Supermassive black holes and their host spheroids \\ 
II. The red and blue sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram'' (2016, ApJ, 817, 21)}

\author{Giulia A.~D.~Savorgnan and Alister W.~Graham}
\affil{Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia.}
\email{gsavorgn@astro.swin.edu.au}
\author{Alessandro Marconi}
\affil{Dipartimento di Fisica e Astronomia, Universit\'a di Firenze, via G. Sansone 1, I-50019 Sesto Fiorentino, Firenze, Italy.}
\and
\author{Eleonora Sani}
\affil{European Southern Observatory, Alonso de Cordova, Vitacura 3107, Santiago, Chile.}

\keywords{}

\section*{}

The header of column 5 in Table 4 is incorrectly given as $\langle MAG_{\rm sph} \rangle$. 
It should read $\langle M_{\rm *,sph} \rangle$.

\end{document}