### 2016-03-22 thesis/declaration.tex

\chapter*{Declaration}

I herewith declare that this thesis contains no material that has been accepted 
for the award to the candidate of any other degree or diploma. 
To the best of my knowledge, 
this thesis contains no material previously published or written by another author, 
except where due reference is made.
The work presented in this thesis has been carried out 
in the Centre for Astrophysics and Supercomputing 
at Swinburne University of Technology between September 2012 and April 2016. 
The content of the chapters listed below has appeared or will appear in refereed journals.

\begin{itemize}

\item Chapter \ref{ch:recov-mn} has been published as 
``The supermassive black hole mass -- S\'ersic index relations for bulges and elliptical galaxies''
in 2013, MNRAS, 434, 387, by Giulia A.~D.~Savorgnan et al.

\item Chapter \ref{ch:galviv} has been published as 
``Supermassive black holes and their host spheroids I. Disassembling galaxies'' 
in 2016, ApJS, 222, 10, by Giulia A.~D.~Savorgnan \& Alister W.~Graham.

\item Chapter \ref{ch:mm} has been published as 
``Supermassive black holes and their host spheroids 
II. The red and blue sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram''
in 2016, ApJ, 817, 21, by Giulia A.~D.~Savorgnan et al.

\item Chapter \ref{ch:mn} has been accepted for publication in \emph{The Astrophysical Journal} as 
``Supermassive black holes and their host spheroids 
III. The $M_{\rm BH} - n_{\rm sph}$ correlation''
by Giulia A.~D.~Savorgnan.

\item Chapter \ref{ch:msigma} has been published as 
``Overmassive black holes in the $M_{\rm BH} - \sigma$ diagram do not belong to over (dry) merged galaxies''
in 2015, MNRAS, 446, 2330, by Giulia A.~D.~Savorgnan \& Alister W.~Graham.

\item Chapter \ref{ch:ellic} has been published as 
``Explaining the reportedly over-massive black holes in early-type galaxies with intermediate-scale discs'' 
in 2016, MNRAS, 457, 320, by Giulia A.~D.~Savorgnan \& Alister W.~Graham.

\end{itemize}