### 2016-04-01 thesis/acknowledgement.tex
end

\chapter*{Acknowledgements}

First thanks go to my supervisor Alister, 
for offering me a PhD project in the country I always dreamt to live in, 
for guiding me during this long journey, 
for his patience and support, 
for all the lessons I've learnt, 
and for being constantly available to all of his students, 
regardless of how many late nights he had to spend in the office correcting my writings. 
Second thanks go to my co-supervisor Duncan for the constructive scientific conversations that we had, 
for his useful comments and suggestions, 
and for his genuine interest in making sure that at any stage everything was going well with me. 
Very special thanks are reserved to Chris Blake, Duncan, Greg, and Karl, 
for their invaluable, immediate, unconditionate support when I most than ever needed it.

%To my family and in particular to zia Anna and Suzy for being such enviable badasses 
%and such great examples. 
%
%To my PhD brothers and sisters, Angela, Anna, Gonzalo, Pierluigi, and Rob, 
%I'm glad I've been on this boat in your excellent company, guys. 
%Many thanks also to Glenn and Jeff for their informal and very comfortable attitude, 
%for always encouraging me to do my best and for all the good laughs we've had. 
%
%To my irreplaceable friends Elisa, Gonzalo, Greg $\heartsuit$ Anna, James, Shaun, Silvia, 
%thank you very much for being real. 
%
%And last, to Garry, because you might have glimpsed a tinder that already was inside me, 
%but you are doing a very good job at making it a blaze.