### 2016-06-28 thesis/publications.tex

\chapter*{List of publications}

In addition to my six lead-author papers that constitute the main body of this thesis, 
as a result of my PhD project, I participated in the following publications:

\begin{itemize}

\item Shankar, F., Bernardi, M., Sheth, R.~K., Ferrarese, L., Graham, A.~W.,
      {\bf Savorgnan, G.~A.~D.}, Allevato, V., Marconi, A., Laesker, R., Lapi, A. \\
      Selection bias in dynamically-measured super-massive black hole samples: 
      its consequences and the quest for the most fundamental relation \\
      Accepted for publication in \emph{MNRAS}. \\

\item Graham, A.~W., Durr\'e, M., {\bf Savorgnan, G.~A.~D.}, Medling, A.~M., 
      Batcheldor, D., Scott, N., Watson, B., Marconi, A. \\
      A Normal Supermassive Black Hole in NGC 1277 \\
      \emph{ApJ}, 819, 43, 2016. \\

\item Graham, A.~W., Dullo, B., {\bf Savorgnan, G.~A.~D.} \\
      Hiding in Plain Sight: An Abundance of Compact Massive Spheroids in the Local Universe \\
      \emph{ApJ}, 804, 32. 2015. \\

\end{itemize}